---
title: "geodata"
output: html_document
date: "2023-06-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) loading point data

```{r}
#install.packages("readxl")
library(readxl)

sheet_names = excel_sheets("../era5_data_2006_2020_thailand_monmean.xlsx")

#laden von den einzelnen sheets
T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
T.tp.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_tp_2006_2020_thailand_monm")
T.t2m.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_t2m_2006_2020_thailand_mon")

```

# a) Example 01: map with coordinate points of the centroids

```{r}
#assuming that longitude and latitude data in climate data are the centroids
centroids = T.huss.mon[,2:4] #including the names of the province to join the data later

# downloading the geo data (borders of districts)
library(sf)
geography <- st_read("/Users/leamrowiec/Desktop/Datenanalyse SS23/other data/gadm36_THA_shp")
```

```{r}
#install.packages("dplyr")
library(dplyr)

# zeilen der dataframe centroids nach der reihenfolge der Distrikte in geography ordnen
centroids.ordered <- arrange(centroids, ADM1_EN) 
#aber: namen der distrikte sind nicht immer gleich -> hinzufügen einer Nummerierung der Dsitrikte, um gleiche spalte zu bekommen
centroids.ordered['numbers'] = c(1:77)
geography["numbers"] = c(1:77)


```

```{r}
#install.packages("dplyr")
library(dplyr)
```

```{r}
# Zeilen beider dataframes (centroids und geography) nach alphabetischer reihenfolge der provinzen ordnen
centroids.sorted = arrange(centroids, ADM1_EN) #geography ist schon nach dem alphabet geordnet

#namen der distrikte sind nicht immer gleich -> hinzufügen einer Nummerierung der Dsitrikte, um gleiche spalte zu bekommen

centroids.sorted['numbers'] = c(1:77)
geography["numbers"] = c(1:77)


```

```{r}
# combining the two dataframes via the common column "numbers"
centroids.geography = merge(centroids.sorted, geography, by= "numbers")
# verview of data structure
class(centroids.geography)
head(centroids.geography)
summary(centroids.geography)
str(centroids.geography)

```

```{r}
#install.packages("ggplot2")
library(ggplot2)
library(sf)
```

```{r}
#install.packages("rgdal")
#install.packages("data.table")

library(rgdal)
library(data.table)
```

```{r}
ggplot() +
  geom_sf(data = centroids.geography, aes(geometry=geometry),col = "black", fill = NA) +  # adds geometric layer of polygons stored in the data object
  geom_point(data = centroids.geography, aes(x = Longitude, y = Latitude)) # adds coordinate points
  
```

# b) Example 02: Map with data plotted over district area

```{r}
# plotting polygon data with ggplot2
# geography is dataframe with polygon data
class(geography)

# to plot the data we need a new dataframe combining geography data and relative humidity data

# 1. putting lines of dataframe T.huss.mon in alphabetical order of the province names
library(dplyr)
T.huss.mon.ordered = arrange(T.huss.mon, ADM1_EN) 

# 2. changing numers in column "Number"
#T.huss.mon.ordered = subset(T.huss.mon.ordered, select = -Number)
T.huss.mon.ordered['numbers'] = c(1:77)
# geography["numbers"] = c(1:77); numbers have already been added


```

```{r}

# using mean data of each province as a mean over all reported months
T.huss.mon.mean = apply(T.huss.mon.ordered[-c(1,2)], 1, mean)
T.huss.mon.mean.df = as.data.frame(T.huss.mon.mean)

# adding another column of numberst to merge the dataframes
T.huss.mon.mean.df['numbers'] = c(1:77)

# 3. merging the two dataframes of relative humidtiy and geography
huss.geography = merge(T.huss.mon.mean.df, geography, by= "numbers")


```

```{r}
ggplot() +
  geom_sf(data = huss.geography,aes(geometry = geometry, fill= T.huss.mon.mean)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "mean relative humidity")
```

```{r}
# using max data of each province as a mean over all reported months
T.huss.mon.max = apply(T.huss.mon.ordered[-c(1,2)], 1, max)
T.huss.mon.max.df = as.data.frame(T.huss.mon.max)

# adding another column of numberst to merge the dataframes
T.huss.mon.max.df['numbers'] = c(1:77)

# merging the two dataframes of relative humidtiy and geography
huss.geography.max = merge(T.huss.mon.max.df, geography, by= "numbers")

# removing non-necessary colums of huss.geography.max
huss.geography.max.rmv = huss.geography.max[-c(3,4),]

# umbenennen der Spalte "T.huss.mon.max" in der dataframe "huss.geography.max.rmv"
huss.goegraphy.max.rmv.1=rename(huss.geography.max.rmv, mon.max= T.huss.mon.max)

# only plotting the map
ggplot(huss.goegraphy.max.rmv.1, aes(geometry = geometry)) +geom_sf()

#
install.packages("colorspace")
library(colorspace)
library(ggplot2)

ggplot(huss.goegraphy.max.rmv.1, aes(geometry=geometry, fill = mon.max)) +
  geom_sf () +
  colorspace::scale_fill_continuous_sequential(palette= "viridis")

ggplot() +
  geom_sf(data = huss.goegraphy.max.rmv.1,aes(geometry = geometry, fill= mon.max)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "mean relative humidity")

ggplot(huss.goegraphy.max.rmv.1, aes(geometry=geometry)) +
  geom_sf (aes(fill = mon.max)) +
  colorspace::scale_fill_continuous_sequential(palette= "viridis")

# wieso geht es immernoch nicht? sieht aus, als ob es latitude plottet, obwohl nicht mehr im dataframe???
print(huss.goegraphy.max.rmv.1)
```

plotting a geographic plot of the minimum temperatures

```{r}

T.huss.mon.min = apply(T.huss.mon.ordered[-c(1,2)], 1, min)
T.huss.mon.min.df = as.data.frame(T.huss.mon.min)

# adding another column of numbers to merge the dataframes
T.huss.mon.min.df['numbers'] = c(1:77)

# 3. merging the two dataframes of relative humidtiy and geography
huss.geography.min = merge(T.huss.mon.min.df, geography, by= "numbers")

# umbenennen der Spalte "T.huss.mon.min" in der dataframe "huss.geography.min"
library(dplyr)
huss.goegraphy.min.1 = rename(huss.geography.min, mon.min= T.huss.mon.min)

library(viridis)
library(ggplot2)

ggplot(huss.goegraphy.min.1, aes(geometry=geometry)) +
  geom_sf (aes(fill = mon.min)) +
  colorspace::scale_fill_continuous_sequential(palette= "viridis")+
  labs(title = "minimum relative humidity (2006-2020)")

```

```{r}
T.huss.mon.median = apply(T.huss.mon.ordered[-c(1,2)], 1, median)
T.huss.mon.median.df = as.data.frame(T.huss.mon.median)

# adding another column of numbers to merge the dataframes
T.huss.mon.median.df['numbers'] = c(1:77)

# 3. merging the two dataframes of relative humidtiy and geography
huss.geography.median = merge(T.huss.mon.median.df, geography, by= "numbers")

# umbenennen der Spalte "T.huss.mon.min" in der dataframe "huss.geography.min"
library(dplyr)
huss.goegraphy.median.1 = rename(huss.geography.median, mon.median= T.huss.mon.median)

library(viridis)
library(ggplot2)

ggplot(huss.goegraphy.median.1, aes(geometry=geometry)) +
  geom_sf (aes(fill = mon.median)) +
  colorspace::scale_fill_continuous_sequential(palette= "viridis")
  
```

## c) geo plot of mean relative humidity of each month

```{r}
install.packages("RColorBrewer")
#install.packages("geofacet")
#install.packages("magrittr")
#install.packages("ggplot2")

library(RColorBrewer)
library(geofacet)
library(ggplot2)
library(dplyr)
library(magrittr)

# Ziel: facet plot mit einzelnen Bannern, die jeweils die mittlere relative humidität über alle jahre in einem bestimmten Monat zeigen

# 1. Erstellen einer dataframe, die alle mittelwerte der jeweiligen Monate für jede Provinz beinhaltet

jan = c("2006.01.", "2007.01.", "2008.01.", "2009.01.", "2010.01.", "2011.01.", "2012.01.", "2013.01.", "2014.01.", "2015.01.", "2016.01.", "2017.01.", "2018.01." , "2019.01.", "2020.01.")

feb = c("2006.02.", "2007.02.", "2008.02.", "2009.02.", "2010.02.", "2011.02.", "2012.02.", "2013.02.", "2014.02.", "2015.02.", "2016.02.", "2017.02.", "2018.02." , "2019.02.", "2020.02.")

mar = c("2006.03.", "2007.03.", "2008.03.", "2009.03.", "2010.03.", "2011.03.", "2012.03.", "2013.03.", "2014.03.", "2015.03.", "2016.03.", "2017.03.", "2018.03." , "2019.03.", "2020.03.")

apr = c("2006.04.", "2007.04.", "2008.04.", "2009.04.", "2010.04.", "2011.04.", "2012.04.", "2013.04.", "2014.04.", "2015.04.", "2016.04.", "2017.04.", "2018.04." , "2019.04.", "2020.04.")

mai = c("2006.05.", "2007.05.", "2008.05.", "2009.05.", "2010.05.", "2011.05.", "2012.05.", "2013.05.", "2014.05.", "2015.05.", "2016.05.", "2017.05.", "2018.05." , "2019.05.", "2020.05.")

jun = c("2006.06.", "2007.06.", "2008.06.", "2009.06.", "2010.06.", "2011.06.", "2012.06.", "2013.06.", "2014.06.", "2015.06.", "2016.06.", "2017.06.", "2018.06." , "2019.06.", "2020.06.")

jul = c("2006.07.", "2007.07.", "2008.07.", "2009.07.", "2010.07.", "2011.07.", "2012.07.", "2013.07.", "2014.07.", "2015.07.", "2016.07.", "2017.07.", "2018.07." , "2019.07.", "2020.07.")

aug = c("2006.08.", "2007.08.", "2008.08.", "2009.08.", "2010.08.", "2011.08.", "2012.08.", "2013.08.", "2014.08.", "2015.08.", "2016.08.", "2017.08.", "2018.08." , "2019.08.", "2020.08.")

sep = c("2006.09.", "2007.09.", "2008.09.", "2009.09.", "2010.09.", "2011.09.", "2012.09.", "2013.09.", "2014.09.", "2015.09.", "2016.09.", "2017.09.", "2018.09." , "2019.09.", "2020.09.")

oct = c("2006.10.", "2007.10.", "2008.10.", "2009.10.", "2010.10.", "2011.10.", "2012.10.", "2013.10.", "2014.10.", "2015.10.", "2016.10.", "2017.10.", "2018.10." , "2019.10.", "2020.10.")

nov = c("2006.11.", "2007.11.", "2008.11.", "2009.11.", "2010.11.", "2011.11.", "2012.11.", "2013.11.", "2014.11.", "2015.11.", "2016.11.", "2017.11.", "2018.11." , "2019.11.", "2020.11.")

dec = c("2006.12.", "2007.12.", "2008.12.", "2009.12.", "2010.12.", "2011.12.", "2012.12.", "2013.12.", "2014.12.", "2015.12.", "2016.12.", "2017.12.", "2018.12." , "2019.12.", "2020.12.")

T.huss.mon.ordered.df = as.data.frame(T.huss.mon.ordered[,-1]) #sicherstellen, dass es sich um dataframe handelt

show(T.huss.mon.ordered)

# definieren von einzelnen dataframes zu jedem Monat, in denen die Mittelwerte jedes Monats für jede Provinz sind
means.jan = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,jan])})
means.feb = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,feb])})
means.mar = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,mar])})
means.apr = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,apr])})
means.mai = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,mai])})
means.apr = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,apr])})
means.jun = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,jun])})
means.jul = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,jul])})
means.aug = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,aug])})
means.sep = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,sep])})
means.oct = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,oct])})
means.nov = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,nov])})
means.dec = apply(T.huss.mon.ordered.df, 1, function(x) {rowMeans(T.huss.mon.ordered.df[,dec])})

library(dplyr)

# erstellen eines Vectors aus den einzelnen werten (in jeder spalte einer zeile stehen die gleichen Werte)
means.jan.v = means.jan[,1]
means.feb.v = means.feb[,1]
means.mar.v = means.mar[,1]
means.apr.v = means.apr[,1]
means.mai.v = means.mai[,1]
means.jun.v = means.jun[,1]
means.jul.v = means.jul[,1]
means.aug.v = means.aug[,1]
means.sep.v = means.sep[,1]
means.oct.v = means.oct[,1]
means.nov.v = means.nov[,1]
means.dec.v = means.dec[,1]

#zusammenfügen der geometry-daten, provinzennummern und provinzennamen mit den Mittelwerten der einzelnen monaten zu einzelnen dataframes zu jedem monat
new_cols = T.huss.mon.ordered.df[, c("ADM1_EN","numbers")]

means.jan.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], jan = means.jan.v)

means.feb.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], feb = means.feb.v)

means.mar.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], mar = means.mar.v)

means.apr.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], apr = means.apr.v)

means.mai.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], mai = means.mai.v)

means.jun.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], jun = means.jun.v)

means.jul.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], jul = means.jul.v)

means.aug.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], aug = means.aug.v)

means.sep.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], sept = means.sep.v)

means.oct.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], oct = means.oct.v)

means.nov.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], nov = means.nov.v)

means.dec.df = cbind(ADM1_EN = T.huss.mon.ordered.df[,"ADM1_EN"], numbers = T.huss.mon.ordered.df[,"numbers"], dec = means.dec.v)

jan.combined = merge(means.jan.df, geography, by= "numbers")
feb.combined = merge(means.feb.df, geography, by= "numbers")
mar.combined = merge(means.mar.df, geography, by= "numbers")
apr.combined = merge(means.apr.df, geography, by= "numbers")
mai.combined = merge(means.mai.df, geography, by= "numbers")
jun.combined = merge(means.jun.df, geography, by= "numbers")
jul.combined = merge(means.jul.df, geography, by= "numbers")
aug.combined = merge(means.aug.df, geography, by= "numbers")
sep.combined = merge(means.sep.df, geography, by= "numbers")
oct.combined = merge(means.oct.df, geography, by= "numbers")
nov.combined = merge(means.nov.df, geography, by= "numbers")
dec.combined = merge(means.dec.df, geography, by= "numbers")

```

### geoplot january to december

```{r}
library(dplyr)
library(colorspace)
library(viridis)
library(ggplot2)

#PLOT JANUARY---------------------------------------------------------------------------------

#definieren als numeric
class(jan.combined$jan)
jan.num = as.numeric(jan.combined$jan)
jan.combined$jan = jan.num

#plotten
ggplot(jan.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = jan.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of jauary (2006 to 2020)", fill = "mean relative humidity")
#purp geht auch als Farbpalette

#PLOT FEBUARY--------------------------------------------------------------------------------

#definieren als numeric
class(feb.combined$feb)
feb.num = as.numeric(feb.combined$feb)
feb.combined$feb = feb.num

#plotten
ggplot(feb.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = feb.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of february (2006 to 2020)", fill = "mean relative humidity")

#PLOT MARCH-----------------------------------------------------------------------------------

#definieren als numeric
class(mar.combined$mar)
mar.num = as.numeric(mar.combined$mar)
mar.combined$mar = mar.num

#plotten
ggplot(mar.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = mar.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")
  labs(title = "mean relative humidity of march (2006 to 2020)", fill = "mean relative humidity")

#PLOT APRIL-----------------------------------------------------------------------------------
  
  #definieren als numeric
class(apr.combined$apr)
apr.num = as.numeric(apr.combined$apr)
apr.combined$apr = apr.num

#plotten
ggplot(apr.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = apr.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of april (2006 to 2020)", fill = "mean relative humidity")

#PLOT MAI-------------------------------------------------------------------------------------
  
#definieren als numeric
class(mai.combined$mai)
mai.num = as.numeric(mai.combined$mai)
mai.combined$mai = mai.num

#plotten
ggplot(mai.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = mai.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of mai (2006 to 2020)", fill = "mean relative humidity")

#PLOT JUNE------------------------------------------------------------------------------------

#definieren als numeric
class(jun.combined$jun)
jun.num = as.numeric(jun.combined$jun)
jun.combined$jun = jun.num

#plotten
ggplot(jun.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = jun.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of june (2006 to 2020)", fill = "mean relative humidity")

#PLOT JULI------------------------------------------------------------------------------------

#definieren als numeric
class(jul.combined$jul)
jul.num = as.numeric(jul.combined$jul)
jul.combined$mai = jul.num

#plotten
ggplot(jul.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = jul.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of juli (2006 to 2020)", fill = "mean relative humidity")
#PLOT AUGUST-----------------------------------------------------------------------------------

#definieren als numeric
class(aug.combined$aug)
aug.num = as.numeric(aug.combined$aug)
aug.combined$mai = mai.num

#plotten
ggplot(aug.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = aug.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of august (2006 to 2020)",fill = "mean relative humidity")

#PLOT SEPTEMBER--------------------------------------------------------------------------------

#definieren als numeric
class(sep.combined$sep)
sep.num = as.numeric(sep.combined$sep)
sep.combined$sep = sep.num

#plotten
ggplot(sep.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = sep.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of september (2006 to 2020)", fill = "mean relative humidity")

#PLOT OCTOBER---------------------------------------------------------------------------------

#definieren als numeric
class(oct.combined$oct)
oct.num = as.numeric(oct.combined$oct)
oct.combined$oct = oct.num

#plotten
ggplot(oct.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = oct.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of october (2006 to 2020)",fill = "mean relative humidity")

#PLOT NOVEMBER--------------------------------------------------------------------------------

#definieren als numeric
class(nov.combined$nov)
nov.num = as.numeric(nov.combined$nov)
nov.combined$nov = nov.num

#plotten
ggplot(nov.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = nov.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of november (2006 to 2020)",fill = "mean relative humidity")

#PLOT DECEMBER---------------------------------------------------------------------------------

#definieren als numeric
class(dec.combined$dec)
dec.num = as.numeric(dec.combined$dec)
dec.combined$dec= dec.num

#plotten
ggplot(dec.combined, aes(geometry=geometry)) +
  geom_sf (aes(fill = dec.num)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "mean relative humidity of december (2006 to 2020) ", fill = "mean relative humidity")

```

### anormality plots: boxplot of the total cases of every year

```{r}
print(T.huss.mon)
#defining a dataframe for each year (means of all months in one year)
T.huss.2006= T.huss.mon[,c(2,5:16)]
T.huss.2007= T.huss.mon[,c(2,17:28)]
T.huss.2008= T.huss.mon[,c(2,29:40)]
T.huss.2009= T.huss.mon[,c(2,41:52)]
T.huss.2010= T.huss.mon[,c(2,53:64)]
T.huss.2011= T.huss.mon[,c(2,65:76)]
T.huss.2012= T.huss.mon[,c(2,77:88)]
T.huss.2013= T.huss.mon[,c(2,89:100)]
T.huss.2014= T.huss.mon[,c(2,101:112)]
T.huss.2015= T.huss.mon[,c(2,113:124)]
T.huss.2016= T.huss.mon[,c(2,125:136)]
T.huss.2017= T.huss.mon[,c(2,137:148)]
T.huss.2018= T.huss.mon[,c(2,149:160)]
T.huss.2019= T.huss.mon[,c(2,161:172)]
T.huss.2020= T.huss.mon[,c(2,173:184)]

m.T.huss.2006= apply(T.huss.2006[,2:13], 1, mean)
m.T.huss.2007= apply(T.huss.2007[,2:13], 1, mean)
m.T.huss.2008= apply(T.huss.2008[,2:13], 1, mean)
m.T.huss.2009= apply(T.huss.2009[,2:13], 1, mean)
m.T.huss.2010= apply(T.huss.2010[,2:13], 1, mean)
m.T.huss.2011= apply(T.huss.2011[,2:13], 1, mean)
m.T.huss.2012= apply(T.huss.2012[,2:13], 1, mean)
m.T.huss.2013= apply(T.huss.2013[,2:13], 1, mean)
m.T.huss.2014= apply(T.huss.2014[,2:13], 1, mean)
m.T.huss.2015= apply(T.huss.2015[,2:13], 1, mean)
m.T.huss.2016= apply(T.huss.2016[,2:13], 1, mean)
m.T.huss.2017= apply(T.huss.2017[,2:13], 1, mean)
m.T.huss.2018= apply(T.huss.2018[,2:13], 1, mean)
m.T.huss.2019= apply(T.huss.2019[,2:13], 1, mean)
m.T.huss.2020= apply(T.huss.2020[,2:13], 1, mean)


#merging the dataframes 
m.T.huss= cbind(m.T.huss.2006, m.T.huss.2007,m.T.huss.2008,m.T.huss.2009, m.T.huss.2010, m.T.huss.2011, m.T.huss.2012,m.T.huss.2013,m.T.huss.2014,m.T.huss.2015,m.T.huss.2016,m.T.huss.2017,m.T.huss.2018,m.T.huss.2019,m.T.huss.2020)
m.T.huss = as.data.frame(m.T.huss)
colnames(m.T.huss) = c("mean of 2006", "mean of 2007", "mean of 2008", "mean of 2009", "mean of 2010", "mean of 2011", "mean of 2012", "mean of 2013", "mean of 2014", "mean of 2015", "mean of 2016", "mean of 2017", "mean of 2018", "mean of 2019", "mean of 2020")
print(m.T.huss)

# adding column with total dengue cases of each province
row.means= apply(m.T.huss, 1, mean)
length (row.means)

m.T.huss.tot= cbind("total mean of each province"= row.means, m.T.huss)
print(m.T.huss.tot)


#starting to plot

#install.packages("tidyverse")
#install.packages("hrbrthemes")
library(ggplot2)
library(tidyverse)
library(hrbrthemes)
library(viridis)
#install.packages("cowplot")
library(cowplot)

#just playing around
ggplot(m.T.huss.tot, aes(y=`mean of 2006`))+
  stat_boxplot(geom="errorbar")+
  scale_x_discrete()+
  geom_boxplot(outlier.color= "gray", outlier.size=2)+
  labs(title= "2006", y= "   ")
  coord_cartesian(ylim = c(75, 110))

plot.2006 <- ggplot(m.T.huss.tot, aes(y = `mean of 2006`)) +
  stat_boxplot(geom = "errorbar") +
  scale_x_discrete()+
  geom_boxplot(outlier.color = "gray", outlier.size = 2) +
  labs(title = "2006", y = "") +
  ylim(70, 90)

plot.2007
m.T.huss.tot$`mean of 2006`

#plot
plot.2006=ggplot(m.T.huss.tot, aes(y=m.T.huss.tot$`mean of 2006`))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2006, y=" ")+
        ylim(70, 90)
plot.2007=ggplot(m.T.huss.tot, aes(y=m.T.huss.tot$"mean of 2007"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2007, y=" ")
plot.2008=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2008"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2008,y=" ")
plot.2009=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2009"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2009,y=" ")
plot.2010=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2010"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2010,y=" ")
plot.2011=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2011"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2011,y=" ")
plot.2012=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2012"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2012,y=" ")
plot.2013=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2013"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2013,y=" ")
plot.2014=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2014"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2014,y=" ")
plot.2015=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2015"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2015,y=" ")
plot.2016=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2016"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2016,y=" ")
plot.2017=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2017"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2017,y=" ")
plot.2018=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2018"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2018,y=" ")
plot.2019=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2019"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2019,y=" ")
plot.2020=ggplot(m.T.huss.tot, aes(y=m.T.huss$"mean of 2020"))+
        stat_boxplot(geom="errorbar")+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        scale_x_discrete()+
        ylim(70, 90)+
        labs(title=2020,y=" ")

#combining the plots into one big plot makes problems -> arrangement in word dokument   
plot_grid(plot.2006, plot.2007, plot.2008, plot.2009, plot.2010, plot.2011, plot.2012, plot.2013,  nrow = 2, align = "h")

plot_grid(plot.2014, plot.2015, plot.2016, plot.2017, plot.2018, plot.2019, plot.2020, nrow=2, align="h")
 
#install.packages("patchwork")
#library(patchwork)
#plot1 = plot.2006 + plot.2007 + plot.2008 + plot.2009 + plot.2010 + plot.2011 
#plot1  

install.packages("gridExtra")
library(gridExtra)
grid.arrange(plot.2006, plot.2007, plot.2008, plot.2009, nrow=1)  
grid.arrange(plot.2010, plot.2011, plot.2012, plot.2013, nrow=1)  
grid.arrange(plot.2014, plot.2015, plot.2016, plot.2017, nrow=1)  
grid.arrange(plot.2018, plot.2019, plot.2020, nrow=1)
        
#overview of box plots
boxplot(m.T.huss.tot[,-1], 
        main="mean yearly relative humidity (2006-2020)",
        xlab= "year",
        ylab="mean relative humidity") 
 


```

### plot of relative humidity yearly around 2018 and 2009

these years show the highest median values of relative humidity. I want to compare the yearly pattern of relative humidity between the mentioned years and the surrounding years

```{r}
#loading necessary library
library(ggplot2)
library(dplyr)
#defining  data to plot
mean.mon.2009 = apply(T.huss.2009[,-1], 2, mean)
mean.mon.2008 = apply(T.huss.2008[,-1], 2, mean)
mean.mon.2010 = apply(T.huss.2010[,-1], 2, mean)
mean.mon.2018 = apply(T.huss.2018[,-1], 2, mean)
mean.mon.2017 = apply(T.huss.2017[,-1], 2, mean)
mean.mon.2019 = apply(T.huss.2019[,-1], 2, mean)
# combining data in tibbles
df1 = as_tibble(data.frame(mean.mon.2008, mean.mon.2009, mean.mon.2010))
df2 = as_tibble(data.frame(mean.mon.2017, mean.mon.2018, mean.mon.2019))
#creating the plots
df1%>%
  ggplot(aes(x=seq(1,12,1), y=mean.mon.2009,color ="2009"), linetype=1)+
  geom_line(aes(y=mean.mon.2009), linetype=1)+
  geom_point(aes(y=mean.mon.2009))+
  geom_line(y=mean.mon.2008, color="violet")+ 
  geom_point(aes(y=mean.mon.2008,color ="2008"), linetype=1)+
  geom_line(y=mean.mon.2010, color="orange")+
  geom_point(aes(y=mean.mon.2010,color ="2010"), linetype=1)+
  ylim(55,100)+
  ylab("mean monthly relative humidity [%]")+
  xlab("month")+
  ggtitle("monthly mean relative humidity of all provinces (2008, 2009, 2010)")+
  scale_color_manual(values = c("violet", "pink", "orange"))+
  scale_x_continuous(minor_breaks = seq(1, 12, 0.5))
  
df2%>%
  ggplot(aes(x=seq(1,12,1), y=mean.mon.2018,color ="2018"), linetype=1)+
  geom_line(aes(y=mean.mon.2018, ), linetype=1)+
  geom_point(aes(y=mean.mon.2018))+
  geom_line(y=mean.mon.2017, color="violet")+ 
  geom_point(aes(y=mean.mon.2017,color ="2017"), linetype=1)+
  geom_line(y=mean.mon.2019, color="orange")+
  geom_point(aes(y=mean.mon.2019,color ="2019"), linetype=1)+
  ylim(55,100)+
  ylab("mean monthly relative humidity [%]")+
  xlab("month")+
  ggtitle("monthly mean relative humidity of all provinces (2017, 2018, 2019)")+
  scale_color_manual(values = c("violet", "pink", "orange"))+
  scale_x_continuous(minor_breaks = seq(1, 12, 0.5))
```

## plot of

To understand the impact of relative humidity on different kinds of areas in Thailand I chose exemplary provinces with different characteristics (position in the country, height of land):

-   Maha-Sarakham (north-east), 100-200m above sea level

-   Chiang Mai (north to north-west), 1000-1500m above sea level

-   Kalasin (central), 0-100m above sea level

-   Chumphon

-   Songkhla (south), 100-200m above sea level

    *Diercke Weltatlas* (1st ed.). (2008). westermann

```{r}
# dataframe combining yearly mean of each province 

huss.means = as_tibble(data.frame(T.huss.mon.ordered.df[,"ADM1_EN"], means.jan.v, means.feb.v, means.mar.v, means.apr.v, means.mai.v, means.jun.v, means.jul.v, means.aug.v, means.sep.v, means.oct.v, means.nov.v, means.dec.v)

#colnames(huss.means.df)= T.huss.mon.ordered.df[,"ADM1_EN"]
#huss.means.df = as.data.frame(huss.means)
huss.means


#choosing the mentioned exemplary provinces and plotting them
huss.means.df%>%
  ggplot(aes(x=seq(1,12,1), y=huss.means[,24], color ="Maha Sarakham"), linetype=1)+
  geom_line(aes(y=huss.means[,24]), linetype=1)+
  geom_point(aes(y=huss.means[,24]))+
  geom_line(y=huss.means[,10], color="violet")+ 
  geom_point(aes(y=huss.means[,10],color ="Chiang Mai"), linetype=1)+
  geom_line(y=huss.means[,60], color="orange")+
  geom_point(aes(y=huss.means[,60],color ="Saraburi"), linetype=1)+
  geom_line(y=huss.means[,13], color="blue")+
  geom_point(aes(y=huss.means[,13],color ="Chumphon"), linetype=1)+
  geom_line(y=huss.means[,64], color="green")+
  geom_point(aes(y=huss.means[,64],color ="Songkhla"), linetype=1)+
  ylim(55,100)+
  ylab("mean monthly relative humidity [%]")+
  xlab("month")+
  ggtitle("monthly mean relative humidity of all provinces (Maha Sarakham,Chiang Mai, Saraburi, Chumphon, Songkhala)")+
  scale_color_manual(values = c("violet", "orange", "blue", "green"))+
  scale_x_continuous(minor_breaks = seq(1, 12, 0.5))
```
