---
title: "Dengue_Climate_Plots_New"
author: "Manon Manddernach"
date: "2023-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 0. Loading the Data 

```{r}
## Dengue Data 
dengue.norm <- subset(dengue.norm, select = setdiff(colnames(dengue.norm), c("X", "X.1","X2006.total", "X2007.total", "X2008.total","X2009.total", "X2010.total", "X2011.total","X2012.total", "X2013.total", "X2014.total","X2015.total", "X2016.total", "X2017.total","X2018.total", "X2019.total", "X2020.total")))
dengue.norm['numbers']=c(1:77)
# Replace Bungkan with Nong Khai Data 
dengue.norm[4,2:181] <- dengue.norm[35,2:181]
#dengue.norm <- replace(dengue.norm, is.na(dengue.norm), 0)
as.data.frame(dengue.norm)

## Humidity Data 
T.huss.mon.ordered <- read.csv("../geo_data/T.huss.mon_correct_provinces.csv")
T.huss.mon.ordered <- subset(T.huss.mon.ordered, select = -X)
T.huss.mon.ordered$Number <- c(1:77)
names(T.huss.mon.ordered)[names(T.huss.mon.ordered) == "Number"] <- "numbers"

## Geo Data 
shapefilePath <- "./geo_data/gadm36_THA_shp"
layers <- st_layers(shapefilePath)
geography <- st_read(shapefilePath, layer = "gadm36_THA_1")
numbers <- c(1:77)
geography <- cbind(geography,numbers)

```

```{r}
View(dengue.norm)
```


## 0.1 Preparing the Data 

```{r}
## Years and Months  

Years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)
Months <- c(1,2,3,4,5,6,7,8,9,10,11,12)


## Dengue Data Mean values

dengue.norm.2006.mean.absolute <- as.data.frame(apply(dengue.norm.2006.mean,1,mean))
dengue.norm.2007.mean.absolute <- as.data.frame(apply(dengue.norm.2007.mean,1,mean))
dengue.norm.2008.mean.absolute <- as.data.frame(apply(dengue.norm.2008.mean,1,mean))
dengue.norm.2009.mean.absolute <- as.data.frame(apply(dengue.norm.2009.mean,1,mean))
dengue.norm.2010.mean.absolute <- as.data.frame(apply(dengue.norm.2010.mean,1,mean))
dengue.norm.2011.mean.absolute <- as.data.frame(apply(dengue.norm.2011.mean,1,mean))
dengue.norm.2012.mean.absolute <- as.data.frame(apply(dengue.norm.2012.mean,1,mean))
dengue.norm.2013.mean.absolute <- as.data.frame(apply(dengue.norm.2013.mean,1,mean))
dengue.norm.2014.mean.absolute <- as.data.frame(apply(dengue.norm.2014.mean,1,mean))
dengue.norm.2015.mean.absolute <- as.data.frame(apply(dengue.norm.2015.mean,1,mean))
dengue.norm.2016.mean.absolute <- as.data.frame(apply(dengue.norm.2016.mean,1,mean))
dengue.norm.2017.mean.absolute <- as.data.frame(apply(dengue.norm.2017.mean,1,mean))
dengue.norm.2018.mean.absolute <- as.data.frame(apply(dengue.norm.2018.mean,1,mean))
dengue.norm.2019.mean.absolute <- as.data.frame(apply(dengue.norm.2019.mean,1,mean))
dengue.norm.2020.mean.absolute <- as.data.frame(apply(dengue.norm.2020.mean,1,mean))

Dengue.mean <- c(dengue.norm.2006.mean.absolute,
                 dengue.norm.2007.mean.absolute,
                 dengue.norm.2008.mean.absolute,
                 dengue.norm.2009.mean.absolute, 
                 dengue.norm.2010.mean.absolute,
                 dengue.norm.2011.mean.absolute,
                 dengue.norm.2012.mean.absolute, 
                 dengue.norm.2013.mean.absolute, 
                 dengue.norm.2014.mean.absolute, 
                 dengue.norm.2015.mean.absolute, 
                 dengue.norm.2016.mean.absolute,
                 dengue.norm.2017.mean.absolute, 
                 dengue.norm.2018.mean.absolute, 
                 dengue.norm.2019.mean.absolute,
                 dengue.norm.2020.mean.absolute)

Dengue.mean <- lapply(Dengue.mean, unlist)
Dengue.mean <- lapply(Dengue.mean, mean)

Dengue.mean <- list(
  `apply(dengue.norm.2006.mean, 1, mean)` = 5.570221,
  `apply(dengue.norm.2007.mean, 1, mean)` = 8.742932,
  `apply(dengue.norm.2008.mean, 1, mean)` = 11.76473,
  `apply(dengue.norm.2009.mean, 1, mean)` = 7.219884,
  `apply(dengue.norm.2010.mean, 1, mean)` = 14.7395,
  `apply(dengue.norm.2011.mean, 1, mean)` = 8.785111,
  `apply(dengue.norm.2012.mean, 1, mean)` = 9.955709,
  `apply(dengue.norm.2013.mean, 1, mean)` = 20.17376,
  `apply(dengue.norm.2014.mean, 1, mean)` = 5.596716,
  `apply(dengue.norm.2015.mean, 1, mean)` = 17.39642,
  `apply(dengue.norm.2016.mean, 1, mean)` = 9.211036,
  `apply(dengue.norm.2017.mean, 1, mean)` = 6.315424,
  `apply(dengue.norm.2018.mean, 1, mean)` = 10.47807,
  `apply(dengue.norm.2019.mean, 1, mean)` = 14.60038,
  `apply(dengue.norm.2020.mean, 1, mean)` = 14.0717
)

Dengue.mean <- data.frame(Code = names(Dengue.mean), Value = unlist(Dengue.mean))
Dengue.mean <- Dengue.mean$Value


## Humidity Data Mean values 


huss.2006<- as.data.frame(T.huss.mon.ordered[,5:16])
huss.2007<- as.data.frame(T.huss.mon.ordered[,17:28])
huss.2008<- as.data.frame(T.huss.mon.ordered[,29:40])
huss.2009<- as.data.frame(T.huss.mon.ordered[,41:52])
huss.2010<- as.data.frame(T.huss.mon.ordered[,53:64])
huss.2011<- as.data.frame(T.huss.mon.ordered[,65:76])
huss.2012<- as.data.frame(T.huss.mon.ordered[,77:88])
huss.2013<- as.data.frame(T.huss.mon.ordered[,89:100])
huss.2014<- as.data.frame(T.huss.mon.ordered[,101:112])
huss.2015<- as.data.frame(T.huss.mon.ordered[,113:124])
huss.2016<- as.data.frame(T.huss.mon.ordered[,125:136])
huss.2017<- as.data.frame(T.huss.mon.ordered[,137:148])
huss.2018<- as.data.frame(T.huss.mon.ordered[,149:150])
huss.2019<- as.data.frame(T.huss.mon.ordered[,151:162])
huss.2020<- as.data.frame(T.huss.mon.ordered[,163:174])




huss.2006.mean.absolute <- as.data.frame(apply(huss.2006.mean,1,mean))
huss.2007.mean.absolute <- as.data.frame(apply(huss.2007.mean,1,mean))
huss.2008.mean.absolute <- as.data.frame(apply(huss.2008.mean,1,mean))
huss.2009.mean.absolute <- as.data.frame(apply(huss.2009.mean,1,mean))
huss.2010.mean.absolute <- as.data.frame(apply(huss.2010.mean,1,mean))
huss.2011.mean.absolute <- as.data.frame(apply(huss.2011.mean,1,mean))
huss.2012.mean.absolute <- as.data.frame(apply(huss.2012.mean,1,mean))
huss.2013.mean.absolute <- as.data.frame(apply(huss.2013.mean,1,mean))
huss.2014.mean.absolute <- as.data.frame(apply(huss.2014.mean,1,mean))
huss.2015.mean.absolute <- as.data.frame(apply(huss.2015.mean,1,mean))
huss.2016.mean.absolute <- as.data.frame(apply(huss.2016.mean,1,mean))
huss.2017.mean.absolute <- as.data.frame(apply(huss.2017.mean,1,mean))
huss.2018.mean.absolute <- as.data.frame(apply(huss.2018.mean,1,mean))
huss.2019.mean.absolute <- as.data.frame(apply(huss.2019.mean,1,mean))
huss.2020.mean.absolute <- as.data.frame(apply(huss.2020.mean,1,mean))

Huss.mean <- c(huss.2006.mean.absolute,
               huss.2007.mean.absolute,
               huss.2008.mean.absolute,
               huss.2009.mean.absolute, 
               huss.2010.mean.absolute,
               huss.2011.mean.absolute,
               huss.2012.mean.absolute, 
               huss.2013.mean.absolute, 
               huss.2014.mean.absolute, 
               huss.2015.mean.absolute, 
               huss.2016.mean.absolute,
               huss.2017.mean.absolute,
               huss.2018.mean.absolute, 
               huss.2019.mean.absolute,
               huss.2020.mean.absolute)

Huss.mean <- lapply(Huss.mean,unlist)
Huss.mean <- lapply(Huss.mean,mean)

Huss.mean <- list(
  `$apply(huss.2006.mean, 1, mean)` = 74.8919,
  `$apply(huss.2007.mean, 1, mean)` = 56.53615,
  `$apply(huss.2008.mean, 1, mean)` = 57.60083,
  `$apply(huss.2009.mean, 1, mean)` = 57.32692,
  `$apply(huss.2010.mean, 1, mean)` = 57.13848,
  `$apply(huss.2011.mean, 1, mean)` = 58.08443,
  `$apply(huss.2012.mean, 1, mean)` = 57.97239,
  `$apply(huss.2013.mean, 1, mean)` = 57.40518,
  `$apply(huss.2014.mean, 1, mean)` = 57.16852,
  `$apply(huss.2015.mean, 1, mean)` = 56.8422,
  `$apply(huss.2016.mean, 1, mean)` = 56.88025,
  `$apply(huss.2017.mean, 1, mean)` = 58.00465,
  `$apply(huss.2018.mean, 1, mean)` = 55.85306,
  `$apply(huss.2019.mean, 1, mean)` = 57.80723,
  `$apply(huss.2020.mean, 1, mean)` = 55.99571
)

Huss.mean <- data.frame(Code = names(Huss.mean), Value = unlist(Huss.mean))
Huss.mean <- Huss.mean$Value


Dengue.Humidity.Plot <- data.frame(Years=Years,
                                   Dengue.mean = Dengue.mean,
                                   Huss.mean = Huss.mean
                                   )


```

## 0.2 Defining single Provinces 
```{r}

MahaSarakham <- data.frame(T.huss.mon.ordered[24, ])
MahaSarakhamD <- data.frame(dengue.norm[24, ])

ChiangMai <- data.frame(T.huss.mon.ordered[10, ])
ChiangMaiD <- data.frame(dengue.norm[10, ])

Kalasin <- data.frame(T.huss.mon.ordered[14, ])
KalasinD <- data.frame(dengue.norm[14, ])

Chumphon <- data.frame(T.huss.mon.ordered[13, ])
ChumphonD <- data.frame(dengue.norm[13, ])

Songkhla <- data.frame(T.huss.mon.ordered[64, ])
SongkhlaD <- data.frame(dengue.norm[64, ])

####

NakhonSiThammarat <- data.frame(T.huss.mon.ordered[31, ])
NakhonSiThammaratD <- data.frame(dengue.norm[31, ])

SuratThani <- data.frame(T.huss.mon.ordered[67, ])
SuratThaniD <- data.frame(dengue.norm[67, ])


##

Phayao <- data.frame(T.huss.mon.ordered[67, ])
PhayaoD <- data.frame(dengue.norm[67, ])


```


# 1. Geographic Provinces Diffrent Sea levels


To understand the connection between dengue cases and different kinds of areas in Thailand exemplary provinces with different characteristics (position in the country, height of land) were chosen:

-   Maha-Sarakham (north-east), 100-200m above sea level

-   Chiang Mai (north to north-west), 1000-1500m above sea level

-   Kalasin (central), 0-100m above sea level

-   Chumphon

-   Songkhla (south), 100-200m above sea level

    *Diercke Weltatlas* (1st ed.). (2008). westermann
    
    
## 1.1 Maha Sarakham
```{r}
## Creating Dataframes and Plotting GGPlots for Maha Sarakham
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(MahaSarakham[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(MahaSarakhamD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("MahaSarakham.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  MahaSarakham_year <- get(paste0("MahaSarakham.", year))
  MahaSarakham_year$Dengue <- as.numeric(MahaSarakham_year$Dengue)
  MahaSarakham_year$Humidity <- as.numeric(MahaSarakham_year$Humidity)
  
  correlation <- cor(MahaSarakham_year$Dengue, MahaSarakham_year$Humidity)
  
  plot <- ggplot(MahaSarakham_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Maha Sarakham", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}


```

## 1.2 Chiang Mai
```{r}
## Creating Dataframes and Plotting GGPlots for Chiang Mai
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(ChiangMai[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(ChiangMaiD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("ChiangMai.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  ChiangMai_year <- get(paste0("ChiangMai.", year))
  ChiangMai_year$Dengue <- as.numeric(ChiangMai_year$Dengue)
  ChiangMai_year$Humidity <- as.numeric(ChiangMai_year$Humidity)
  
  correlation <- cor(ChiangMai_year$Dengue, ChiangMai_year$Humidity)
  
  plot <- ggplot(ChiangMai_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Chiang Mai", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```

## 1.3 Kalasin 
```{r}

## Creating Dataframes and Plotting GGPlots for Kalasin
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Kalasin[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(KalasinD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Kalasin.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Kalasin_year <- get(paste0("Kalasin.", year))
  Kalasin_year$Dengue <- as.numeric(Kalasin_year$Dengue)
  Kalasin_year$Humidity <- as.numeric(Kalasin_year$Humidity)
  
  correlation <- cor(Kalasin_year$Dengue, Kalasin_year$Humidity)
  
  plot <- ggplot(Kalasin_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Kalasin", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}
```


## 1.4 Chumphom
```{r}
## Creating Dataframes and Plotting GGPlots for Chumphon
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Chumphon[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(ChumphonD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Chumphon.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Chumphon_year <- get(paste0("Chumphon.", year))
  Chumphon_year$Dengue <- as.numeric(Chumphon_year$Dengue)
  Chumphon_year$Humidity <- as.numeric(Chumphon_year$Humidity)
  
  correlation <- cor(Chumphon_year$Dengue, Chumphon_year$Humidity)
  
  plot <- ggplot(Chumphon_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Chumphon", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```

## 1.5 Songkhla 
```{r}
## Creating Dataframes and Plotting GGPlots for Songkhla
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Songkhla[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(SongkhlaD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Songkhla.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Songkhla_year <- get(paste0("Songkhla.", year))
  Songkhla_year$Dengue <- as.numeric(Songkhla_year$Dengue)
  Songkhla_year$Humidity <- as.numeric(Songkhla_year$Humidity)
  
  correlation <- cor(Songkhla_year$Dengue, Songkhla_year$Humidity)
  
  plot <- ggplot(Songkhla_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Songkhla", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```


# 2. Possible to be hit by the gulf of Thailand 

## 2.1 Nakhon Si Thammarat
```{r}
## Creating Dataframes and Plotting GGPlots for Nakhon Si Thammarat
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Nakhon.Si.Thammarat[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(Nakhon.Si.ThammaratD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Nakhon.Si.Thammarat.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Nakhon.Si.Thammarat_year <- get(paste0("Nakhon.Si.Thammarat.", year))
  Nakhon.Si.Thammarat_year$Dengue <- as.numeric(Nakhon.Si.Thammarat_year$Dengue)
  Nakhon.Si.Thammarat_year$Humidity <- as.numeric(Nakhon.Si.Thammarat_year$Humidity)
  
  correlation <- cor(Nakhon.Si.Thammarat_year$Dengue, Nakhon.Si.Thammarat_year$Humidity)
  
  plot <- ggplot(Nakhon.Si.Thammarat_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Nakhon Si Thammarat", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}
```



## 2.2 Surat Thani
```{r}
## Creating Dataframes and Plotting GGPlots for Surat Thani
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(SuratThani[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(SuratThaniD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("SuratThani.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Surat.Thani_year <- get(paste0("SuratThani.", year))
  Surat.Thani_year$Dengue <- as.numeric(Surat.Thani_year$Dengue)
  Surat.Thani_year$Humidity <- as.numeric(Surat.Thani_year$Humidity)
  
  correlation <- cor(Surat.Thani_year$Dengue, Surat.Thani_year$Humidity)
  
  plot <- ggplot(Surat.Thani_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Surat Thani", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```

## 2.3 Chumphon 
## 2.4 Songkhla 

# 3. Focusing in high Provinces 

```{r}
## Creating Dataframes and Plotting GGPlots for Phayao
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Phayao[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(PhayaoD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Phayao.", i + 2005), data_frames[[i]])
}

plots <- list()

for (year in years) {
  Phayao_year <- get(paste0("Phayao.", year))
  Phayao_year$Dengue <- as.numeric(Phayao_year$Dengue)
  Phayao_year$Humidity <- as.numeric(Phayao_year$Humidity)
  
  correlation <- cor(Phayao_year$Dengue, Phayao_year$Humidity)
  
  plot <- ggplot(Phayao_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Phayao", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```


