---
title: "Dengue_cases_mapping_clustering"
author: "Paula, Lea"
date: "2023-06-13"
output: html_document
---

# 1) loading the data

```{r}
#loading dengue data 
filePath = paste(getwd(), "dengue_data_com.csv",sep = "/")
combined_df <- read.csv("../dengue_data/dengue_data_com.csv")
print(combined_df)

#converting dengue cases into numbers
combined_df$dengue_cases <- as.integer(combined_df$dengue_cases)
print(combined_df)
```

```{r}
#replacing NA with 0
## Frage hier: sollen wir NA values als 0 annehmen oder lieber Mittelwert oder sonst irgendwie??
#combined_df[is.na(combined_df)] <- 0
#print(combined_df)
```

```{r}
#opening packages
library(sp)
library(sf)
library(terra)
library(raster)
library(ggplot2)
library(exactextractr)
```

```{r}
#loading geo data
library(sf)
geography <- st_read("/Users/leamrowiec/Desktop/Datenanalyse SS23/other data/gadm36_THA_shp")
```

# 2) basic plots

## 2.1) defining variables for each time span

```{r}
# the defined variables always refer to the chosen rows

tot_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Total")
tot_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Total")
tot_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Total")
tot_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Total")
tot_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Total")
## 77 provinces
tot_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Total")
tot_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Total")
tot_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Total")
tot_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Total")
tot_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Total")
tot_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Total")
tot_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Total")
tot_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Total")
tot_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Total")
tot_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Total")

jan_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Jan")
jan_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Jan")
jan_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Jan")
jan_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Jan")
jan_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Jan")
jan_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Jan")
jan_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Jan")
jan_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Jan")
jan_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Jan")
jan_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Jan")
jan_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Jan")
jan_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Jan")
jan_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Jan")
jan_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Jan")
jan_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Jan")

feb_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Feb")
feb_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Feb")
feb_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Feb")
feb_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Feb")
feb_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Feb")
feb_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Feb")
feb_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Feb")
feb_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Feb")
feb_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Feb")
feb_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Feb")
feb_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Feb")
feb_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Feb")
feb_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Feb")
feb_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Feb")
feb_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Feb")

mar_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Mar")
mar_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Mar")
mar_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Mar")
mar_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Mar")
mar_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Mar")
mar_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Mar")
mar_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Mar")
mar_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Mar")
mar_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Mar")
mar_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Mar")
mar_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Mar")
mar_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Mar")
mar_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Mar")
mar_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Mar")
mar_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Mar")

apr_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Apr")
apr_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Apr")
apr_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Apr")
apr_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Apr")
apr_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Apr")
apr_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Apr")
apr_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Apr")
apr_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Apr")
apr_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Apr")
apr_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Apr")
apr_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Apr")
apr_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Apr")
apr_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Apr")
apr_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Apr")
apr_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Apr")

may_2006 = which(combined_df$Year=="2006" & combined_df$Month == "May")
may_2007 = which(combined_df$Year=="2007" & combined_df$Month == "May")
may_2008 = which(combined_df$Year=="2008" & combined_df$Month == "May")
may_2009 = which(combined_df$Year=="2009" & combined_df$Month == "May")
may_2010 = which(combined_df$Year=="2010" & combined_df$Month == "May")
may_2011 = which(combined_df$Year=="2011" & combined_df$Month == "May")
may_2012 = which(combined_df$Year=="2012" & combined_df$Month == "May")
may_2013 = which(combined_df$Year=="2013" & combined_df$Month == "May")
may_2014 = which(combined_df$Year=="2014" & combined_df$Month == "May")
may_2015 = which(combined_df$Year=="2015" & combined_df$Month == "May")
may_2016 = which(combined_df$Year=="2016" & combined_df$Month == "May")
may_2017 = which(combined_df$Year=="2017" & combined_df$Month == "May")
may_2018 = which(combined_df$Year=="2018" & combined_df$Month == "May")
may_2019 = which(combined_df$Year=="2019" & combined_df$Month == "May")
may_2020 = which(combined_df$Year=="2020" & combined_df$Month == "May")

jun_2006 = which(combined_df$Year=="2006" & combined_df$Month == "June")
jun_2007 = which(combined_df$Year=="2007" & combined_df$Month == "June")
jun_2008 = which(combined_df$Year=="2008" & combined_df$Month == "June")
jun_2009 = which(combined_df$Year=="2009" & combined_df$Month == "June")
jun_2010 = which(combined_df$Year=="2010" & combined_df$Month == "June")
jun_2011 = which(combined_df$Year=="2011" & combined_df$Month == "June")
jun_2012 = which(combined_df$Year=="2012" & combined_df$Month == "June")
jun_2013 = which(combined_df$Year=="2013" & combined_df$Month == "June")
jun_2014 = which(combined_df$Year=="2014" & combined_df$Month == "June")
jun_2015 = which(combined_df$Year=="2015" & combined_df$Month == "June")
jun_2016 = which(combined_df$Year=="2016" & combined_df$Month == "June")
jun_2017 = which(combined_df$Year=="2017" & combined_df$Month == "June")
jun_2018 = which(combined_df$Year=="2018" & combined_df$Month == "June")
jun_2019 = which(combined_df$Year=="2019" & combined_df$Month == "June")
jun_2020 = which(combined_df$Year=="2020" & combined_df$Month == "June")

jul_2006 = which(combined_df$Year=="2006" & combined_df$Month == "July")
jul_2007 = which(combined_df$Year=="2007" & combined_df$Month == "July")
jul_2008 = which(combined_df$Year=="2008" & combined_df$Month == "July")
jul_2009 = which(combined_df$Year=="2009" & combined_df$Month == "July")
jul_2010 = which(combined_df$Year=="2010" & combined_df$Month == "July")
jul_2011 = which(combined_df$Year=="2011" & combined_df$Month == "July")
jul_2012 = which(combined_df$Year=="2012" & combined_df$Month == "July")
jul_2013 = which(combined_df$Year=="2013" & combined_df$Month == "July")
jul_2014 = which(combined_df$Year=="2014" & combined_df$Month == "July")
jul_2015 = which(combined_df$Year=="2015" & combined_df$Month == "July")
jul_2016 = which(combined_df$Year=="2016" & combined_df$Month == "July")
jul_2017 = which(combined_df$Year=="2017" & combined_df$Month == "July")
jul_2018 = which(combined_df$Year=="2018" & combined_df$Month == "July")
jul_2019 = which(combined_df$Year=="2019" & combined_df$Month == "July")
jul_2020 = which(combined_df$Year=="2020" & combined_df$Month == "July")

aug_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Aug")
aug_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Aug")
aug_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Aug")
aug_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Aug")
aug_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Aug")
aug_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Aug")
aug_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Aug")
aug_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Aug")
aug_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Aug")
aug_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Aug")
aug_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Aug")
aug_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Aug")
aug_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Aug")
aug_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Aug")
aug_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Aug")

sep_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Sept")
sep_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Sept")
sep_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Sept")
sep_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Sept")
sep_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Sept")
sep_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Sept")
sep_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Sept")
sep_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Sept")
sep_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Sept")
sep_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Sept")
sep_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Sept")
sep_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Sept")
sep_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Sept")
sep_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Sept")
sep_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Sept")

oct_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Oct")
oct_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Oct")
oct_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Oct")
oct_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Oct")
oct_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Oct")
oct_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Oct")
oct_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Oct")
oct_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Oct")
oct_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Oct")
oct_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Oct")
oct_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Oct")
oct_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Oct")
oct_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Oct")
oct_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Oct")
oct_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Oct")

nov_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Nov")
nov_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Nov")
nov_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Nov")
nov_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Nov")
nov_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Nov")
nov_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Nov")
nov_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Nov")
nov_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Nov")
nov_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Nov")
nov_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Nov")
nov_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Nov")
nov_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Nov")
nov_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Nov")
nov_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Nov")
nov_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Nov")

dec_2006 = which(combined_df$Year=="2006" & combined_df$Month == "Dec")
dec_2007 = which(combined_df$Year=="2007" & combined_df$Month == "Dec")
dec_2008 = which(combined_df$Year=="2008" & combined_df$Month == "Dec")
dec_2009 = which(combined_df$Year=="2009" & combined_df$Month == "Dec")
dec_2010 = which(combined_df$Year=="2010" & combined_df$Month == "Dec")
dec_2011 = which(combined_df$Year=="2011" & combined_df$Month == "Dec")
dec_2012 = which(combined_df$Year=="2012" & combined_df$Month == "Dec")
dec_2013 = which(combined_df$Year=="2013" & combined_df$Month == "Dec")
dec_2014 = which(combined_df$Year=="2014" & combined_df$Month == "Dec")
dec_2015 = which(combined_df$Year=="2015" & combined_df$Month == "Dec")
dec_2016 = which(combined_df$Year=="2016" & combined_df$Month == "Dec")
dec_2017 = which(combined_df$Year=="2017" & combined_df$Month == "Dec")
dec_2018 = which(combined_df$Year=="2018" & combined_df$Month == "Dec")
dec_2019 = which(combined_df$Year=="2019" & combined_df$Month == "Dec")
dec_2020 = which(combined_df$Year=="2020" & combined_df$Month == "Dec")

#tot_2006 = combined_df[1:76,]
#tot_2006 = combined_df[1:76,]
#jan_2006 = combined_df[77:152,]
#feb_2006 = combined_df[153:228,]
#mar_2006 = combined_df[229:304,]
#apr_2006 = combined_df[307:383]
  
```

## 

# 3) plotting mean dengue cases of each year of 2006 to 2020

## 3.1) defining new columns

the province Bueng Kan separated from Nong Khain in 2011 \--\> 2011 to 2020: 77 provinces, before: 76 provinces

```{r}
# setting dengue data cases of Bueng Kan and Nong Khain in 2006 to 2010 to "NA"

new_row_tot_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Total")
new_row_tot_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Total")
new_row_tot_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Total")
new_row_tot_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Total")
new_row_tot_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Total")

new_row_jan_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Jan")
new_row_feb_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Feb")
new_row_mar_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Mar")
new_row_apr_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Apr")
new_row_may_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "May")
new_row_jun_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "June")
new_row_jul_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "July")
new_row_aug_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Aug")
new_row_sep_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Sept")
new_row_oct_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Oct")
new_row_nov_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Nov")
new_row_dec_2006 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2006", Month= "Dec")


new_row_jan_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Jan")
new_row_feb_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Feb")
new_row_mar_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Mar")
new_row_apr_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Apr")
new_row_may_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "May")
new_row_jun_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "June")
new_row_jul_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "July")
new_row_aug_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Aug")
new_row_sep_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Sept")
new_row_oct_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Oct")
new_row_nov_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Nov")
new_row_dec_2007 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2007", Month= "Dec")


new_row_jan_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Jan")
new_row_feb_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Feb")
new_row_mar_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Mar")
new_row_apr_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Apr")
new_row_may_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "May")
new_row_jun_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "June")
new_row_jul_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "July")
new_row_aug_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Aug")
new_row_sep_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Sept")
new_row_oct_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Oct")
new_row_nov_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Nov")
new_row_dec_2008 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2008", Month= "Dec")


new_row_jan_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Jan")
new_row_feb_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Feb")
new_row_mar_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Mar")
new_row_apr_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Apr")
new_row_may_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "May")
new_row_jun_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "June")
new_row_jul_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "July")
new_row_aug_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Aug")
new_row_sep_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Sept")
new_row_oct_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Oct")
new_row_nov_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Nov")
new_row_dec_2009 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2009", Month= "Dec")

new_row_jan_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Jan")
new_row_feb_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Feb")
new_row_mar_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Mar")
new_row_apr_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Apr")
new_row_may_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "May")
new_row_jun_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "June")
new_row_jul_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "July")
new_row_aug_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Aug")
new_row_sep_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Sept")
new_row_oct_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Oct")
new_row_nov_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Nov")
new_row_dec_2010 = data.frame(X=0, Reporting_area= "Bueng Kan", dengue_cases= NA, Year ="2010", Month= "Dec")
```

## 3.2) plotting dengue cases of each month (mean of all years)

```{r}
#2006------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2006 = combined_df[tot_2006,]
# adding row of the new province
dengue.tot.2006.1 <- rbind(dengue.tot.2006[1:(4-1), ], new_row_tot_2006, dengue.tot.2006[4:nrow(dengue.tot.2006), ])

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2006, length(tot_2006)+1)
dengue.tot.2006.geo = merge(dengue.tot.2006.1, geography, by= "X")

ggplot(dengue.tot.2006.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2006", fill = "total reported dengue cases")

#2007------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2007 = combined_df[tot_2007,]
# adding row of the new province
dengue.tot.2007.1 <- rbind(dengue.tot.2007[1:(4-1), ], new_row_tot_2007, dengue.tot.2007[4:nrow(dengue.tot.2007), ])

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2007, length(tot_2007)+1)
dengue.tot.2007.geo = merge(dengue.tot.2007.1, geography, by= "X")

ggplot(dengue.tot.2007.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2007", fill = "total reported dengue cases")

#2008------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2008 = combined_df[tot_2008,]
# adding row of the new province
dengue.tot.2008.1 <- rbind(dengue.tot.2006[1:(4-1), ], new_row_tot_2008, dengue.tot.2008[4:nrow(dengue.tot.2008), ])

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2008, length(tot_2008)+1)
dengue.tot.2008.geo = merge(dengue.tot.2008.1, geography, by= "X")

ggplot(dengue.tot.2008.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2008", fill = "total reported dengue cases")

#2009------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2009 = combined_df[tot_2009,]
# adding row of the new province
dengue.tot.2009.1 <- rbind(dengue.tot.2009[1:(4-1), ], new_row_tot_2009, dengue.tot.2009[4:nrow(dengue.tot.2009), ])

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2009, length(tot_2009)+1)
dengue.tot.2009.geo = merge(dengue.tot.2009.1, geography, by= "X")

ggplot(dengue.tot.2009.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2009", fill = "total reported dengue cases")

#2010------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2010 = combined_df[tot_2010,]
# adding row of the new province
dengue.tot.2010.1 <- rbind(dengue.tot.2010[1:(4-1), ], new_row_tot_2010, dengue.tot.2010[4:nrow(dengue.tot.2010), ])

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2010, length(tot_2010)+1)
dengue.tot.2010.geo = merge(dengue.tot.2010.1, geography, by= "X")

ggplot(dengue.tot.2010.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2010", fill = "total reported dengue cases")










#2011------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2011 = combined_df[tot_2011,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2011)
dengue.tot.2011.geo = merge(dengue.tot.2011, geography, by= "X")

ggplot(dengue.tot.2011.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2011", fill = "total reported dengue cases")

#2012------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2012 = combined_df[tot_2012,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2012)
dengue.tot.2012.geo = merge(dengue.tot.2012, geography, by= "X")

ggplot(dengue.tot.2012.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2012", fill = "total reported dengue cases")

#2013------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2013 = combined_df[tot_2013,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2013)
dengue.tot.2013.geo = merge(dengue.tot.2013, geography, by= "X")

ggplot(dengue.tot.2013.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2013", fill = "total reported dengue cases")

#2014------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2014 = combined_df[tot_2014,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2014)
dengue.tot.2014.geo = merge(dengue.tot.2014, geography, by= "X")

ggplot(dengue.tot.2014.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2014", fill = "total reported dengue cases")

#2015------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2015 = combined_df[tot_2015,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2015)
dengue.tot.2015.geo = merge(dengue.tot.2015, geography, by= "X")

ggplot(dengue.tot.2015.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2015", fill = "total reported dengue cases")


#2016------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2016 = combined_df[tot_2016,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2016)
dengue.tot.2016.geo = merge(dengue.tot.2016, geography, by= "X")

ggplot(dengue.tot.2016.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2016", fill = "total reported dengue cases")

#2017------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2017 = combined_df[tot_2017,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2017)
dengue.tot.2017.geo = merge(dengue.tot.2017, geography, by= "X")

ggplot(dengue.tot.2017.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2017", fill = "total reported dengue cases")

#2018------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2018 = combined_df[tot_2018,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2018)
dengue.tot.2018.geo = merge(dengue.tot.2018, geography, by= "X")

ggplot(dengue.tot.2018.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2018", fill = "total reported dengue cases")

#2019------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2019 = combined_df[tot_2019,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2019)
dengue.tot.2019.geo = merge(dengue.tot.2019, geography, by= "X")

ggplot(dengue.tot.2019.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2019", fill = "total reported dengue cases")

#2020------------------------------------------------------------------------------------------
#coosing the lines that show total data for each district 2006 to 2020
dengue.tot.2020 = combined_df[tot_2020,]

# merging the dataframe with the geodata (geography is ordered in alphabetical order, dengue data too)
geography["X"] = c(tot_2020)
dengue.tot.2020.geo = merge(dengue.tot.2020, geography, by= "X")

ggplot(dengue.tot.2020.geo, aes(geometry=geometry)) +
  geom_sf (aes(fill = dengue_cases)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "total reported dengue cases in 2020", fill = "total reported dengue cases")

```

-\> Bankok is very dark because population density is very high -\> overlay with population density

# 4) plotting time series

## 4.1) plotting mean dengue cases of the whole country in a time series

preparing a dataframe containing mean value of each month over all provinces (2006 to 2020)

```{r}
print(combined_df)

mean_jan_2006 = mean(combined_df$dengue_cases[jan_2006])
mean_feb_2006 = mean(combined_df$dengue_cases[feb_2006])
mean_mar_2006 = mean(combined_df$dengue_cases[mar_2006])
mean_apr_2006 = mean(combined_df$dengue_cases[apr_2006])
mean_may_2006 = mean(combined_df$dengue_cases[may_2006])
mean_jun_2006 = mean(combined_df$dengue_cases[jun_2006])
mean_jul_2006 = mean(combined_df$dengue_cases[jul_2006])
mean_aug_2006 = mean(combined_df$dengue_cases[aug_2006])
mean_sep_2006 = mean(combined_df$dengue_cases[sep_2006])
mean_oct_2006 = mean(combined_df$dengue_cases[oct_2006])
mean_nov_2006 = mean(combined_df$dengue_cases[nov_2006])
mean_dec_2006 = mean(combined_df$dengue_cases[dec_2006])

mean_jan_2007 = mean(combined_df$dengue_cases[jan_2007])
mean_feb_2007 = mean(combined_df$dengue_cases[feb_2007])
mean_mar_2007 = mean(combined_df$dengue_cases[mar_2007])
mean_apr_2007 = mean(combined_df$dengue_cases[apr_2007])
mean_may_2007 = mean(combined_df$dengue_cases[may_2007])
mean_jun_2007 = mean(combined_df$dengue_cases[jun_2007])
mean_jul_2007 = mean(combined_df$dengue_cases[jul_2007])
mean_aug_2007 = mean(combined_df$dengue_cases[aug_2007])
mean_sep_2007 = mean(combined_df$dengue_cases[sep_2007])
mean_oct_2007 = mean(combined_df$dengue_cases[oct_2007])
mean_nov_2007 = mean(combined_df$dengue_cases[nov_2007])
mean_dec_2007 = mean(combined_df$dengue_cases[dec_2007])

mean_jan_2008 = mean(combined_df$dengue_cases[jan_2008])
mean_feb_2008 = mean(combined_df$dengue_cases[feb_2008])
mean_mar_2008 = mean(combined_df$dengue_cases[mar_2008])
mean_apr_2008 = mean(combined_df$dengue_cases[apr_2008])
mean_may_2008 = mean(combined_df$dengue_cases[may_2008])
mean_jun_2008 = mean(combined_df$dengue_cases[jun_2008])
mean_jul_2008 = mean(combined_df$dengue_cases[jul_2008])
mean_aug_2008 = mean(combined_df$dengue_cases[aug_2008])
mean_sep_2008 = mean(combined_df$dengue_cases[sep_2008])
mean_oct_2008 = mean(combined_df$dengue_cases[oct_2008])
mean_nov_2008 = mean(combined_df$dengue_cases[nov_2008])
mean_dec_2008 = mean(combined_df$dengue_cases[dec_2008])

mean_jan_2009 = mean(combined_df$dengue_cases[jan_2009])
mean_feb_2009 = mean(combined_df$dengue_cases[feb_2009])
mean_mar_2009 = mean(combined_df$dengue_cases[mar_2009])
mean_apr_2009 = mean(combined_df$dengue_cases[apr_2009])
mean_may_2009 = mean(combined_df$dengue_cases[may_2009])
mean_jun_2009 = mean(combined_df$dengue_cases[jun_2009])
mean_jul_2009 = mean(combined_df$dengue_cases[jul_2009])
mean_aug_2009 = mean(combined_df$dengue_cases[aug_2009])
mean_sep_2009 = mean(combined_df$dengue_cases[sep_2009])
mean_oct_2009 = mean(combined_df$dengue_cases[oct_2009])
mean_nov_2009 = mean(combined_df$dengue_cases[nov_2009])
mean_dec_2009 = mean(combined_df$dengue_cases[dec_2009])

mean_jan_2010 = mean(combined_df$dengue_cases[jan_2010])
mean_feb_2010 = mean(combined_df$dengue_cases[feb_2010])
mean_mar_2010 = mean(combined_df$dengue_cases[mar_2010])
mean_apr_2010 = mean(combined_df$dengue_cases[apr_2010])
mean_may_2010 = mean(combined_df$dengue_cases[may_2010])
mean_jun_2010 = mean(combined_df$dengue_cases[jun_2010])
mean_jul_2010 = mean(combined_df$dengue_cases[jul_2010])
mean_aug_2010 = mean(combined_df$dengue_cases[aug_2010])
mean_sep_2010 = mean(combined_df$dengue_cases[sep_2010])
mean_oct_2010 = mean(combined_df$dengue_cases[oct_2010])
mean_nov_2010 = mean(combined_df$dengue_cases[nov_2010])
mean_dec_2010 = mean(combined_df$dengue_cases[dec_2010])

mean_jan_2011 = mean(combined_df$dengue_cases[jan_2011])
mean_feb_2011 = mean(combined_df$dengue_cases[feb_2011])
mean_mar_2011 = mean(combined_df$dengue_cases[mar_2011])
mean_apr_2011 = mean(combined_df$dengue_cases[apr_2011])
mean_may_2011 = mean(combined_df$dengue_cases[may_2011])
mean_jun_2011 = mean(combined_df$dengue_cases[jun_2011])
mean_jul_2011 = mean(combined_df$dengue_cases[jul_2011])
mean_aug_2011 = mean(combined_df$dengue_cases[aug_2011])
mean_sep_2011 = mean(combined_df$dengue_cases[sep_2011])
mean_oct_2011 = mean(combined_df$dengue_cases[oct_2011])
mean_nov_2011 = mean(combined_df$dengue_cases[nov_2011])
mean_dec_2011 = mean(combined_df$dengue_cases[dec_2011])

mean_jan_2012 = mean(combined_df$dengue_cases[jan_2012])
mean_feb_2012 = mean(combined_df$dengue_cases[feb_2012])
mean_mar_2012 = mean(combined_df$dengue_cases[mar_2012])
mean_apr_2012 = mean(combined_df$dengue_cases[apr_2012])
mean_may_2012 = mean(combined_df$dengue_cases[may_2012])
mean_jun_2012 = mean(combined_df$dengue_cases[jun_2012])
mean_jul_2012 = mean(combined_df$dengue_cases[jul_2012])
mean_aug_2012 = mean(combined_df$dengue_cases[aug_2012])
mean_sep_2012 = mean(combined_df$dengue_cases[sep_2012])
mean_oct_2012 = mean(combined_df$dengue_cases[oct_2012])
mean_nov_2012 = mean(combined_df$dengue_cases[nov_2012])
mean_dec_2012 = mean(combined_df$dengue_cases[dec_2012])

mean_jan_2013 = mean(combined_df$dengue_cases[jan_2013])
mean_feb_2013 = mean(combined_df$dengue_cases[feb_2013])
mean_mar_2013 = mean(combined_df$dengue_cases[mar_2013])
mean_apr_2013 = mean(combined_df$dengue_cases[apr_2013])
mean_may_2013 = mean(combined_df$dengue_cases[may_2013])
mean_jun_2013 = mean(combined_df$dengue_cases[jun_2013])
mean_jul_2013 = mean(combined_df$dengue_cases[jul_2013])
mean_aug_2013 = mean(combined_df$dengue_cases[aug_2013])
mean_sep_2013 = mean(combined_df$dengue_cases[sep_2013])
mean_oct_2013 = mean(combined_df$dengue_cases[oct_2013])
mean_nov_2013 = mean(combined_df$dengue_cases[nov_2013])
mean_dec_2013 = mean(combined_df$dengue_cases[dec_2013])

mean_jan_2014 = mean(combined_df$dengue_cases[jan_2014])
mean_feb_2014 = mean(combined_df$dengue_cases[feb_2014])
mean_mar_2014 = mean(combined_df$dengue_cases[mar_2014])
mean_apr_2014 = mean(combined_df$dengue_cases[apr_2014])
mean_may_2014 = mean(combined_df$dengue_cases[may_2014])
mean_jun_2014 = mean(combined_df$dengue_cases[jun_2014])
mean_jul_2014 = mean(combined_df$dengue_cases[jul_2014])
mean_aug_2014 = mean(combined_df$dengue_cases[aug_2014])
mean_sep_2014 = mean(combined_df$dengue_cases[sep_2014])
mean_oct_2014 = mean(combined_df$dengue_cases[oct_2014])
mean_nov_2014 = mean(combined_df$dengue_cases[nov_2014])
mean_dec_2014 = mean(combined_df$dengue_cases[dec_2014])

mean_jan_2015 = mean(combined_df$dengue_cases[jan_2015])
mean_feb_2015 = mean(combined_df$dengue_cases[feb_2015])
mean_mar_2015 = mean(combined_df$dengue_cases[mar_2015])
mean_apr_2015 = mean(combined_df$dengue_cases[apr_2015])
mean_may_2015 = mean(combined_df$dengue_cases[may_2015])
mean_jun_2015 = mean(combined_df$dengue_cases[jun_2015])
mean_jul_2015 = mean(combined_df$dengue_cases[jul_2015])
mean_aug_2015 = mean(combined_df$dengue_cases[aug_2015])
mean_sep_2015 = mean(combined_df$dengue_cases[sep_2015])
mean_oct_2015 = mean(combined_df$dengue_cases[oct_2015])
mean_nov_2015 = mean(combined_df$dengue_cases[nov_2015])
mean_dec_2015 = mean(combined_df$dengue_cases[dec_2015])

mean_jan_2016 = mean(combined_df$dengue_cases[jan_2016])
mean_feb_2016 = mean(combined_df$dengue_cases[feb_2016])
mean_mar_2016 = mean(combined_df$dengue_cases[mar_2016])
mean_apr_2016 = mean(combined_df$dengue_cases[apr_2016])
mean_may_2016 = mean(combined_df$dengue_cases[may_2016])
mean_jun_2016 = mean(combined_df$dengue_cases[jun_2016])
mean_jul_2016 = mean(combined_df$dengue_cases[jul_2016])
mean_aug_2016 = mean(combined_df$dengue_cases[aug_2016])
mean_sep_2016 = mean(combined_df$dengue_cases[sep_2016])
mean_oct_2016 = mean(combined_df$dengue_cases[oct_2016])
mean_nov_2016 = mean(combined_df$dengue_cases[nov_2016])
mean_dec_2016 = mean(combined_df$dengue_cases[dec_2016])

mean_jan_2017 = mean(combined_df$dengue_cases[jan_2017])
mean_feb_2017 = mean(combined_df$dengue_cases[feb_2017])
mean_mar_2017 = mean(combined_df$dengue_cases[mar_2017])
mean_apr_2017 = mean(combined_df$dengue_cases[apr_2017])
mean_may_2017 = mean(combined_df$dengue_cases[may_2017])
mean_jun_2017 = mean(combined_df$dengue_cases[jun_2017])
mean_jul_2017 = mean(combined_df$dengue_cases[jul_2017])
mean_aug_2017 = mean(combined_df$dengue_cases[aug_2017])
mean_sep_2017 = mean(combined_df$dengue_cases[sep_2017])
mean_oct_2017 = mean(combined_df$dengue_cases[oct_2017])
mean_nov_2017 = mean(combined_df$dengue_cases[nov_2017])
mean_dec_2017 = mean(combined_df$dengue_cases[dec_2017])

mean_jan_2018 = mean(combined_df$dengue_cases[jan_2018])
mean_feb_2018 = mean(combined_df$dengue_cases[feb_2018])
mean_mar_2018 = mean(combined_df$dengue_cases[mar_2018])
mean_apr_2018 = mean(combined_df$dengue_cases[apr_2018])
mean_may_2018 = mean(combined_df$dengue_cases[may_2018])
mean_jun_2018 = mean(combined_df$dengue_cases[jun_2018])
mean_jul_2018 = mean(combined_df$dengue_cases[jul_2018])
mean_aug_2018 = mean(combined_df$dengue_cases[aug_2018])
mean_sep_2018 = mean(combined_df$dengue_cases[sep_2018])
mean_oct_2018 = mean(combined_df$dengue_cases[oct_2018])
mean_nov_2018 = mean(combined_df$dengue_cases[nov_2018])
mean_dec_2018 = mean(combined_df$dengue_cases[dec_2018])

mean_jan_2019 = mean(combined_df$dengue_cases[jan_2019])
mean_feb_2019 = mean(combined_df$dengue_cases[feb_2019])
mean_mar_2019 = mean(combined_df$dengue_cases[mar_2019])
mean_apr_2019 = mean(combined_df$dengue_cases[apr_2019])
mean_may_2019 = mean(combined_df$dengue_cases[may_2019])
mean_jun_2019 = mean(combined_df$dengue_cases[jun_2019])
mean_jul_2019 = mean(combined_df$dengue_cases[jul_2019])
mean_aug_2019 = mean(combined_df$dengue_cases[aug_2019])
mean_sep_2019 = mean(combined_df$dengue_cases[sep_2019])
mean_oct_2019 = mean(combined_df$dengue_cases[oct_2019])
mean_nov_2019 = mean(combined_df$dengue_cases[nov_2019])
mean_dec_2019 = mean(combined_df$dengue_cases[dec_2019])

mean_jan_2020 = mean(combined_df$dengue_cases[jan_2020])
mean_feb_2020 = mean(combined_df$dengue_cases[feb_2020])
mean_mar_2020 = mean(combined_df$dengue_cases[mar_2020])
mean_apr_2020 = mean(combined_df$dengue_cases[apr_2020])
mean_may_2020 = mean(combined_df$dengue_cases[may_2020])
mean_jun_2020 = mean(combined_df$dengue_cases[jun_2020])
mean_jul_2020 = mean(combined_df$dengue_cases[jul_2020])
mean_aug_2020 = mean(combined_df$dengue_cases[aug_2020])
mean_sep_2020 = mean(combined_df$dengue_cases[sep_2020])
mean_oct_2020 = mean(combined_df$dengue_cases[oct_2020])
mean_nov_2020 = mean(combined_df$dengue_cases[nov_2020])
mean_dec_2020 = mean(combined_df$dengue_cases[dec_2020])
```

```{r}
#assembling dataframe
combined_df_mean = rbind(
  "1.2006" = mean_jan_2006,
  "2.2006" = mean_feb_2006, 
  "3.2006" = mean_mar_2006, 
  "4.2006" = mean_apr_2006,
  "5.2006" = mean_may_2006, 
  "6.2006" = mean_jun_2006,
  "7.2006" = mean_jul_2006,
  "8.2006" = mean_aug_2006, 
  "9.2006" = mean_sep_2006, 
  "10.2006" = mean_oct_2006, 
  "11.2006" = mean_nov_2006, 
  "12.2006" = mean_dec_2006, 

"1.2007" = mean_jan_2007, 
"2.2007" = mean_feb_2007, 
"3.2007" = mean_mar_2007, 
"4.2007" = mean_apr_2007, 
"5.2007" = mean_may_2007, 
"6.2007" = mean_jun_2007, 
"7.2007" = mean_jul_2007, 
"8.2007" = mean_aug_2007, 
"9.2007" = mean_sep_2007, 
"10.2007" = mean_oct_2007, 
"11.2007" = mean_nov_2007, 
"12.2007" = mean_dec_2007, 

"1.2008" = mean_jan_2008, 
"2.2008" = mean_feb_2008, 
"3.2008" = mean_mar_2008, 
"4.2008" = mean_apr_2008, 
"5.2008" = mean_may_2008, 
"6.2008" = mean_jun_2008, 
"7.2008" = mean_jul_2008, 
"8.2008" = mean_aug_2008, 
"9.2008" = mean_sep_2008, 
"10.2008" = mean_oct_2008, 
"11.2008" = mean_nov_2008, 
"12.2008" = mean_dec_2008, 

"1.2009" = mean_jan_2009, 
"2.2009" = mean_feb_2009,
"3.2009" = mean_mar_2009, 
"4.2009" = mean_apr_2009, 
"5.2009" = mean_may_2009, 
"6.2009" = mean_jun_2009, 
"7.2009" = mean_jul_2009,
"8.2009" = mean_aug_2009, 
"9.2009" = mean_sep_2009, 
"10.2009" = mean_oct_2009,
"11.2009" = mean_nov_2009, 
"12.2009" = mean_dec_2009, 

"1.2010" = mean_jan_2010, 
"2.2010" = mean_feb_2010,
"3.2010" = mean_mar_2010,
"4.2010" = mean_apr_2010,
"5.2010" = mean_may_2010,
"6.2010" = mean_jun_2010,
"7.2010" = mean_jul_2010,
"8.2010" = mean_aug_2010,
"9.2010" = mean_sep_2010,
"10.2010" = mean_oct_2010,
"11.2010" = mean_nov_2010,
"12.2010" = mean_dec_2010,

"1.2011" = mean_jan_2011, 
"2.2011" = mean_feb_2011, 
"3.2011" = mean_mar_2011,
"4.2011" = mean_apr_2011, 
"5.2011" = mean_may_2011,
"6.2011" = mean_jun_2011,
"7.2011" = mean_jul_2011,
"8.2011" = mean_aug_2011,
"9.2011" = mean_sep_2011,
"10.2011" = mean_oct_2011,
"11.2011" = mean_nov_2011,
"12.2011" = mean_dec_2011,

"1.2012" = mean_jan_2012, 
"2.2012" = mean_feb_2012, 
"3.2012" = mean_mar_2012, 
"4.2012" = mean_apr_2012, 
"5.2012" = mean_may_2012, 
"6.2012" = mean_jun_2012, 
"7.2012" = mean_jul_2012, 
"8.2012" = mean_aug_2012, 
"9.2012" = mean_sep_2012, 
"10.2012" = mean_oct_2012, 
"11.2012" = mean_nov_2012, 
"12.2012" = mean_dec_2012, 

"1.2013" = mean_jan_2013, 
"2.2013" = mean_feb_2013, 
"3.2013" = mean_mar_2013, 
"4.2013" = mean_apr_2013, 
"5.2013" = mean_may_2013, 
"6.2013" = mean_jun_2013, 
"7.2013" = mean_jul_2013, 
"8.2013" = mean_aug_2013, 
"9.2013" = mean_sep_2013, 
"10.2013" = mean_oct_2013, 
"11.2013" = mean_nov_2013, 
"12.2013" = mean_dec_2013, 

"1.2014" = mean_jan_2014, 
"2.2014" = mean_feb_2014, 
"3.2014" = mean_mar_2014, 
"4.2014" = mean_apr_2014, 
"5.2014" = mean_may_2014, 
"6.2014" = mean_jun_2014, 
"7.2014" = mean_jul_2014, 
"8.2014" = mean_aug_2014, 
"9.2014" = mean_sep_2014, 
"10.2014" = mean_oct_2014, 
"11.2014" = mean_nov_2014, 
"12.2014" = mean_dec_2014, 

"1.2015" = mean_jan_2015, 
"2.2015" = mean_feb_2015, 
"3.2015" = mean_mar_2015, 
"4.2015" = mean_apr_2015, 
"5.015" = mean_may_2015, 
"6.2015" = mean_jun_2015, 
"7.2015" = mean_jul_2015, 
"8.2015" = mean_aug_2015, 
"9.2015" = mean_sep_2015, 
"10.2015" = mean_oct_2015, 
"11.2015" = mean_nov_2015, 
"12.2015" = mean_dec_2015, 

"1.2016" = mean_jan_2016, 
"2.2016" = mean_feb_2016, 
"3.2016" = mean_mar_2016, 
"4.2016" = mean_apr_2016, 
"5.2016" = mean_may_2016, 
"6.2016" = mean_jun_2016, 
"7.2016" = mean_jul_2016,
"8.2016" = mean_aug_2016, 
"9.2016" = mean_sep_2016, 
"10.2016" = mean_oct_2016, 
"11.2016" = mean_nov_2016, 
"12.2016" = mean_dec_2016, 

"1.2017" = mean_jan_2017, 
"2.2017" = mean_feb_2017, 
"3.2017" = mean_mar_2017, 
"4.2017" = mean_apr_2017, 
"5.2017" = mean_may_2017, 
"6.2017" = mean_jun_2017, 
"7.2017" = mean_jul_2017, 
"8.2017" = mean_aug_2017, 
"9.2017" = mean_sep_2017, 
"10.2017" = mean_oct_2017, 
"11.2017" = mean_nov_2017, 
"12.2017" = mean_dec_2017, 

"1.2018" = mean_jan_2018, 
"2.2018" = mean_feb_2018, 
"3.2018" = mean_mar_2018, 
"4.2018" = mean_apr_2018, 
"5.2018" = mean_may_2018, 
"6.2018" = mean_jun_2018, 
"7.2018" = mean_jul_2018, 
"8.2018" = mean_aug_2018, 
"9.2018" = mean_sep_2018, 
"10.2018" = mean_oct_2018, 
"11.2018" = mean_nov_2018, 
"12.2018" = mean_dec_2018, 

"1.2019" = mean_jan_2019, 
"2.2019" = mean_feb_2019, 
"3.2019" = mean_mar_2019, 
"4.2019" = mean_apr_2019, 
"5.2019" = mean_may_2019, 
"6.2019" = mean_jun_2019, 
"7.2019" = mean_jul_2019, 
"8.2019" = mean_aug_2019, 
"9.2019" = mean_sep_2019, 
"10.2019" = mean_oct_2019, 
"11.2019" = mean_nov_2019, 
"12.2019" = mean_dec_2019, 

"2.2020" = mean_jan_2020,
"2.2020" = mean_feb_2020, 
"3.2020" = mean_mar_2020, 
"4.2020" = mean_apr_2020, 
"5.2020" = mean_may_2020, 
"6.2020" = mean_jun_2020, 
"7.2020" = mean_jul_2020, 
"8.2020" = mean_aug_2020, 
"9.2020" = mean_sep_2020, 
"10.2020" = mean_oct_2020, 
"11.2020" = mean_nov_2020, 
"12.2020" = mean_dec_2020)


colnames(combined_df_mean) = c("average_dengue_cases")
print(combined_df_mean)
```

```{r}
plot.ts(combined_df_mean$average_dengue_cases,
        main= "mean dengue cases of all provinces (2006 to 2020)", 
        ylab= "mean dengue cases",
        xlab= "time",
        ylim = c(0,2500)
        ) +
  abline(v= c(0,12,24,36,48,60,72,84,96,108,120,132,144,156,168), col="violet")

```

## 4.2) plotting sum of dengue cases of the whole country in a time series

```{r}
print(combined_df)

sum_jan_2006 = sum(combined_df$dengue_cases[jan_2006])
sum_feb_2006 = sum(combined_df$dengue_cases[feb_2006])
sum_mar_2006 = sum(combined_df$dengue_cases[mar_2006])
sum_apr_2006 = sum(combined_df$dengue_cases[apr_2006])
sum_may_2006 = sum(combined_df$dengue_cases[may_2006])
sum_jun_2006 = sum(combined_df$dengue_cases[jun_2006])
sum_jul_2006 = sum(combined_df$dengue_cases[jul_2006])
sum_aug_2006 = sum(combined_df$dengue_cases[aug_2006])
sum_sep_2006 = sum(combined_df$dengue_cases[sep_2006])
sum_oct_2006 = sum(combined_df$dengue_cases[oct_2006])
sum_nov_2006 = sum(combined_df$dengue_cases[nov_2006])
sum_dec_2006 = sum(combined_df$dengue_cases[dec_2006])

sum_jan_2007 = sum(combined_df$dengue_cases[jan_2007])
sum_feb_2007 = sum(combined_df$dengue_cases[feb_2007])
sum_mar_2007 = sum(combined_df$dengue_cases[mar_2007])
sum_apr_2007 = sum(combined_df$dengue_cases[apr_2007])
sum_may_2007 = sum(combined_df$dengue_cases[may_2007])
sum_jun_2007 = sum(combined_df$dengue_cases[jun_2007])
sum_jul_2007 = sum(combined_df$dengue_cases[jul_2007])
sum_aug_2007 = sum(combined_df$dengue_cases[aug_2007])
sum_sep_2007 = sum(combined_df$dengue_cases[sep_2007])
sum_oct_2007 = sum(combined_df$dengue_cases[oct_2007])
sum_nov_2007 = sum(combined_df$dengue_cases[nov_2007])
sum_dec_2007 = sum(combined_df$dengue_cases[dec_2007])

sum_jan_2008 = sum(combined_df$dengue_cases[jan_2008])
sum_feb_2008 = sum(combined_df$dengue_cases[feb_2008])
sum_mar_2008 = sum(combined_df$dengue_cases[mar_2008])
sum_apr_2008 = sum(combined_df$dengue_cases[apr_2008])
sum_may_2008 = sum(combined_df$dengue_cases[may_2008])
sum_jun_2008 = sum(combined_df$dengue_cases[jun_2008])
sum_jul_2008 = sum(combined_df$dengue_cases[jul_2008])
sum_aug_2008 = sum(combined_df$dengue_cases[aug_2008])
sum_sep_2008 = sum(combined_df$dengue_cases[sep_2008])
sum_oct_2008 = sum(combined_df$dengue_cases[oct_2008])
sum_nov_2008 = sum(combined_df$dengue_cases[nov_2008])
sum_dec_2008 = sum(combined_df$dengue_cases[dec_2008])

sum_jan_2009 = sum(combined_df$dengue_cases[jan_2009])
sum_feb_2009 = sum(combined_df$dengue_cases[feb_2009])
sum_mar_2009 = sum(combined_df$dengue_cases[mar_2009])
sum_apr_2009 = sum(combined_df$dengue_cases[apr_2009])
sum_may_2009 = sum(combined_df$dengue_cases[may_2009])
sum_jun_2009 = sum(combined_df$dengue_cases[jun_2009])
sum_jul_2009 = sum(combined_df$dengue_cases[jul_2009])
sum_aug_2009 = sum(combined_df$dengue_cases[aug_2009])
sum_sep_2009 = sum(combined_df$dengue_cases[sep_2009])
sum_oct_2009 = sum(combined_df$dengue_cases[oct_2009])
sum_nov_2009 = sum(combined_df$dengue_cases[nov_2009])
sum_dec_2009 = sum(combined_df$dengue_cases[dec_2009])

sum_jan_2010 = sum(combined_df$dengue_cases[jan_2010])
sum_feb_2010 = sum(combined_df$dengue_cases[feb_2010])
sum_mar_2010 = sum(combined_df$dengue_cases[mar_2010])
sum_apr_2010 = sum(combined_df$dengue_cases[apr_2010])
sum_may_2010 = sum(combined_df$dengue_cases[may_2010])
sum_jun_2010 = sum(combined_df$dengue_cases[jun_2010])
sum_jul_2010 = sum(combined_df$dengue_cases[jul_2010])
sum_aug_2010 = sum(combined_df$dengue_cases[aug_2010])
sum_sep_2010 = sum(combined_df$dengue_cases[sep_2010])
sum_oct_2010 = sum(combined_df$dengue_cases[oct_2010])
sum_nov_2010 = sum(combined_df$dengue_cases[nov_2010])
sum_dec_2010 = sum(combined_df$dengue_cases[dec_2010])

sum_jan_2011 = sum(combined_df$dengue_cases[jan_2011])
sum_feb_2011 = sum(combined_df$dengue_cases[feb_2011])
sum_mar_2011 = sum(combined_df$dengue_cases[mar_2011])
sum_apr_2011 = sum(combined_df$dengue_cases[apr_2011])
sum_may_2011 = sum(combined_df$dengue_cases[may_2011])
sum_jun_2011 = sum(combined_df$dengue_cases[jun_2011])
sum_jul_2011 = sum(combined_df$dengue_cases[jul_2011])
sum_aug_2011 = sum(combined_df$dengue_cases[aug_2011])
sum_sep_2011 = sum(combined_df$dengue_cases[sep_2011])
sum_oct_2011 = sum(combined_df$dengue_cases[oct_2011])
sum_nov_2011 = sum(combined_df$dengue_cases[nov_2011])
sum_dec_2011 = sum(combined_df$dengue_cases[dec_2011])

sum_jan_2012 = sum(combined_df$dengue_cases[jan_2012])
sum_feb_2012 = sum(combined_df$dengue_cases[feb_2012])
sum_mar_2012 = sum(combined_df$dengue_cases[mar_2012])
sum_apr_2012 = sum(combined_df$dengue_cases[apr_2012])
sum_may_2012 = sum(combined_df$dengue_cases[may_2012])
sum_jun_2012 = sum(combined_df$dengue_cases[jun_2012])
sum_jul_2012 = sum(combined_df$dengue_cases[jul_2012])
sum_aug_2012 = sum(combined_df$dengue_cases[aug_2012])
sum_sep_2012 = sum(combined_df$dengue_cases[sep_2012])
sum_oct_2012 = sum(combined_df$dengue_cases[oct_2012])
sum_nov_2012 = sum(combined_df$dengue_cases[nov_2012])
sum_dec_2012 = sum(combined_df$dengue_cases[dec_2012])

sum_jan_2013 = sum(combined_df$dengue_cases[jan_2013])
sum_feb_2013 = sum(combined_df$dengue_cases[feb_2013])
sum_mar_2013 = sum(combined_df$dengue_cases[mar_2013])
sum_apr_2013 = sum(combined_df$dengue_cases[apr_2013])
sum_may_2013 = sum(combined_df$dengue_cases[may_2013])
sum_jun_2013 = sum(combined_df$dengue_cases[jun_2013])
sum_jul_2013 = sum(combined_df$dengue_cases[jul_2013])
sum_aug_2013 = sum(combined_df$dengue_cases[aug_2013])
sum_sep_2013 = sum(combined_df$dengue_cases[sep_2013])
sum_oct_2013 = sum(combined_df$dengue_cases[oct_2013])
sum_nov_2013 = sum(combined_df$dengue_cases[nov_2013])
sum_dec_2013 = sum(combined_df$dengue_cases[dec_2013])

sum_jan_2014 = sum(combined_df$dengue_cases[jan_2014])
sum_feb_2014 = sum(combined_df$dengue_cases[feb_2014])
sum_mar_2014 = sum(combined_df$dengue_cases[mar_2014])
sum_apr_2014 = sum(combined_df$dengue_cases[apr_2014])
sum_may_2014 = sum(combined_df$dengue_cases[may_2014])
sum_jun_2014 = sum(combined_df$dengue_cases[jun_2014])
sum_jul_2014 = sum(combined_df$dengue_cases[jul_2014])
sum_aug_2014 = sum(combined_df$dengue_cases[aug_2014])
sum_sep_2014 = sum(combined_df$dengue_cases[sep_2014])
sum_oct_2014 = sum(combined_df$dengue_cases[oct_2014])
sum_nov_2014 = sum(combined_df$dengue_cases[nov_2014])
sum_dec_2014 = sum(combined_df$dengue_cases[dec_2014])

sum_jan_2015 = sum(combined_df$dengue_cases[jan_2015])
sum_feb_2015 = sum(combined_df$dengue_cases[feb_2015])
sum_mar_2015 = sum(combined_df$dengue_cases[mar_2015])
sum_apr_2015 = sum(combined_df$dengue_cases[apr_2015])
sum_may_2015 = sum(combined_df$dengue_cases[may_2015])
sum_jun_2015 = sum(combined_df$dengue_cases[jun_2015])
sum_jul_2015 = sum(combined_df$dengue_cases[jul_2015])
sum_aug_2015 = sum(combined_df$dengue_cases[aug_2015])
sum_sep_2015 = sum(combined_df$dengue_cases[sep_2015])
sum_oct_2015 = sum(combined_df$dengue_cases[oct_2015])
sum_nov_2015 = sum(combined_df$dengue_cases[nov_2015])
sum_dec_2015 = sum(combined_df$dengue_cases[dec_2015])

sum_jan_2016 = sum(combined_df$dengue_cases[jan_2016])
sum_feb_2016 = sum(combined_df$dengue_cases[feb_2016])
sum_mar_2016 = sum(combined_df$dengue_cases[mar_2016])
sum_apr_2016 = sum(combined_df$dengue_cases[apr_2016])
sum_may_2016 = sum(combined_df$dengue_cases[may_2016])
sum_jun_2016 = sum(combined_df$dengue_cases[jun_2016])
sum_jul_2016 = sum(combined_df$dengue_cases[jul_2016])
sum_aug_2016 = sum(combined_df$dengue_cases[aug_2016])
sum_sep_2016 = sum(combined_df$dengue_cases[sep_2016])
sum_oct_2016 = sum(combined_df$dengue_cases[oct_2016])
sum_nov_2016 = sum(combined_df$dengue_cases[nov_2016])
sum_dec_2016 = sum(combined_df$dengue_cases[dec_2016])

sum_jan_2017 = sum(combined_df$dengue_cases[jan_2017])
sum_feb_2017 = sum(combined_df$dengue_cases[feb_2017])
sum_mar_2017 = sum(combined_df$dengue_cases[mar_2017])
sum_apr_2017 = sum(combined_df$dengue_cases[apr_2017])
sum_may_2017 = sum(combined_df$dengue_cases[may_2017])
sum_jun_2017 = sum(combined_df$dengue_cases[jun_2017])
sum_jul_2017 = sum(combined_df$dengue_cases[jul_2017])
sum_aug_2017 = sum(combined_df$dengue_cases[aug_2017])
sum_sep_2017 = sum(combined_df$dengue_cases[sep_2017])
sum_oct_2017 = sum(combined_df$dengue_cases[oct_2017])
sum_nov_2017 = sum(combined_df$dengue_cases[nov_2017])
sum_dec_2017 = sum(combined_df$dengue_cases[dec_2017])

sum_jan_2018 = sum(combined_df$dengue_cases[jan_2018])
sum_feb_2018 = sum(combined_df$dengue_cases[feb_2018])
sum_mar_2018 = sum(combined_df$dengue_cases[mar_2018])
sum_apr_2018 = sum(combined_df$dengue_cases[apr_2018])
sum_may_2018 = sum(combined_df$dengue_cases[may_2018])
sum_jun_2018 = sum(combined_df$dengue_cases[jun_2018])
sum_jul_2018 = sum(combined_df$dengue_cases[jul_2018])
sum_aug_2018 = sum(combined_df$dengue_cases[aug_2018])
sum_sep_2018 = sum(combined_df$dengue_cases[sep_2018])
sum_oct_2018 = sum(combined_df$dengue_cases[oct_2018])
sum_nov_2018 = sum(combined_df$dengue_cases[nov_2018])
sum_dec_2018 = sum(combined_df$dengue_cases[dec_2018])

sum_jan_2019 = sum(combined_df$dengue_cases[jan_2019])
sum_feb_2019 = sum(combined_df$dengue_cases[feb_2019])
sum_mar_2019 = sum(combined_df$dengue_cases[mar_2019])
sum_apr_2019 = sum(combined_df$dengue_cases[apr_2019])
sum_may_2019 = sum(combined_df$dengue_cases[may_2019])
sum_jun_2019 = sum(combined_df$dengue_cases[jun_2019])
sum_jul_2019 = sum(combined_df$dengue_cases[jul_2019])
sum_aug_2019 = sum(combined_df$dengue_cases[aug_2019])
sum_sep_2019 = sum(combined_df$dengue_cases[sep_2019])
sum_oct_2019 = sum(combined_df$dengue_cases[oct_2019])
sum_nov_2019 = sum(combined_df$dengue_cases[nov_2019])
sum_dec_2019 = sum(combined_df$dengue_cases[dec_2019])

sum_jan_2020 = sum(combined_df$dengue_cases[jan_2020])
sum_feb_2020 = sum(combined_df$dengue_cases[feb_2020])
sum_mar_2020 = sum(combined_df$dengue_cases[mar_2020])
sum_apr_2020 = sum(combined_df$dengue_cases[apr_2020])
sum_may_2020 = sum(combined_df$dengue_cases[may_2020])
sum_jun_2020 = sum(combined_df$dengue_cases[jun_2020])
sum_jul_2020 = sum(combined_df$dengue_cases[jul_2020])
sum_aug_2020 = sum(combined_df$dengue_cases[aug_2020])
sum_sep_2020 = sum(combined_df$dengue_cases[sep_2020])
sum_oct_2020 = sum(combined_df$dengue_cases[oct_2020])
sum_nov_2020 = sum(combined_df$dengue_cases[nov_2020])
sum_dec_2020 = sum(combined_df$dengue_cases[dec_2020])
```

```{r}
#assembling dataframe
combined_df_sum = rbind(
  "1.2006" = sum_jan_2006,
  "2.2006" = sum_feb_2006, 
  "3.2006" = sum_mar_2006, 
  "4.2006" = sum_apr_2006,
  "5.2006" = sum_may_2006, 
  "6.2006" = sum_jun_2006,
  "7.2006" = sum_jul_2006,
  "8.2006" = sum_aug_2006, 
  "9.2006" = sum_sep_2006, 
  "10.2006" = sum_oct_2006, 
  "11.2006" = sum_nov_2006, 
  "12.2006" = sum_dec_2006, 

"1.2007" = sum_jan_2007, 
"2.2007" = sum_feb_2007, 
"3.2007" = sum_mar_2007, 
"4.2007" = sum_apr_2007, 
"5.2007" = sum_may_2007, 
"6.2007" = sum_jun_2007, 
"7.2007" = sum_jul_2007, 
"8.2007" = sum_aug_2007, 
"9.2007" = sum_sep_2007, 
"10.2007" = sum_oct_2007, 
"11.2007" = sum_nov_2007, 
"12.2007" = sum_dec_2007, 

"1.2008" = sum_jan_2008, 
"2.2008" = sum_feb_2008, 
"3.2008" = sum_mar_2008, 
"4.2008" = sum_apr_2008, 
"5.2008" = sum_may_2008, 
"6.2008" = sum_jun_2008, 
"7.2008" = sum_jul_2008, 
"8.2008" = sum_aug_2008, 
"9.2008" = sum_sep_2008, 
"10.2008" = sum_oct_2008, 
"11.2008" = sum_nov_2008, 
"12.2008" = sum_dec_2008, 

"1.2009" = sum_jan_2009, 
"2.2009" = sum_feb_2009,
"3.2009" = sum_mar_2009, 
"4.2009" = sum_apr_2009, 
"5.2009" = sum_may_2009, 
"6.2009" = sum_jun_2009, 
"7.2009" = sum_jul_2009,
"8.2009" = sum_aug_2009, 
"9.2009" = sum_sep_2009, 
"10.2009" = sum_oct_2009,
"11.2009" = sum_nov_2009, 
"12.2009" = sum_dec_2009, 

"1.2010" = sum_jan_2010, 
"2.2010" = sum_feb_2010,
"3.2010" = sum_mar_2010,
"4.2010" = sum_apr_2010,
"5.2010" = sum_may_2010,
"6.2010" = sum_jun_2010,
"7.2010" = sum_jul_2010,
"8.2010" = sum_aug_2010,
"9.2010" = sum_sep_2010,
"10.2010" = sum_oct_2010,
"11.2010" = sum_nov_2010,
"12.2010" = sum_dec_2010,

"1.2011" = sum_jan_2011, 
"2.2011" = sum_feb_2011, 
"3.2011" = sum_mar_2011,
"4.2011" = sum_apr_2011, 
"5.2011" = sum_may_2011,
"6.2011" = sum_jun_2011,
"7.2011" = sum_jul_2011,
"8.2011" = sum_aug_2011,
"9.2011" = sum_sep_2011,
"10.2011" = sum_oct_2011,
"11.2011" = sum_nov_2011,
"12.2011" = sum_dec_2011,

"1.2012" = sum_jan_2012, 
"2.2012" = sum_feb_2012, 
"3.2012" = sum_mar_2012, 
"4.2012" = sum_apr_2012, 
"5.2012" = sum_may_2012, 
"6.2012" = sum_jun_2012, 
"7.2012" = sum_jul_2012, 
"8.2012" = sum_aug_2012, 
"9.2012" = sum_sep_2012, 
"10.2012" = sum_oct_2012, 
"11.2012" = sum_nov_2012, 
"12.2012" = sum_dec_2012, 

"1.2013" = sum_jan_2013, 
"2.2013" = sum_feb_2013, 
"3.2013" = sum_mar_2013, 
"4.2013" = sum_apr_2013, 
"5.2013" = sum_may_2013, 
"6.2013" = sum_jun_2013, 
"7.2013" = sum_jul_2013, 
"8.2013" = sum_aug_2013, 
"9.2013" = sum_sep_2013, 
"10.2013" = sum_oct_2013, 
"11.2013" = sum_nov_2013, 
"12.2013" = sum_dec_2013, 

"1.2014" = sum_jan_2014, 
"2.2014" = sum_feb_2014, 
"3.2014" = sum_mar_2014, 
"4.2014" = sum_apr_2014, 
"5.2014" = sum_may_2014, 
"6.2014" = sum_jun_2014, 
"7.2014" = sum_jul_2014, 
"8.2014" = sum_aug_2014, 
"9.2014" = sum_sep_2014, 
"10.2014" = sum_oct_2014, 
"11.2014" = sum_nov_2014, 
"12.2014" = sum_dec_2014, 

"1.2015" = sum_jan_2015, 
"2.2015" = sum_feb_2015, 
"3.2015" = sum_mar_2015, 
"4.2015" = sum_apr_2015, 
"5.2015" = sum_may_2015, 
"6.2015" = sum_jun_2015, 
"7.2015" = sum_jul_2015, 
"8.2015" = sum_aug_2015, 
"9.2015" = sum_sep_2015, 
"10.2015" = sum_oct_2015, 
"11.2015" = sum_nov_2015, 
"12.2015" = sum_dec_2015, 

"1.2016" = sum_jan_2016, 
"2.2016" = sum_feb_2016, 
"3.2016" = sum_mar_2016, 
"4.2016" = sum_apr_2016, 
"5.2016" = sum_may_2016, 
"6.2016" = sum_jun_2016, 
"7.2016" = sum_jul_2016,
"8.2016" = sum_aug_2016, 
"9.2016" = sum_sep_2016, 
"10.2016" = sum_oct_2016, 
"11.2016" = sum_nov_2016, 
"12.2016" = sum_dec_2016, 

"1.2017" = sum_jan_2017, 
"2.2017" = sum_feb_2017, 
"3.2017" = sum_mar_2017, 
"4.2017" = sum_apr_2017, 
"5.2017" = sum_may_2017, 
"6.2017" = sum_jun_2017, 
"7.2017" = sum_jul_2017, 
"8.2017" = sum_aug_2017, 
"9.2017" = sum_sep_2017, 
"10.2017" = sum_oct_2017, 
"11.2017" = sum_nov_2017, 
"12.2017" = sum_dec_2017, 

"1.2018" = sum_jan_2018, 
"2.2018" = sum_feb_2018, 
"3.2018" = sum_mar_2018, 
"4.2018" = sum_apr_2018, 
"5.2018" = sum_may_2018, 
"6.2018" = sum_jun_2018, 
"7.2018" = sum_jul_2018, 
"8.2018" = sum_aug_2018, 
"9.2018" = sum_sep_2018, 
"10.2018" = sum_oct_2018, 
"11.2018" = sum_nov_2018, 
"12.2018" = sum_dec_2018, 

"1.2019" = sum_jan_2019, 
"2.2019" = sum_feb_2019, 
"3.2019" = sum_mar_2019, 
"4.2019" = sum_apr_2019, 
"5.2019" = sum_may_2019, 
"6.2019" = sum_jun_2019, 
"7.2019" = sum_jul_2019, 
"8.2019" = sum_aug_2019, 
"9.2019" = sum_sep_2019, 
"10.2019" = sum_oct_2019, 
"11.2019" = sum_nov_2019, 
"12.2019" = sum_dec_2019, 

"2.2020" = sum_jan_2020,
"2.2020" = sum_feb_2020, 
"3.2020" = sum_mar_2020, 
"4.2020" = sum_apr_2020, 
"5.2020" = sum_may_2020, 
"6.2020" = sum_jun_2020, 
"7.2020" = sum_jul_2020, 
"8.2020" = sum_aug_2020, 
"9.2020" = sum_sep_2020, 
"10.2020" = sum_oct_2020, 
"11.2020" = sum_nov_2020, 
"12.2020" = sum_dec_2020)


colnames(combined_df_sum) = c("sum_dengue_cases")
print(combined_df_sum)
```
