---
title: "general data analysis: characteristics of the data"
output: html_document
date: "2023-05-16"
---

# General analysis: characteristics of the data

## loading the data

```{r}
#loading mean monthly relative humidity, temperature, precipitation data

install.packages("readxl")
library(readxl)

sheet_names = excel_sheets("/Users/leamrowiec/Desktop/Datenanalyse SS23/era5_data_2006_2020_thailand_monmean.xlsx")

#laden von den einzelnen sheets
T.huss.mon = read_excel("/Users/leamrowiec/Desktop/Datenanalyse SS23/era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
T.tp.mon = read_excel("/Users/leamrowiec/Desktop/Datenanalyse SS23/era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_tp_2006_2020_thailand_monm")
T.t2m.mon = read_excel("/Users/leamrowiec/Desktop/Datenanalyse SS23/era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_t2m_2006_2020_thailand_mon")




```

```{r}
show(T.huss.mon)
show(T.tp.mon)
show(T.t2m.mon)
```

## 2) data cleanup

### 2.1) naming columns and rows

```{r}
# naming each row with the name of the province
rownames(T.huss.mon) <- T.huss.mon[,2]
rownames(T.tp.mon) <- T.tp.mon[,2]
rownames(T.t2m.mon) <- T.t2m.mon[,2]

#naming each colum with years of year, numbers of month
show(colnames(T.huss.mon)) # -> colums are already named 
```

```{r}
#naming each colum with years of year, numbers of month
show(colnames(T.huss.mon))
```

### 2.2) removing unnecessary columns

```{r}
#removing unnecessary colums: numbers of provinces (col 1), lonitude and latitude data (col 3,4)
T.huss.mon <- T.huss.mon[,c(-1,-3,-4)]
T.tp.mon <- T.tp.mon[,c(-1,-3,-4)]
T.t2m.mon <-T.t2m.mon[,c(-1,-3,-4)]

show(T.huss.mon)
show(T.tp.mon)
show(T.t2m.mon)
```

```{r}
# summary of each dataset
summary(T.huss.mon)
summary(T.tp.mon)
summary(T.t2m.mon)
```

## 3) computing the mean and median monthly relative humidity, temperature and precipitation (mean of province data)

```{r}
# instaling ggplot2
install.packages("ggplot2")
library(ggplot2)

```

```{r}
# computing monthly mean per province
mean.huss.monthly = sapply(T.huss.mon[,-(1)],function(x) {mean(x)})
mean.tp.monthly = sapply(T.tp.mon[,-(1)],function(x) {mean(x)})
mean.t2m.monthly = sapply(T.t2m.mon[,-(1)], function(x) {mean(x)})

#plotting monthly mean 
plot.ts(mean.huss.monthly, main= "relative humidity (mean of all provinces) per month", ylab="mean of relative humidity (monthly)", xlab="month")
plot.ts(mean.tp.monthly, main= "temperature (mean of all provinces) per month", ylab="mean precipitation (monthly)", xlab="month")
plot.ts(mean.t2m.monthly, main= "temperature (t2m) (mean of all provinces) per month", ylab="mean temperature (monthly)", xlab="month")




# plotting monthly mean of relative humidity, precipitation and temperature in one plot
x= colnames(T.huss.mon[, -(1)])

#runden
new.mean.huss.monthly = round( mean.huss.monthly, 0)
new.mean.tp.monthly = round(mean.tp.monthly, 0)
new.mean.t2m.monthly =round(mean.t2m.monthly, 0)
#converting into numbers
new.mean.huss.monthly.num = as.numeric(new.mean.huss.monthly)
new.mean.tp.monthly.num = as.numeric(new.mean.tp.monthly)
new.mean.t2m.monthly.num = as.numeric(new.mean.t2m.monthly)
#plotten
plot(x, new.mean.huss.monthly.num, col = "red", pch="o", lty=2, xlim = c(1, 200), ylim = c(1, 300), xlab = "months", ylab = "mean huss/mean tp/mean t2m")

lines(x, new.mean.tp.monthly.num, col = "blue",lty=2)
points(x, new.mean.tp.monthly.num, col = "blue", pch="o")

lines(x, new.mean.t2m.monthly.num, col = "green",lty=2)
points(x, new.mean.t2m.monthly.num, col = "green", pch="o")

legend("topright", legend = c("mean monthly huss", "mean monthly precipitation","mean monthly temperature"), col = c("red", "blue", "green"), lwd = c(2, 2, 2, 2), pch = c(16, 16, 16))

#wieso funktioniert es nicht? gleiche länge,numercs, keine characters, keine kommas
```

```{r}
# computing monthly mean per province
median.huss.monthly = sapply(T.huss.mon[,-(1)],function(x) {median(x)})
median.tp.monthly = sapply(T.tp.mon[,-(1)],function(x) {median(x)})
median.t2m.monthly = sapply(T.t2m.mon[,-(1)], function(x) {median(x)})

#plotting monthly mean 
plot.ts(median.huss.monthly, main= "relative humidity (median of all provinces) per month", ylab="median of relative humidity (monthly)", xlab="month")
plot.ts(median.tp.monthly, main= "temperature (median of all provinces) per month", ylab="median precipitation (monthly)", xlab="month")
plot.ts(median.t2m.monthly, main= "temperature (t2m) (median of all provinces) per month", ylab="median temperature (monthly)", xlab="month")
```

## 4) plotting mean of values (of each month) for each province

```{r}
# computing monthly mean of climate values for province over complete time span of available data

#step1: transposing the dataframe to use sapply rowwise
t.T.huss.mon = t(T.huss.mon)
t.T.tp.mon = t(T.tp.mon)
t.T.t2m.mon = t(T.t2m.mon)
#step2: computing the mean of each value in each province
mean.huss.prov = apply(T.huss.mon[,-(1)],2,function(x) {mean(x)})
mean.tp.prov = apply(T.tp.mon[,-(1)],2,function(x) {mean(x)})
mean.t2m.prov = apply(T.t2m.mon[,-(1)],2,function(x) {mean(x)})

```

## 5) computing the maximum/minimum value for each month of each year

```{r}
# computing monthly maximum over all provinces
T.huss.mon.max = apply(T.huss.mon[,-(1)], 1, function(x) {max(x)})
T.tp.mon.max = apply(T.tp.mon[,-(1)], 1, function(x) {max(x)})
T.t2m.mon.max = apply(T.t2m.mon[,-(1)], 1, function(x) {max(x)})

# computing monthly minimum over all provinces
T.huss.mon.min = apply(T.huss.mon[,-(1)],1, function(x) {min(x)})
T.tp.mon.min = apply(T.tp.mon[,-(1)],1, function(x) {min(x)})
T.t2m.mon.min = apply(T.t2m.mon[,-(1)],1, function(x) {min(x)})

# max and min return the mximum/minimum value of the dataframe, not the index

# plotting maximum monthly value of all provinces 
plot.ts(T.huss.mon.max, main= "maximum monthly relative humidity of all provinces", ylab= "maximum relative humidity (monthly)", xlab="month")
plot.ts(T.tp.mon.max, main= "maximum monthly precipitation of all provinces", ylab= "maximum precipitation (monthly)", xlab="month")
plot.ts(T.t2m.mon.max, main= "maximum monthly temperature of all provinces", ylab= "maximum temperature (monthly)", xlab="month")

# plotting minimim monthly value of all provinces
plot.ts(T.huss.mon.min, main= "minimum monthly relative humidity of all provinces", ylab= "minimum temperature (monthly)", xlab="month")
plot.ts(T.tp.mon.min, main= "minimum monthly precipitation of all provinces", ylab= "minimum temperature (monthly)", xlab="month")
plot.ts(T.t2m.mon.min, main= "minimum monthly temperature of all provinces", ylab= "minimum temperature (monthly)", xlab="month")


```

## 6) plotting abnormalities

```{r}
# 1) computing the mean of all entries in the dataframe
mean.huss = mean(unlist(T.huss.mon))
mean.tp = mean(unlist(T.tp.mon))
mean.t2m = mean(unlist(T.t2m.mon))

# 2) computing abnormality
abn.T.huss.mon = apply(T.huss.mon, 1, function(x)
  {x-mean.huss})
abn.T.tp.mon = apply(T.tp.mon, 1, function(x)
  {x-mean.tp})
abn.T.t2m.mon = apply(T.t2m.mon, 1, function(x)
  {x-mean.t2m})

# 3) plotting abnormalities of each month in a histogramm
hist(abn.T.huss.mon)
hist(abn.T.tp.mon)
hist(abn.T.t2m.mon)

```
