---
title: "Plot mit Yearly Dengue Data and Clima Data"
author: "Manon Manddernach"
date: "2023-07-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Preparing the Data 

## 0.1 Preparing the Data for yearly means

```{r}
library(ggplot2)
library(scales)
```


```{r}

## Years and Months  

Years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)
Months <- c(1,2,3,4,5,6,7,8,9,10,11,12)


## Dengue Data Mean values

dengue.norm.2006.mean.absolute <- as.data.frame(apply(dengue.norm.2006.mean,1,mean))
dengue.norm.2007.mean.absolute <- as.data.frame(apply(dengue.norm.2007.mean,1,mean))
dengue.norm.2008.mean.absolute <- as.data.frame(apply(dengue.norm.2008.mean,1,mean))
dengue.norm.2009.mean.absolute <- as.data.frame(apply(dengue.norm.2009.mean,1,mean))
dengue.norm.2010.mean.absolute <- as.data.frame(apply(dengue.norm.2010.mean,1,mean))
dengue.norm.2011.mean.absolute <- as.data.frame(apply(dengue.norm.2011.mean,1,mean))
dengue.norm.2012.mean.absolute <- as.data.frame(apply(dengue.norm.2012.mean,1,mean))
dengue.norm.2013.mean.absolute <- as.data.frame(apply(dengue.norm.2013.mean,1,mean))
dengue.norm.2014.mean.absolute <- as.data.frame(apply(dengue.norm.2014.mean,1,mean))
dengue.norm.2015.mean.absolute <- as.data.frame(apply(dengue.norm.2015.mean,1,mean))
dengue.norm.2016.mean.absolute <- as.data.frame(apply(dengue.norm.2016.mean,1,mean))
dengue.norm.2017.mean.absolute <- as.data.frame(apply(dengue.norm.2017.mean,1,mean))
dengue.norm.2018.mean.absolute <- as.data.frame(apply(dengue.norm.2018.mean,1,mean))
dengue.norm.2019.mean.absolute <- as.data.frame(apply(dengue.norm.2019.mean,1,mean))
dengue.norm.2020.mean.absolute <- as.data.frame(apply(dengue.norm.2020.mean,1,mean))

Dengue.mean <- c(dengue.norm.2006.mean.absolute,
                 dengue.norm.2007.mean.absolute,
                 dengue.norm.2008.mean.absolute,
                 dengue.norm.2009.mean.absolute, 
                 dengue.norm.2010.mean.absolute,
                 dengue.norm.2011.mean.absolute,
                 dengue.norm.2012.mean.absolute, 
                 dengue.norm.2013.mean.absolute, 
                 dengue.norm.2014.mean.absolute, 
                 dengue.norm.2015.mean.absolute, 
                 dengue.norm.2016.mean.absolute,
                 dengue.norm.2017.mean.absolute, 
                 dengue.norm.2018.mean.absolute, 
                 dengue.norm.2019.mean.absolute,
                 dengue.norm.2020.mean.absolute)

Dengue.mean <- lapply(Dengue.mean, unlist)
Dengue.mean <- lapply(Dengue.mean, mean)

Dengue.mean <- list(
  `apply(dengue.norm.2006.mean, 1, mean)` = 5.570221,
  `apply(dengue.norm.2007.mean, 1, mean)` = 8.742932,
  `apply(dengue.norm.2008.mean, 1, mean)` = 11.76473,
  `apply(dengue.norm.2009.mean, 1, mean)` = 7.219884,
  `apply(dengue.norm.2010.mean, 1, mean)` = 14.7395,
  `apply(dengue.norm.2011.mean, 1, mean)` = 8.785111,
  `apply(dengue.norm.2012.mean, 1, mean)` = 9.955709,
  `apply(dengue.norm.2013.mean, 1, mean)` = 20.17376,
  `apply(dengue.norm.2014.mean, 1, mean)` = 5.596716,
  `apply(dengue.norm.2015.mean, 1, mean)` = 17.39642,
  `apply(dengue.norm.2016.mean, 1, mean)` = 9.211036,
  `apply(dengue.norm.2017.mean, 1, mean)` = 6.315424,
  `apply(dengue.norm.2018.mean, 1, mean)` = 10.47807,
  `apply(dengue.norm.2019.mean, 1, mean)` = 14.60038,
  `apply(dengue.norm.2020.mean, 1, mean)` = 14.0717
)

Dengue.mean <- data.frame(Code = names(Dengue.mean), Value = unlist(Dengue.mean))
Dengue.mean <- Dengue.mean$Value


## Humidity Data Mean values 


huss.2006<- as.data.frame(T.huss.mon.ordered[,5:16])
huss.2007<- as.data.frame(T.huss.mon.ordered[,17:28])
huss.2008<- as.data.frame(T.huss.mon.ordered[,29:40])
huss.2009<- as.data.frame(T.huss.mon.ordered[,41:52])
huss.2010<- as.data.frame(T.huss.mon.ordered[,53:64])
huss.2011<- as.data.frame(T.huss.mon.ordered[,65:76])
huss.2012<- as.data.frame(T.huss.mon.ordered[,77:88])
huss.2013<- as.data.frame(T.huss.mon.ordered[,89:100])
huss.2014<- as.data.frame(T.huss.mon.ordered[,101:112])
huss.2015<- as.data.frame(T.huss.mon.ordered[,113:124])
huss.2016<- as.data.frame(T.huss.mon.ordered[,125:136])
huss.2017<- as.data.frame(T.huss.mon.ordered[,137:148])
huss.2018<- as.data.frame(T.huss.mon.ordered[,149:150])
huss.2019<- as.data.frame(T.huss.mon.ordered[,151:162])
huss.2020<- as.data.frame(T.huss.mon.ordered[,163:174])




huss.2006.mean.absolute <- as.data.frame(apply(huss.2006.mean,1,mean))
huss.2007.mean.absolute <- as.data.frame(apply(huss.2007.mean,1,mean))
huss.2008.mean.absolute <- as.data.frame(apply(huss.2008.mean,1,mean))
huss.2009.mean.absolute <- as.data.frame(apply(huss.2009.mean,1,mean))
huss.2010.mean.absolute <- as.data.frame(apply(huss.2010.mean,1,mean))
huss.2011.mean.absolute <- as.data.frame(apply(huss.2011.mean,1,mean))
huss.2012.mean.absolute <- as.data.frame(apply(huss.2012.mean,1,mean))
huss.2013.mean.absolute <- as.data.frame(apply(huss.2013.mean,1,mean))
huss.2014.mean.absolute <- as.data.frame(apply(huss.2014.mean,1,mean))
huss.2015.mean.absolute <- as.data.frame(apply(huss.2015.mean,1,mean))
huss.2016.mean.absolute <- as.data.frame(apply(huss.2016.mean,1,mean))
huss.2017.mean.absolute <- as.data.frame(apply(huss.2017.mean,1,mean))
huss.2018.mean.absolute <- as.data.frame(apply(huss.2018.mean,1,mean))
huss.2019.mean.absolute <- as.data.frame(apply(huss.2019.mean,1,mean))
huss.2020.mean.absolute <- as.data.frame(apply(huss.2020.mean,1,mean))

Huss.mean <- c(huss.2006.mean.absolute,
               huss.2007.mean.absolute,
               huss.2008.mean.absolute,
               huss.2009.mean.absolute, 
               huss.2010.mean.absolute,
               huss.2011.mean.absolute,
               huss.2012.mean.absolute, 
               huss.2013.mean.absolute, 
               huss.2014.mean.absolute, 
               huss.2015.mean.absolute, 
               huss.2016.mean.absolute,
               huss.2017.mean.absolute,
               huss.2018.mean.absolute, 
               huss.2019.mean.absolute,
               huss.2020.mean.absolute)

Huss.mean <- lapply(Huss.mean,unlist)
Huss.mean <- lapply(Huss.mean,mean)

Huss.mean <- list(
  `$apply(huss.2006.mean, 1, mean)` = 74.8919,
  `$apply(huss.2007.mean, 1, mean)` = 56.53615,
  `$apply(huss.2008.mean, 1, mean)` = 57.60083,
  `$apply(huss.2009.mean, 1, mean)` = 57.32692,
  `$apply(huss.2010.mean, 1, mean)` = 57.13848,
  `$apply(huss.2011.mean, 1, mean)` = 58.08443,
  `$apply(huss.2012.mean, 1, mean)` = 57.97239,
  `$apply(huss.2013.mean, 1, mean)` = 57.40518,
  `$apply(huss.2014.mean, 1, mean)` = 57.16852,
  `$apply(huss.2015.mean, 1, mean)` = 56.8422,
  `$apply(huss.2016.mean, 1, mean)` = 56.88025,
  `$apply(huss.2017.mean, 1, mean)` = 58.00465,
  `$apply(huss.2018.mean, 1, mean)` = 55.85306,
  `$apply(huss.2019.mean, 1, mean)` = 57.80723,
  `$apply(huss.2020.mean, 1, mean)` = 55.99571
)

Huss.mean <- data.frame(Code = names(Huss.mean), Value = unlist(Huss.mean))
Huss.mean <- Huss.mean$Value


Dengue.Humidity.Plot <- data.frame(Years=Years,
                                   Dengue.mean = Dengue.mean,
                                   Huss.mean = Huss.mean
                                   )



```

## Code for Provinces
```{r}

Bangkok <- data.frame(T.huss.mon.ordered[3, ])
BangkokD <- data.frame(dengue.norm[3, ])

Nakhon.Si.Thammarat <- data.frame(T.huss.mon.ordered[31, ])
Nakhon.Si.ThammaratD <- data.frame(dengue.norm[31, ])


```
 
 
```{r}
## Creating Dataframes for each Year
data_frames <- list()

for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Bangkok[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(BangkokD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

for (i in seq_along(data_frames)) {
  assign(paste0("Bangkok.", i + 2005), data_frames[[i]])
}



## Plotting GGPlots
plots <- list()

for (year in years) {
  Bangkok_year <- get(paste0("Bangkok.", year))
  Bangkok_year$Dengue <- as.numeric(Bangkok_year$Dengue)
  Bangkok_year$Humidity <- as.numeric(Bangkok_year$Humidity)
  
  # Caliculating correlation to print in GGPlot 
  correlation <- cor(Bangkok_year$Dengue, Bangkok_year$Humidity)

  plot <- ggplot(Bangkok_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Bangkok", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}

```
 
### Creating Data Frames for each Year for a province 
 
```{r}
data_frames <- list()


for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Nakhon.Si.Thammarat[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(Nakhon.Si.ThammaratD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  

  data_frames[[year - 2005]] <- data_frame
}


for (i in seq_along(data_frames)) {
  assign(paste0("Nakhon.Si.Thammarat.", i + 2005), data_frames[[i]])
}

correlation <- cor(SuratThani_year$Dengue, SuratThani_year$Humidity)
```

###Correlation 
```{r}
correlation <- cor(SuratThani_year$Dengue, SuratThani_year$Humidity)
```
 
### Plotting GGPlots
```{r}


## Creating Dataframes for each Year 

data_frames <- list()


for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Nakhon.Si.Thammarat[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(Nakhon.Si.ThammaratD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  

  data_frames[[year - 2005]] <- data_frame
}


for (i in seq_along(data_frames)) {
  assign(paste0("Nakhon.Si.Thammarat.", i + 2005), data_frames[[i]])
}

correlation <- cor(SuratThani_year$Dengue, SuratThani_year$Humidity)


## Correlation

correlation <- cor(SuratThani_year$Dengue, SuratThani_year$Humidity)

## Plotting GGPlots 

plots <- list()

for (year in years) {

  Nakhon.Si.Thammarat_year <- get(paste0("Nakhon.Si.Thammarat.", year))
  Nakhon.Si.Thammarat_year$Dengue <- as.numeric(Nakhon.Si.Thammarat_year$Dengue)
  Nakhon.Si.Thammarat_year$Humidity <- as.numeric(Nakhon.Si.Thammarat_year$Humidity)

 
  plot <- ggplot(Nakhon.Si.Thammarat_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100.000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Nakhon.Si.Thammarat", year),
         x = "Months",
         y = "Dengue Cases per 100.000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100.000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100.000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))+
     annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)

  plots[[year - 2005]] <- plot
}

for (i in seq_along(plots)) {
  print(plots[[i]])
}
```



## 0.2 Preparing the Data for Monsoon Seasons 

### Bangkok
```{r}
#Months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
Months <- c(1,2,3,4,5,6,7,8,9,10,11,12)
Bangkok <- data.frame(T.huss.mon.ordered[3, ])
BangkokD <- data.frame(dengue.norm[3, ])


# Bangkok 2007
Humidity.2006 <- as.data.frame(t(Bangkok[,5:16]))
colnames(Humidity.2006)<-"Humidity.2006"

Dengue.2006 <- as.data.frame(t(BangkokD[,2:13]))
colnames(Dengue.2006) <- "Dengue.2006"

Bangkok.2006 <- data.frame(Humidity.2006=Humidity.2006,
                           Dengue.2006=Dengue.2006,
                           Months=Months)
# Bangkok 2007
Humidity.2007 <- as.data.frame(t(Bangkok[,17:28]))
colnames(Humidity.2007) <- "Humidity.2007"

Dengue.2007 <- as.data.frame(t(BangkokD[,14:25]))
colnames(Dengue.2007) <- "Dengue.2007"

Bangkok.2007 <- data.frame(Humidity.2007 = Humidity.2007,
                           Dengue.2007 = Dengue.2007,
                           Months = Months)

# Bangkok 2008
Humidity.2008 <- as.data.frame(t(Bangkok[,29:40]))
colnames(Humidity.2008) <- "Humidity.2008"

Dengue.2008 <- as.data.frame(t(BangkokD[,26:37]))
colnames(Dengue.2008) <- "Dengue.2008"

Bangkok.2008 <- data.frame(Humidity.2008 = Humidity.2008,
                           Dengue.2008 = Dengue.2008,
                           Months = Months)

# Bangkok 2009
Humidity.2009 <- as.data.frame(t(Bangkok[,41:52]))
colnames(Humidity.2009) <- "Humidity.2009"

Dengue.2009 <- as.data.frame(t(BangkokD[,38:49]))
colnames(Dengue.2009) <- "Dengue.2009"

Bangkok.2009 <- data.frame(Humidity.2009 = Humidity.2009,
                           Dengue.2009 = Dengue.2009,
                           Months = Months)

# Bangkok 2010 
Humidity.2010 <- as.data.frame(t(Bangkok[,53:64]))
colnames(Humidity.2010)<-"Humidity.2010"

Dengue.2010 <- as.data.frame(t(BangkokD[,50:61]))
colnames(Dengue.2010) <- "Dengue.2010"

Bangkok.2010 <- data.frame(Humidity.2010=Humidity.2010,
                           Dengue.2010=Dengue.2010,
                           Months=Months)


# Bangkok 2011
Humidity.2011 <- as.data.frame(t(Bangkok[,65:76]))
colnames(Humidity.2011) <- "Humidity.2011"

Dengue.2011 <- as.data.frame(t(BangkokD[,62:73]))
colnames(Dengue.2011) <- "Dengue.2011"

Bangkok.2011 <- data.frame(Humidity.2011 = Humidity.2011,
                           Dengue.2011 = Dengue.2011,
                           Months = Months)

# Bangkok 2012
Humidity.2012 <- as.data.frame(t(Bangkok[,77:88]))
colnames(Humidity.2012) <- "Humidity.2012"

Dengue.2012 <- as.data.frame(t(BangkokD[,74:85]))
colnames(Dengue.2012) <- "Dengue.2012"

Bangkok.2012 <- data.frame(Humidity.2012 = Humidity.2012,
                           Dengue.2012 = Dengue.2012,
                           Months = Months)

# Bangkok 2013
Humidity.2013 <- as.data.frame(t(Bangkok[,89:100]))
colnames(Humidity.2013) <- "Humidity.2013"

Dengue.2013 <- as.data.frame(t(BangkokD[,86:97]))
colnames(Dengue.2013) <- "Dengue.2013"

Bangkok.2013 <- data.frame(Humidity.2013 = Humidity.2013,
                           Dengue.2013 = Dengue.2013,
                           Months = Months)

# Bangkok 2014
Humidity.2014 <- as.data.frame(t(Bangkok[,101:112]))
colnames(Humidity.2014) <- "Humidity.2014"

Dengue.2014 <- as.data.frame(t(BangkokD[,98:109]))
colnames(Dengue.2014) <- "Dengue.2014"

Bangkok.2014 <- data.frame(Humidity.2014 = Humidity.2014,
                           Dengue.2014 = Dengue.2014,
                           Months = Months)

# DBangkok 2015
Humidity.2015 <- as.data.frame(t(Bangkok[,113:124]))
colnames(Humidity.2015) <- "Humidity.2015"

Dengue.2015 <- as.data.frame(t(BangkokD[,110:121]))
colnames(Dengue.2015) <- "Dengue.2015"

Bangkok.2015 <- data.frame(Humidity.2015 = Humidity.2015,
                           Dengue.2015 = Dengue.2015,
                           Months = Months)

# Bangkok 2016
Humidity.2016 <- as.data.frame(t(Bangkok[,125:136]))
colnames(Humidity.2016) <- "Humidity.2016"

Dengue.2016 <- as.data.frame(t(BangkokD[,122:133]))
colnames(Dengue.2016) <- "Dengue.2016"

Bangkok.2016 <- data.frame(Humidity.2016 = Humidity.2016,
                           Dengue.2016 = Dengue.2016,
                           Months = Months)

# Bangkok 2017
Humidity.2017 <- as.data.frame(t(Bangkok[,137:148]))
colnames(Humidity.2017) <- "Humidity.2017"

Dengue.2017 <- as.data.frame(t(BangkokD[,134:145]))
colnames(Dengue.2017) <- "Dengue.2017"

Bangkok.2017 <- data.frame(Humidity.2017 = Humidity.2017,
                           Dengue.2017 = Dengue.2017,
                           Months = Months)

# Bangkok 2018
Humidity.2018 <- as.data.frame(t(Bangkok[,149:160]))
colnames(Humidity.2018) <- "Humidity.2018"

Dengue.2018 <- as.data.frame(t(BangkokD[,146:157]))
colnames(Dengue.2018) <- "Dengue.2018"

Bangkok.2018 <- data.frame(Humidity.2018 = Humidity.2018,
                           Dengue.2018 = Dengue.2018,
                           Months = Months)

# Bangkok 2019
Humidity.2019 <- as.data.frame(t(Bangkok[,161:172]))
colnames(Humidity.2019) <- "Humidity.2019"

Dengue.2019 <- as.data.frame(t(BangkokD[,158:169]))
colnames(Dengue.2019) <- "Dengue.2019"

Bangkok.2019 <- data.frame(Humidity.2019 = Humidity.2019,
                           Dengue.2019 = Dengue.2019,
                           Months = Months)

# Bangkok 2020
Humidity.2020 <- as.data.frame(t(Bangkok[,173:184]))
colnames(Humidity.2020) <- "Humidity.2020"

Dengue.2020 <- as.data.frame(t(BangkokD[,170:181]))
colnames(Dengue.2020) <- "Dengue.2020"

Bangkok.2020 <- data.frame(Humidity.2020 = Humidity.2020,
                           Dengue.2020 = Dengue.2020,
                           Months = Months)


```

### Nakhon Si Thammarat

```{r}


Nakhon.Si.Thammarat <- data.frame(T.huss.mon.ordered[31, ])
Nakhon.Si.ThammaratD <- data.frame(dengue.norm[31, ])



data_frames <- list()


for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  humidity <- as.data.frame(t(Nakhon.Si.Thammarat[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(Nakhon.Si.ThammaratD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  

  data_frames[[year - 2005]] <- data_frame
}


for (i in seq_along(data_frames)) {
  assign(paste0("Nakhon.Si.Thammarat.", i + 2005), data_frames[[i]])
}

```


### Songkhla
```{r}
Months <- c(1,2,3,4,5,6,7,8,9,10,11,12)
Songkhla <- data.frame(T.huss.mon.ordered[64, ])
SongkhlaD <- data.frame(dengue.norm[64, ])
Months <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)


data_frames <- list()


for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  # Dataframe with Humidity , Dengue and Months for each Year
  humidity <- as.data.frame(t(Songkhla[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(SongkhlaD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  

  data_frames[[year - 2005]] <- data_frame
}


for (i in seq_along(data_frames)) {
  assign(paste0("Songkhla.", i + 2005), data_frames[[i]])
}
```

## 1. Plotting Dengue and Humidity Yearly Means
```{r}

ggplot(Dengue.Humidity.Plot, aes(x = Years)) +
  geom_line(aes(y = Dengue.mean, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Huss.mean, color = "Humidity in %"), type = "o") +
  labs(title = "Yearly Dengue Cases and Humidity", x = "Years", y = "Dengue Cases for 100.000") +
  geom_point(aes(y = Dengue.mean, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Huss.mean, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title"))+
   scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent))


```


## 2.1 Plotting Bangkok 
```{r}
# GGPlot for Bangkok 2006
ggplot(Bangkok.2006, aes(x = Months)) +
  geom_line(aes(y = Dengue.2006, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2006, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2006", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2006, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2006, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2007
ggplot(Bangkok.2007, aes(x = Months)) +
  geom_line(aes(y = Dengue.2007, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2007, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2007", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2007, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2007, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2008
ggplot(Bangkok.2008, aes(x = Months)) +
  geom_line(aes(y = Dengue.2008, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2008, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2008", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2008, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2008, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2009
ggplot(Bangkok.2009, aes(x = Months)) +
  geom_line(aes(y = Dengue.2009, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2009, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2009", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2009, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2009, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2010
ggplot(Bangkok.2010, aes(x = Months)) +
  geom_line(aes(y = Dengue.2010, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2010, color = "Humidity in %"), type = "o") +
  labs(title = " Dengue Cases and Humidity in Bangkok 2010", x = "Months", y = "Dengue Cases for 100.000") +
  geom_point(aes(y = Dengue.2010, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2010, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title"))+
   scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent))+
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2011
ggplot(Bangkok.2011, aes(x = Months)) +
  geom_line(aes(y = Dengue.2011, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2011, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2011", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2011, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2011, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2012
ggplot(Bangkok.2012, aes(x = Months)) +
  geom_line(aes(y = Dengue.2012, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2012, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2012", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2012, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2012, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2013
ggplot(Bangkok.2013, aes(x = Months)) +
  geom_line(aes(y = Dengue.2013, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2013, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2013", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2013, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2013, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2014
ggplot(Bangkok.2014, aes(x = Months)) +
  geom_line(aes(y = Dengue.2014, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2014, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2014", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2014, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2014, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2015
ggplot(Bangkok.2015, aes(x = Months)) +
  geom_line(aes(y = Dengue.2015, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2015, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2015", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2015, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2015, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2016
ggplot(Bangkok.2016, aes(x = Months)) +
  geom_line(aes(y = Dengue.2016, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2016, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2016", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2016, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2016, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2017 
ggplot(Bangkok.2017, aes(x = Months)) +
  geom_line(aes(y = Dengue.2017, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2017, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2017", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2017, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2017, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))


# GGPlot for Bangkok 2018
ggplot(Bangkok.2018, aes(x = Months)) +
  geom_line(aes(y = Dengue.2018, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2018, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2018", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2018, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2018, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2019
ggplot(Bangkok.2019, aes(x = Months)) +
  geom_line(aes(y = Dengue.2019, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2019, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2019", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2019, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2019, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Bangkok 2020
ggplot(Bangkok.2020, aes(x = Months)) +
  geom_line(aes(y = Dengue.2020, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2020, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Bangkok 2020", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2020, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2020, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```
## 2.1 Plotting Geographic different Regions 

### Maha- Sarakham
### Chiang Mai
```{r}
#DataFrame

Chiang.Mai <- data.frame(T.huss.mon.ordered[10, ])
Chiang.MaiD <- data.frame(dengue.norm[10, ])


# Chiang Mai 2006
Humidity.2006 <- as.data.frame(t(Chiang.Mai[,5:16]))
colnames(Humidity.2006) <- "Humidity.2006"

Dengue.2006 <- as.data.frame(t(Chiang.MaiD[,2:13]))
colnames(Dengue.2006) <- "Dengue.2006"

Chiang.Mai.2006 <- data.frame(Humidity.2006 = Humidity.2006,
                              Dengue.2006 = Dengue.2006,
                              Months = Months)

# Chiang Mai 2007
Humidity.2007 <- as.data.frame(t(Chiang.Mai[,17:28]))
colnames(Humidity.2007) <- "Humidity.2007"

Dengue.2007 <- as.data.frame(t(Chiang.MaiD[,14:25]))
colnames(Dengue.2007) <- "Dengue.2007"

Chiang.Mai.2007 <- data.frame(Humidity.2007 = Humidity.2007,
                              Dengue.2007 = Dengue.2007,
                              Months = Months)

# Chiang Mai 2008
Humidity.2008 <- as.data.frame(t(Chiang.Mai[,29:40]))
colnames(Humidity.2008) <- "Humidity.2008"

Dengue.2008 <- as.data.frame(t(Chiang.MaiD[,26:37]))
colnames(Dengue.2008) <- "Dengue.2008"

Chiang.Mai.2008 <- data.frame(Humidity.2008 = Humidity.2008,
                              Dengue.2008 = Dengue.2008,
                              Months = Months)

# Chiang Mai 2009
Humidity.2009 <- as.data.frame(t(Chiang.Mai[,41:52]))
colnames(Humidity.2009) <- "Humidity.2009"

Dengue.2009 <- as.data.frame(t(Chiang.MaiD[,38:49]))
colnames(Dengue.2009) <- "Dengue.2009"

Chiang.Mai.2009 <- data.frame(Humidity.2009 = Humidity.2009,
                              Dengue.2009 = Dengue.2009,
                              Months = Months)

# Chiang Mai 2010
Humidity.2010 <- as.data.frame(t(Chiang.Mai[,53:64]))
colnames(Humidity.2010) <- "Humidity.2010"

Dengue.2010 <- as.data.frame(t(Chiang.MaiD[,50:61]))
colnames(Dengue.2010) <- "Dengue.2010"

Chiang.Mai.2010 <- data.frame(Humidity.2010 = Humidity.2010,
                              Dengue.2010 = Dengue.2010,
                              Months = Months)

# Chiang Mai 2011
Humidity.2011 <- as.data.frame(t(Chiang.Mai[,65:76]))
colnames(Humidity.2011) <- "Humidity.2011"

Dengue.2011 <- as.data.frame(t(Chiang.MaiD[,62:73]))
colnames(Dengue.2011) <- "Dengue.2011"

Chiang.Mai.2011 <- data.frame(Humidity.2011 = Humidity.2011,
                              Dengue.2011 = Dengue.2011,
                              Months = Months)

# Chiang Mai 2012
Humidity.2012 <- as.data.frame(t(Chiang.Mai[,77:88]))
colnames(Humidity.2012) <- "Humidity.2012"

Dengue.2012 <- as.data.frame(t(Chiang.MaiD[,74:85]))
colnames(Dengue.2012) <- "Dengue.2012"

Chiang.Mai.2012 <- data.frame(Humidity.2012 = Humidity.2012,
                              Dengue.2012 = Dengue.2012,
                              Months = Months)

# Chiang Mai 2013
Humidity.2013 <- as.data.frame(t(Chiang.Mai[,89:100]))
colnames(Humidity.2013) <- "Humidity.2013"

Dengue.2013 <- as.data.frame(t(Chiang.MaiD[,86:97]))
colnames(Dengue.2013) <- "Dengue.2013"

Chiang.Mai.2013 <- data.frame(Humidity.2013 = Humidity.2013,
                              Dengue.2013 = Dengue.2013,
                              Months = Months)

# Chiang Mai 2014
Humidity.2014 <- as.data.frame(t(Chiang.Mai[,101:112]))
colnames(Humidity.2014) <- "Humidity.2014"

Dengue.2014 <- as.data.frame(t(Chiang.MaiD[,98:109]))
colnames(Dengue.2014) <- "Dengue.2014"

Chiang.Mai.2014 <- data.frame(Humidity.2014 = Humidity.2014,
                              Dengue.2014 = Dengue.2014,
                              Months = Months)

# Chiang Mai 2015
Humidity.2015 <- as.data.frame(t(Chiang.Mai[,113:124]))
colnames(Humidity.2015) <- "Humidity.2015"

Dengue.2015 <- as.data.frame(t(Chiang.MaiD[,110:121]))
colnames(Dengue.2015) <- "Dengue.2015"

Chiang.Mai.2015 <- data.frame(Humidity.2015 = Humidity.2015,
                              Dengue.2015 = Dengue.2015,
                              Months = Months)

# Chiang Mai 2016
Humidity.2016 <- as.data.frame(t(Chiang.Mai[,125:136]))
colnames(Humidity.2016) <- "Humidity.2016"

Dengue.2016 <- as.data.frame(t(Chiang.MaiD[,122:133]))
colnames(Dengue.2016) <- "Dengue.2016"

Chiang.Mai.2016 <- data.frame(Humidity.2016 = Humidity.2016,
                              Dengue.2016 = Dengue.2016,
                              Months = Months)

# Chiang Mai 2017
Humidity.2017 <- as.data.frame(t(Chiang.Mai[,137:148]))
colnames(Humidity.2017) <- "Humidity.2017"

Dengue.2017 <- as.data.frame(t(Chiang.MaiD[,134:145]))
colnames(Dengue.2017) <- "Dengue.2017"

Chiang.Mai.2017 <- data.frame(Humidity.2017 = Humidity.2017,
                              Dengue.2017 = Dengue.2017,
                              Months = Months)

# Chiang Mai 2018
Humidity.2018 <- as.data.frame(t(Chiang.Mai[,149:160]))
colnames(Humidity.2018) <- "Humidity.2018"

Dengue.2018 <- as.data.frame(t(Chiang.MaiD[,146:157]))
colnames(Dengue.2018) <- "Dengue.2018"

Chiang.Mai.2018 <- data.frame(Humidity.2018 = Humidity.2018,
                              Dengue.2018 = Dengue.2018,
                              Months = Months)

# Chiang Mai 2019
Humidity.2019 <- as.data.frame(t(Chiang.Mai[,161:172]))
colnames(Humidity.2019) <- "Humidity.2019"

Dengue.2019 <- as.data.frame(t(Chiang.MaiD[,158:169]))
colnames(Dengue.2019) <- "Dengue.2019"

Chiang.Mai.2019 <- data.frame(Humidity.2019 = Humidity.2019,
                              Dengue.2019 = Dengue.2019,
                              Months = Months)

# Chiang Mai 2020
Humidity.2020 <- as.data.frame(t(Chiang.Mai[,173:184]))
colnames(Humidity.2020) <- "Humidity.2020"

Dengue.2020 <- as.data.frame(t(Chiang.MaiD[,170:181]))
colnames(Dengue.2020) <- "Dengue.2020"

Chiang.Mai.2020 <- data.frame(Humidity.2020 = Humidity.2020,
                              Dengue.2020 = Dengue.2020,
                              Months = Months)

```

```{r}
#Plotting
# GGPlot for Chiang.Mai 2006
ggplot(Chiang.Mai.2006, aes(x = Months)) +
  geom_line(aes(y = Dengue.2006, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2006, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2006", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2006, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2006, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2007
ggplot(Chiang.Mai.2007, aes(x = Months)) +
  geom_line(aes(y = Dengue.2007, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2007, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2007", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2007, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2007, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2008
ggplot(Chiang.Mai.2008, aes(x = Months)) +
  geom_line(aes(y = Dengue.2008, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2008, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2008", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2008, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2008, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2009
ggplot(Chiang.Mai.2009, aes(x = Months)) +
  geom_line(aes(y = Dengue.2009, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2009, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2009", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2009, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2009, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2010
ggplot(Chiang.Mai.2010, aes(x = Months)) +
  geom_line(aes(y = Dengue.2010, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2010, color = "Humidity in %"), type = "o") +
  labs(title = " Dengue Cases and Humidity in Chiang.Mai 2010", x = "Months", y = "Dengue Cases for 100.000") +
  geom_point(aes(y = Dengue.2010, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2010, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title"))+
   scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent))+
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2011
ggplot(Chiang.Mai.2011, aes(x = Months)) +
  geom_line(aes(y = Dengue.2011, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2011, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2011", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2011, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2011, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2012
ggplot(Chiang.Mai.2012, aes(x = Months)) +
  geom_line(aes(y = Dengue.2012, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2012, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2012", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2012, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2012, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2013
ggplot(Chiang.Mai.2013, aes(x = Months)) +
  geom_line(aes(y = Dengue.2013, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2013, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2013", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2013, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2013, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2014
ggplot(Chiang.Mai.2014, aes(x = Months)) +
  geom_line(aes(y = Dengue.2014, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2014, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2014", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2014, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2014, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2015
ggplot(Chiang.Mai.2015, aes(x = Months)) +
  geom_line(aes(y = Dengue.2015, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2015, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2015", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2015, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2015, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2016
ggplot(Chiang.Mai.2016, aes(x = Months)) +
  geom_line(aes(y = Dengue.2016, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2016, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2016", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2016, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2016, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2017 
ggplot(Chiang.Mai.2017, aes(x = Months)) +
  geom_line(aes(y = Dengue.2017, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2017, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2017", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2017, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2017, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))


# GGPlot for Chiang.Mai 2018
ggplot(Chiang.Mai.2018, aes(x = Months)) +
  geom_line(aes(y = Dengue.2018, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2018, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2018", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2018, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2018, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2019
ggplot(Chiang.Mai.2019, aes(x = Months)) +
  geom_line(aes(y = Dengue.2019, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2019, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2019", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2019, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2019, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# GGPlot for Chiang.Mai 2020
ggplot(Chiang.Mai.2020, aes(x = Months)) +
  geom_line(aes(y = Dengue.2020, color = "Dengue Cases per 100.000"), type = "o") +
  geom_line(aes(y = Humidity.2020, color = "Humidity in %"), type = "o") +
  labs(title = "Dengue Cases and Humidity in Chiang.Mai 2020", x = "Months", y = "Dengue Cases per 100.000") +
  geom_point(aes(y = Dengue.2020, color = "Dengue Cases per 100.000"), shape = 16) +
  geom_point(aes(y = Humidity.2020, color = "Humidity in %"), shape = 16) +
  scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                     labels = c("Dengue Cases per 100.000", "Humidity in %"),
                     guide = guide_legend(title = "Legend Title")) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 100, name = "Humidity in %", labels = scales::percent)) +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```
```{r}
print(dengue.norm$X2020.07)
print(dengue.norm$X2010.07)

cor(dengue.norm$X2020.07,dengue.norm$X2010.07)

print(dengue.norm$X2020.02)
print(dengue.norm$X2010.02)

cor(dengue.norm$X2020.03,dengue.norm$X2010.03)
cor(dengue.norm$X2020.12,dengue.norm$X2010.12)
```

### Kalasi
### Chumphon
### Songkhla

## 2.2 Plotting Nakhon Si Thammarat,Songkhla, and Surat Thani 

The gulfstor hits Thailand the highest in the gulf of mexio . This is why I decided to focus on these three specific regions. I wanted to see if there were yearly differences fe. the Typhoon Ketsana hit Thailand in in 200

### 2.2.1 Nakhon Si Thammarat
```{r}

# Create a vector of years from 2006 to 2020
years <- 2006:2020

# Create an empty list to store the plots
plots <- list()

# Loop through each year
for (year in years) {
  # Convert the variables to numeric
  Nakhon.Si.Thammarat_year <- get(paste0("Nakhon.Si.Thammarat.", year))
  Nakhon.Si.Thammarat_year$Dengue <- as.numeric(Nakhon.Si.Thammarat_year$Dengue)
  Nakhon.Si.Thammarat_year$Humidity <- as.numeric(Nakhon.Si.Thammarat_year$Humidity)

  # Create the plot for the given year
  plot <- ggplot(Nakhon.Si.Thammarat_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100.000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Nakhon.Si.Thammarat", year),
         x = "Months",
         y = "Dengue Cases per 100.000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100.000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100.000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

  # Append the plot to the list
  plots[[year - 2005]] <- plot
}

# Print the plots
for (i in seq_along(plots)) {
  print(plots[[i]])
}

```

### 2.2.2 Songkhla
```{r}

# Create a vector of years from 2006 to 2020
years <- 2006:2020

# Create an empty list to store the plots
plots <- list()

# Loop through each year
for (year in years) {
  # Convert the variables to numeric
  Songkhla_year <- get(paste0("Songkhla.", year))
  Songkhla_year$Dengue <- as.numeric(Songkhla_year$Dengue)
  Songkhla_year$Humidity <- as.numeric(Songkhla_year$Humidity)

  # Create the plot for the given year
  plot <- ggplot(Songkhla_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100.000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Songkhla", year),
         x = "Months",
         y = "Dengue Cases per 100.000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100.000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100.000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100.000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(sec.axis = sec_axis(~. / 100, name = "Humidity in %", labels = scales::percent)) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

  # Append the plot to the list
  plots[[year - 2005]] <- plot
}

# Print the plots
for (i in seq_along(plots)) {
  print(plots[[i]])
}
```
###2.2.3 Surat Thani
```{r}
library(ggplot2)
library(scales)

# Create a vector of months
Months <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)

# Select data for Surat Thani
SuratThani <- data.frame(T.huss.mon.ordered[67, ])
SuratThaniD <- data.frame(dengue.norm[67, ])

# Create an empty list to store the data frames
data_frames <- list()

# Loop through each year
for (year in 2006:2020) {
  humidity_cols <- ((year - 2006) * 12 + 5):((year - 2006) * 12 + 16)
  dengue_cols <- ((year - 2006) * 12 + 2):((year - 2006) * 12 + 13)
  
  # Dataframe with Humidity, Dengue, and Months for each year
  humidity <- as.data.frame(t(SuratThani[, humidity_cols]))
  colnames(humidity) <- paste0("Humidity.", year)
  
  dengue <- as.data.frame(t(SuratThaniD[, dengue_cols]))
  colnames(dengue) <- paste0("Dengue.", year)
  
  data_frame <- data.frame(Humidity = humidity, Dengue = dengue, Months = Months)
  
  data_frames[[year - 2005]] <- data_frame
}

# Assign data frames to variables
for (i in seq_along(data_frames)) {
  assign(paste0("SuratThani.", i + 2005), data_frames[[i]])
}

# Create a vector of years from 2006 to 2020
years <- 2006:2020

# Create an empty list to store the plots
plots <- list()

# Loop through each year
for (year in years) {
  # Convert the variables to numeric
  SuratThani_year <- get(paste0("SuratThani.", year))
  SuratThani_year$Dengue <- as.numeric(SuratThani_year$Dengue)
  SuratThani_year$Humidity <- as.numeric(SuratThani_year$Humidity)
  
  # Calculate correlation
  correlation <- cor(SuratThani_year$Dengue, SuratThani_year$Humidity)
  
  # Create the plot for the given year
  plot <- ggplot(SuratThani_year, aes(x = Months)) +
    geom_line(aes(y = Dengue, color = "Dengue Cases per 100,000"), type = "o") +
    geom_line(aes(y = Humidity, color = "Humidity in %"), type = "o") +
    labs(title = paste("Dengue Cases and Humidity in Surat Thani", year),
         x = "Months",
         y = "Dengue Cases per 100,000") +
    geom_point(aes(y = Dengue, color = "Dengue Cases per 100,000"), shape = 16) +
    geom_point(aes(y = Humidity, color = "Humidity in %"), shape = 16) +
    scale_color_manual(values = c("Dengue Cases per 100,000" = "violet", "Humidity in %" = "orange"),
                       labels = c("Dengue Cases per 100,000", "Humidity in %"),
                       guide = guide_legend(title = "Legend Title")) +
    scale_y_continuous(
      name = "Dengue Cases per 100,000",
      breaks = seq(0, 200, by = 25),
      limits = c(0, 200)
    ) +
    scale_y_continuous(
      sec.axis = sec_axis(
        ~ .,
        name = "Humidity in %",
        labels = scales::percent,
        breaks = seq(0, 10, by = 1)
      )
    ) +
    scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
    annotate("text", x = 6, y = 50, label = paste0("Correlation: ", round(correlation, 2)), color = "black", size = 4, vjust = -0.5)
  
  # Append the plot to the list
  plots[[year - 2005]] <- plot
}

# Print the plots
for (i in seq_along(plots)) {
  print(plots[[i]])
}

```


## 4. Neuer Plot Try out 

```{r}
seals$z <- with(seals, sqrt(delta_long^2 + delta_lat^2)) m <- ggplot(seals, aes(long, lat))
m + geom_contour(aes(z = z))
x, y, z, alpha, colour, linetype, size, weight
```


