---
title: "ARIMA"
output: html_document
date: "2023-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "GAM"
output: html_document
date: "2023-06-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. load the dengue dataframe combined -> dengue_data_com
```{r}
dengue_data_com = read.csv("../dengue_data/dengue_data_com.csv")

head(dengue_data_com)

```

#2. get the data into a diffrent shape we need for time series analysis 
```{r}
#überlegung: dengue daten erstmal auf ganz thailand bezogen also für jeden monat der Jahre 2006-2020 die summer der dengue cases im ganzen land thailand berechnen 

# load the needed package 
library(tidyverse)

# Group the dataframe after year, month and calculate the dengue cases sum 
dengue_sum <- dengue_data_com %>%
  group_by(Year, Month) %>%
  summarise(`Dengue_Total` = sum(`dengue_cases`))

print(dengue_sum)

# generaten a new dataframe with coulmns year, month, dengue cases 
ts_dengue = data.frame 
```

#2.1 Remove the Total-values 
```{r}
i.remove = which(dengue_sum$Month %in% c("Total"))
i.remove 

dengue_sum = dengue_sum[-i.remove,]

#lets turn the months into numerics 

# Define the Change 
monat_umwandlung <- c(Jan = "01", Feb = "02", Mar = "03", Apr = "04", May = "05", June = "06",
                      July = "07", Aug = "08", Sept = "09", Oct = "10", Nov = "11", Dec = "12")

# Change the months into numbers 
dengue_sum <- dengue_sum %>%
  mutate(Month = recode(Month, !!!monat_umwandlung))

#Chnage the class from chr into dbl 
dengue_sum$Month = as.numeric(dengue_sum$Month)
print(dengue_sum)

#order the dataframe 

dengue_sum = dengue_sum[order(dengue_sum$Year, dengue_sum$Month),]
print(dengue_sum) 

```

#3. now lets convert the dataframe into a ts 
```{r}
# Convert to time series object
myTs <- ts(dengue_sum$Dengue_Total, frequency = 12, start = c(2006, 1))
class(myTs) # ts"  (if you don't just feed in one but more data points per time point e.g. env then it is mts)

# Plot time series
TSstudio::ts_plot(myTs, title = "Dengue Cases in Thailand between the years 2006 - 2020", Xgrid = TRUE)

```

#4. lets have a look on the data 
```{r}
# Loess smoothing, A periodic decomposition
# decompose() assumes the seasonal term to be fixed and repeating (constant throughout the entire time series)
# stl() is more flexible and automatically determines the seasonal period from the data
decomposedSeasonsal <- stl(myTs, s.window = "periodic")

# Set the desired width, height, and resolution
width <- 1200
height <- 1200
res <- 300

# Save the plot with higher resolution
png("decomposition_plot.png", width = width, height = height, res = res)
plot(decomposedSeasonsal, main = "Decomposition of a periodic time series", xaxt = "n")
grid(lty = 2, col = "gray")
dev.off()

```
#5. Test for stationarity 
```{r}
# Test for Stationarity
# Stationarity is an important assumption in many statistical tests  and ts analysis models. 
# A ts is stationary when the mean and variance, do not change over time. 

# Augmented Dickey-Fuller (ADF)
# includes additional lagged differences of the series in the regression equation, 
# capturing different types of non-stationarity, identifying both trend and autoregressive components

library(tseries)
adf.test(myTs) # p-value < 0.05 indicates the TS is stationary --> suggests it is stationary !!!
```

#5.1 Test auf subset of myTS (differnet years) 
```{r}
# Subset of the time series 
start_year <- 2012
end_year <- 2015
subset_ts <- window(myTs, start = c(start_year, 1), end = c(end_year, 12))

# Dickey-Fuller-Test on the choosen subset 
 adf.test(subset_ts)
```
#6. Test for Autocorr 
```{r}
# PACF & ACF: 
# Which lag fits the data best? --> order
# Autocorrelation is the correlation of a Time Series with lags of itself
width = 1400
height = 1000
res = 300 

# autocorrelation
# correlation of the time series with a lag of itself
acfRes <- acf(dengue_sum$Dengue_Total)  # generates plots by default
# Save the plot as an image file (e.g., PNG format)
png("acf_plot.png", width = width, height = height, res = res)
plot(acfRes, main = "Autocorrelation Function (ACF) Plot")
dev.off()

# partial autocorrelation
# correlation of the time series with a lag of itself, 
# with the linear dependence of all the lags between them removed
pacfRes <- pacf(dengue_sum$Dengue_Total)# generates plots by default
# Save the plot as an image file (e.g., PNG format)
png("pacf_plot.png", width = width, height = height, res = res)
plot(pacfRes, main = "Partial autocorrelation Function (pACF) Plot")
dev.off()
```

#7. Lets model 
```{r}
library(forecast)
best_model_auto  <- auto.arima(myTs, seasonal = TRUE)
best_model_auto
```
```{r}
summary(best_model_auto)
```

#8. FORCASTING
```{r}
# Generate forecasts
library(forecast)
horizon <- 24 # number of units you want to forecast --> here months 
forecasted <- forecast(best_model_auto, h = horizon)
class(forecasted) # forcast

# Access forecasted values and intervals
forecastedTs <- forecasted$mean
forecastedTs  <- ts(forecastedTs, frequency = 12, start = c(2020, 12) )
class(forecastedTs) 
plot(forecastedTs)

# some formatting for gglotting
forecastDf <- as.data.frame(forecasted)
forecastDf$time <- time(forecastedTs)
names(forecastDf) <- c("point", "lo_80", "hi_80" , "lo_95", "hi_95", "time")
class(forecastDf)

# Plot both the original ts and the forecasts
Last_plot <- ggplot() +
  geom_line(data = myTs, aes(x = time(myTs), y = myTs), color = "black") +
  geom_ribbon(data = forecastDf, aes(x = time, ymin = lo_95, ymax = hi_95), fill = "gray80") +
  geom_line(data = forecastDf, aes(x = time, y = point), color = "blue") +
  labs(title = "Time Series Forecast") +
  xlab("Year") +
  ylab("Dengue Cases") +
  theme_minimal()
  ggsave("arima_model.png",plot = Last_plot)
Last_plot
```

#9. PREDICTION EVALUATION
```{r}
# Here are some evaluation metrics to compare our predicted and our actual data
original <- ts(dengue_sum$Dengue_Total[157:180], frequency = 12 , start = c(2019,1) ) # create an object containing the original ts of 2018 - 2020

keep = which(dengue_sum_2$Year %in% c(2019:2020))

dengue_sum_test = dengue_sum_2[keep,]

head(dengue_sum_test)
dengue_sum_test

predicted <- forecastedTs

library(Metrics)
# lower values indicate better accuracy:

plot(abs(original-predicted)) # absolute difference

# Mean Absolute Error (MAE):
mae_value <- mae(predicted, original)
mae_value
# mean(abs(predicted - original))
# -> on average, the forecasted cases deviate from the actual case number by approximately 524.74 cases

# Root Mean Squared Error (RMSE):
rmse_value <- rmse(predicted, original)
rmse_value
# sqrt(mean((predicted - original)^2))
# -> a typical magnitude of the forecast errors was 601.65 cases

# Mean Absolute Percentage Error (MAPE):
mape <- mape(predicted, original)*100
mape
# mean(abs((predicted - original) / original)) * 100
# -> the average percentage difference between the predicted values and the actual values was 12.44%


```

