---
title: "GAM"
output: html_document
date: "2023-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "GAM_modeling"
output: html_document
date: "2023-07-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#install the packages 
```{r}
install.packages("readxl")
install.packages("nlme")
install.packages("gam")
install.packages("sdm")
install.packages("raster")
install.packages("ncdf4")
```

```{r}
library(readxl)
library(nlme)
library(gam)
library(sdm)
library(raster)
library(ncdf4)
library(mgcv)
```

#first lets vis. the dengue_cases and the corresponding relative_humidity 
```{r}
# Lade die erforderlichen Bibliotheken
#library(ggplot2)

# Erstelle den Plot
#ggplot(dengue_hum, aes(x = Year, y = Dengue_Cases, z = Relative_Humidity)) +
  #geom_point() +
  #labs(y = "Dengue Cases", z = "Relative Humidity", title = "Dengue Cases vs. Relative Humidity")
```

#1.load and attache the new data 
```{r}
data2 <- read_excel("../dengue_data/dengue_hum.xlsx")
View(data2)
newdata = data.frame(data2)
```


```{r}
#colnames(newdata)[8] <- "RH"
```


##1.1 Model our gam 
```{r}
model1 <- gam(newdata$Dengue_Cases~s(newdata$Relative_Humidity, k = 9),family="quasipoisson")
```

##1.2 Plot the GAM Model
```{r}
plot.gam(model1)
```
##1.3 Have a look at the parametrics 
```{r}
summary(model1)
```

##1.4 GAM evaluation - k value 
```{r}
k.check(model1)
```
-> the model indicates that k = 9 is the best fit 

##1.5 GAM evaluation 
```{r}
gam.check(model1)
```
in a “perfect” model: - the q-q plot should no curves- the histogram have to be bell-shaped- residuals values should be evenly distributed around 0- the response vs fitted values should clusters around a one-one line- p-value should not be too low -> indicates that k is too low, especially if edf is close to k- gcv should be small 


```{r}
linear_model <- lm(newdata$Dengue_Cases~newdata$Relative_Humidity)
smooth_model <- gam(newdata$Dengue_Cases~s(newdata$Relative_Humidity))
```


```{r}
AIC(linear_model, smooth_model)
```
-> the AIC value is slightly smaller for the smooth_model (184511.8) compared to the liner_model (184555.0)
-> a smooth_model is be a better fit to model the data 

#1.6 predict the values 
```{r}
hurs <- stack ("hurs_SEA22_MPI_rcp85_2021-2040_grid_daymean_monmean_swm_timmean_thai.nc")
names(hurs) <- "Relative_Humidity"

newdata = data.frame(newdata)

p1 <- raster::predict(data = hurs, model1, type="response")

```


#1.7 raster plot 
```{r}
thailand <- raster::getData('GADM', country='THA', level=1) 
my.area <- extent(thailand) 
```
```{r}
library(terra)
my.p1 <- crop(p1, my.area)
my.p1.mask <- mask(my.p1, thailand)

?crop
```

# the extent of a map describes the area visible on the 
mapmy.p0 <- crop(p0, my.area) 
# crop returns a geographic subset of an object as specified by an Extent object 
my.p0.mask <- mask(my.p0, thailand) 
# mask function creates a MASK with values 1 and NULL, actually it plots 
# values only over Thailand, the rest are setting to 
zeropdf(file="thailand_precipitation__dengue.pdf") 
# save the plot in pdf in your 
directoryplot(my.p0.mask, legend.width=1, legend.shrink=0.75, axes= FALSE, box=FALSE) title("Predicted values: Influence of precipitation on dengue", line = -1)  
plot(thailand, add=TRUE) 
dev.off()
```

# 1 months lag ***attach dataframe***

```{r}
attach(GAM_data_1_Lag)
```
