---
title: "Clustering_Dengue_Data"
author: "Manon Manddernach"
date: "2023-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Loading Dengue Data Normalized and Clima Data and Geographic Data

I used the Dengue Data normalized with Population Data, Humidity Data and Geography Data. Because we sometimes used the same names for different Data frames , it is necessary to run this chunk in each Markdown.

```{r}
# Dengue Data
dengue.norm <- read.csv("../dengue_data/dengue_without_NAs_norm.csv")
dengue.norm <- subset(dengue.norm, select = setdiff(colnames(dengue.norm), c("X", "X.1","X2006.total", "X2007.total", "X2008.total","X2009.total", "X2010.total", "X2011.total","X2012.total", "X2013.total", "X2014.total","X2015.total", "X2016.total", "X2017.total","X2018.total", "X2019.total", "X2020.total")))
dengue.norm['numbers']=c(1:77)
# Replace Bungkan with Nong Khai Data 
dengue.norm[4,2:181] <- dengue.norm[35,2:181]
#dengue.norm <- replace(dengue.norm, is.na(dengue.norm), 0)
as.data.frame(dengue.norm)

## Humidity Data 
T.huss.mon.ordered <- read.csv("../geo_data/T.huss.mon_correct_provinces.csv")
T.huss.mon.ordered <- subset(T.huss.mon.ordered, select = -X)
T.huss.mon.ordered$Number <- c(1:77)
names(T.huss.mon.ordered)[names(T.huss.mon.ordered) == "Number"] <- "numbers"

## Geo Data 
shapefilePath <- "../geo_data/gadm36_THA_shp"
layers <- st_layers(shapefilePath)
geography <- st_read(shapefilePath, layer = "gadm36_THA_1")
numbers <- c(1:77)
geography <- cbind(geography,numbers)
```

# 1. Clustering each Years Dengue Cases

## 1.1 Creating a list with yearly Dengue Data

```{r}
dengue.norm.2006 <- as.data.frame(dengue.norm[,2:13])
dengue.norm.2007 <- as.data.frame(dengue.norm[,14:25])
dengue.norm.2008 <- as.data.frame(dengue.norm[,26:37])
dengue.norm.2009 <- as.data.frame(dengue.norm[,38:49])
dengue.norm.2010 <- as.data.frame(dengue.norm[,50:61])
dengue.norm.2011 <- as.data.frame(dengue.norm[,62:73])
dengue.norm.2012 <- as.data.frame(dengue.norm[,74:85])
dengue.norm.2013 <- as.data.frame(dengue.norm[,86:97])
dengue.norm.2014 <- as.data.frame(dengue.norm[,98:109])
dengue.norm.2015 <- as.data.frame(dengue.norm[,110:121])
dengue.norm.2016 <- as.data.frame(dengue.norm[,122:133])
dengue.norm.2017 <- as.data.frame(dengue.norm[,134:145])
dengue.norm.2018 <- as.data.frame(dengue.norm[,146:157])
dengue.norm.2019 <- as.data.frame(dengue.norm[,158:169])
dengue.norm.2020 <- as.data.frame(dengue.norm[,170:181])

dengue.norm.list <- list(
  dengue.norm.2006=dengue.norm.2006,
  dengue.norm.2007=dengue.norm.2007,
  dengue.norm.2008=dengue.norm.2008,
  dengue.norm.2009=dengue.norm.2009,
  dengue.norm.2010=dengue.norm.2010,
  dengue.norm.2011=dengue.norm.2011,
  dengue.norm.2012=dengue.norm.2012,
  dengue.norm.2013=dengue.norm.2013,
  dengue.norm.2014=dengue.norm.2014,
  dengue.norm.2015=dengue.norm.2015,
  dengue.norm.2016=dengue.norm.2016,
  dengue.norm.2017=dengue.norm.2017,
  dengue.norm.2018=dengue.norm.2018,
  dengue.norm.2019=dengue.norm.2019,
  dengue.norm.2020=dengue.norm.2020
)
```

## 1.2.Clustering the yearly Dengue Data from each Province from 2006 to 2020

```{r}


# Überprüfen, ob es NA/NaN/Inf-Werte gibt
print(any(is.na(dengue.norm)))


```

```{r}
#dengue.norm[4,2:61] <- dengue.norm[35,2:61]
print(dengue.norm)

for (year in 2006:2020){

k <-3


dengue.norm.year <- get(paste0("dengue.norm.",year))
dengue.norm.year.mean<-as.data.frame(apply(dengue.norm.year,1, mean))
dengue.norm.year.mean ['numbers'] = c(1:77)

dengue.norm.kmeans.years <-kmeans(dengue.norm.year.mean$`apply(dengue.norm.year, 1, mean)`, centers = k )



## Zuordnen aller original values  of cluster 1,2,3 their corresponding center 

dengue.norm.centers.years <- as.vector(dengue.norm.kmeans.years$centers)
dengue.norm.centers.clusters.years <- data.frame(Value = dengue.norm.kmeans.years$cluster, Center = NA)

for (i in 1:length(dengue.norm.kmeans.years$cluster)) {
  dengue.norm.centers.clusters.years$Center[i] <- dengue.norm.centers.years[dengue.norm.kmeans.years$cluster[i]]
}

dengue.norm.centers.clusters.years['numbers'] = c(1:77)
dengue.norm.centers.clusters.years.geography = merge(dengue.norm.centers.clusters.years, geography, by= "numbers")


## ggplot
print(ggplot() +
  geom_sf(data = dengue.norm.centers.clusters.years.geography ,aes(geometry = geometry, fill= Center)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = paste("kmeans K = 3  Dengue Cases of Provinces",year)) +
  #scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")+
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  guides(fill = guide_legend(title = "cases for 100.000 citizens", override.aes = list(shape = 22)))
)

}


```

```{r}
View(dengue.norm.year.mean)
```

# 2. Plotting the Centers of the Clusters from 2006 to 2020

## 2.1 Plotting the Clusters in 3 single Plots

```{r}


dengue.norm.2006.mean <- as.data.frame(apply(dengue.norm.2006,1,mean))
dengue.norm.2007.mean <- as.data.frame(apply(dengue.norm.2007,1,mean))
dengue.norm.2008.mean <- as.data.frame(apply(dengue.norm.2008,1,mean))
dengue.norm.2009.mean <- as.data.frame(apply(dengue.norm.2009,1,mean))
dengue.norm.2010.mean <- as.data.frame(apply(dengue.norm.2010,1,mean))
dengue.norm.2011.mean <- as.data.frame(apply(dengue.norm.2011,1,mean))
dengue.norm.2012.mean <- as.data.frame(apply(dengue.norm.2012,1,mean))
dengue.norm.2013.mean <- as.data.frame(apply(dengue.norm.2013,1,mean))
dengue.norm.2014.mean <- as.data.frame(apply(dengue.norm.2014,1,mean))
dengue.norm.2015.mean <- as.data.frame(apply(dengue.norm.2015,1,mean))
dengue.norm.2016.mean <- as.data.frame(apply(dengue.norm.2016,1,mean))
dengue.norm.2017.mean <- as.data.frame(apply(dengue.norm.2017,1,mean))
dengue.norm.2018.mean <- as.data.frame(apply(dengue.norm.2018,1,mean))
dengue.norm.2019.mean <- as.data.frame(apply(dengue.norm.2019,1,mean))
dengue.norm.2020.mean <- as.data.frame(apply(dengue.norm.2020,1,mean))



dengue.norm.kmeans.2006 <- kmeans(dengue.norm.2006.mean,centers = k)
dengue.norm.kmeans.2007 <- kmeans(dengue.norm.2007.mean,centers = k)
dengue.norm.kmeans.2008 <- kmeans(dengue.norm.2008.mean,centers = k)
dengue.norm.kmeans.2009 <- kmeans(dengue.norm.2009.mean,centers = k)
dengue.norm.kmeans.2010 <- kmeans(dengue.norm.2010.mean,centers = k)
dengue.norm.kmeans.2011 <- kmeans(dengue.norm.2011.mean,centers = k)
dengue.norm.kmeans.2012 <- kmeans(dengue.norm.2012.mean,centers = k)
dengue.norm.kmeans.2013 <- kmeans(dengue.norm.2013.mean,centers = k)
dengue.norm.kmeans.2014 <- kmeans(dengue.norm.2014.mean,centers = k)
dengue.norm.kmeans.2015 <- kmeans(dengue.norm.2015.mean,centers = k)
dengue.norm.kmeans.2016 <- kmeans(dengue.norm.2016.mean,centers = k)
dengue.norm.kmeans.2017 <- kmeans(dengue.norm.2017.mean,centers = k)
dengue.norm.kmeans.2018 <- kmeans(dengue.norm.2018.mean,centers = k)
dengue.norm.kmeans.2019 <- kmeans(dengue.norm.2019.mean,centers = k)
dengue.norm.kmeans.2020 <- kmeans(dengue.norm.2020.mean,centers = k)

```

```{r}

## Creating a Data Frame of each year with a vector of the sorted centers 

years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)

dengue.norm.Yearly.Centers <- data.frame()

for (year in years) {
  dengue.norm.kmeans.year <- get(paste0("dengue.norm.kmeans.", year))# Annahme: dengue.norm.kmeans.2006, dengue.norm.kmeans.2007, etc do already excist
  dengue.norm.centers.sorted <- sort(dengue.norm.kmeans.year$centers)
  dengue.norm.centers.sorted.vector <- as.vector(dengue.norm.centers.sorted)
  
  if (ncol(dengue.norm.Yearly.Centers) == 0) {
    dengue.norm.Yearly.Centers <- data.frame(dengue.norm.centers.sorted.vector)
  } else {
    dengue.norm.Yearly.Centers <- cbind(dengue.norm.Yearly.Centers, dengue.norm.centers.sorted.vector)
  }
}

colnames(dengue.norm.Yearly.Centers) <- years
print(dengue.norm.Yearly.Centers)
```

```{r}
print(max(dengue.norm.Yearly.Centers))
print(min(dengue.norm.Yearly.Centers))

```

```{r}
## Clusters in 3 single ggPlots 
Dengue1<- dengue.norm.Yearly.Centers[1,]
Dengue2 <- dengue.norm.Yearly.Centers[2,]
Dengue3 <- dengue.norm.Yearly.Centers[3,]

ggplot() +
  plot(years, Dengue1, type="o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020",ylim=c(0,110))+
  plot(years, Dengue2, type = "o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020",ylim=c(0,110))+
  plot(years, Dengue3, type = "o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020", ylim=c (0,110))
```

## 2.2 Plotting the Clusters in one Plot

```{r}
##New Data Frame: Rows switch to columns and columns to rows

DengueLow<- unlist(dengue.norm.Yearly.Centers[1,])
DengueLow<- unname(DengueLow)

DengueMedium<- unlist(dengue.norm.Yearly.Centers[2,])
DengueMedium<- unname(DengueMedium)

DengueHigh<- unlist(dengue.norm.Yearly.Centers[3,])
DengueHigh<- unname(DengueHigh)

dengue.norm.Yearly.Centers.Plot <-data.frame(Years= years,
                                 DengueLow = DengueLow,
                                 DengueMedium = DengueMedium,
                                 DengueHigh = DengueHigh)

ggplot(dengue.norm.Yearly.Centers.Plot,aes(x=years))+
  geom_line(aes(y=DengueLow), color = "pink",type="o")+
  geom_line(aes(y=DengueMedium), color = "orange",type ="o")+
  geom_line(aes(y=DengueHigh), color = "violet", type= "o")+
  labs(title="Yearly Dengue Clusters ",x = "Years" , y= " Dengue Cases for 100.000")+
  geom_point(aes(y = DengueLow), color = "pink", shape = 16) +
  geom_point(aes(y = DengueMedium), color = "orange", shape = 16) +
  geom_point(aes(y = DengueHigh), color = "violet", shape = 16, show.legend = TRUE)+
  scale_color_manual(values = c("low" = "pink", "medium" = "orange", "high" = "violet"),
                     labels = c("low", "medium", "high"),
                     name = "Dengue Cases for 100.00")+
   scale_x_continuous(breaks = 2006:2020, labels = c("2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020"))

```

```{r}
cor()
```
