---
title: "Clustering_Clima_Data"
author: "Manon Manddernach"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sp)
library(sf)
library(terra)
library(dplyr)
library(raster)
library(ggplot2)
library(exactextractr)
```

## 1) loading Vietnam Climadata

```{r}
library(readxl)

sheet_names = excel_sheets("../era5_data_2006_2020_thailand_monmean.xlsx")

#laden von den einzelnen sheets
T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
T.tp.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_tp_2006_2020_thailand_monm")
T.t2m.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_t2m_2006_2020_thailand_mon")

```


### 1.1. Getting an Overwiev over T.huss.mon

```{r}
View(T.huss.mon)
View(T.huss.mon.ordered)
View(centroids.ordered)
```


### 1.2. Creating a map of Thailand

```{r}
# Code adds geometric layer of polygons stored in the data object

ggplot() +
  geom_sf(data = centroids.geography, aes(geometry=geometry),col = "black", fill = NA)+ labs(x= "Longitude",y= "Latitude" )

# ggplot () adds an empty ggplot Object , which can be layerd with different geometric layers
# genom_sf()verwendet geografische Daten , um im Simple Features Format zu plotten  
# aes(geometry= .. ) stellt sicher , das die Geometriedaten (Longlituda und Latitude an)
```


```{r}

# Code adds geometric layer of polygons stored in the data object

ggplot() +
  geom_sf(data = centroids.geography, aes(geometry=geometry),col = "black", fill = NA)+ labs(x= "Longitude",y= "Latitude" )

# ggplot () adds an empty ggplot Object , which can be layerd with different geometric layers
# genom_sf()verwendet geografische Daten , um im Simple Features Format zu plotten  
# aes(geometry= .. ) stellt sicher , das die Geometriedaten (Longlituda und Latitude an)

```

# 2. Try Clustering the means of each province over the whole time intervall

### 2.1 Plotten einer Map für die Means aller Provinzen über diee Jahre 2006 bis 2020

### 2.1.1 Prepearing the Province Mean Data for ggplot

```{r}
# 1.using monthlx data of each province as a mean over all reported months

mean.huss.prov.orderd= apply(T.huss.mon.ordered[-c(1,4)], 1, mean)

mean.huss.prov.orderd.df = as.data.frame(mean.huss.prov.orderd)

# 2.adding another column of numberst to merge the dataframes
mean.huss.prov.orderd.df['numbers'] = c(1:77)

# 3.merging the two dataframes of the mean relative humidtiy and geography
mean.huss.prov.geography = merge(mean.huss.prov.orderd.df, geography, by= "numbers")

View(mean.huss.prov.geography)
```


### 2.1.2 Printing the ggplot for mean humidity of Provinces from 2006 to 2020

```{r}
ggplot() +
  geom_sf(data = mean.huss.prov.geography,aes(geometry = geometry), fill= mean.huss.prov.orderd) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "Mean humidity of Provinces from 2006 to 2020") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
  
```

## 2.2 GGPlotten von den Clustern auf die Map von Thailand ; Ziel jeder Cluster hat eine andere Farbe

### 2.2.1. Clustern für K=3

```{r}
k <-3
mean.huss.prov.orderd.clusterd <-kmeans(mean.huss.prov.orderd.matrix, centers = k )

## print the results 
print(mean.huss.prov.orderd.clusterd$cluster)
print(mean.huss.prov.orderd.clusterd$centers)

## Wichtig! Umwandeln in Data Frame 
mean.huss.prov.orderd.clusters = as.data.frame(mean.huss.prov.orderd.clusterd$cluster)

## Vorbereiten zum Plotten :

View(mean.huss.prov.orderd.clusters)

# adding another column of numberst to merge the dataframes
mean.huss.prov.orderd.clusters['numbers'] = c(1:77)

#merging the two dataframes of the mean relative humidtiy and geography
mean.huss.prov.orderd.clusters.geography = merge(mean.huss.prov.orderd.clusters, geography, by= "numbers")

```


```{r}
View(mean.huss.prov.orderd.clusters)

```

```{r}
## ggPlotting the clusters (k=3)von den Mittelwertwn aller Provinzen von 2006 bis 2020

ggplot() +
  geom_sf(data = mean.huss.prov.orderd.clusters.geography ,aes(geometry = geometry), fill= mean.huss.prov.orderd.clusterd$cluster) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces from 2006 to 2020") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
```

###  2.2.2 für k= 4

```{r}
class(mean.huss.prov.orderd.matrix)
```

```{r}
## Commuting forkmeans  k=4
k <-4
mean.huss.prov.orderd.clusterd.4 <-kmeans(mean.huss.prov.orderd.matrix, centers = k )

## print the results 
print(mean.huss.prov.orderd.clusterd.4$cluster)
print(mean.huss.prov.orderd.clusterd.4$centers)

## Wichtig! Umwandeln in Data Frame 
mean.huss.prov.orderd.clusters.4 = as.data.frame(mean.huss.prov.orderd.clusterd.4$cluster)

## Vorbereiten zum Plotten :

View(mean.huss.prov.orderd.clusters.4)

# adding another column of numberst to merge the dataframes
mean.huss.prov.orderd.clusters.4['numbers'] = c(1:77)

#merging the two dataframes of the mean relative humidtiy and geography
mean.huss.prov.orderd.clusters.geography.4 = merge(mean.huss.prov.orderd.clusters.4, geography, by= "numbers")
```

```{r}
## ggPlotting the clusters (k=4)von den Mittelwertwn aller Provinzen von 2006 bis 2020

ggplot() +
  geom_sf(data = mean.huss.prov.orderd.clusters.geography.4 ,aes(geometry = geometry), fill= mean.huss.prov.orderd.clusterd.4$cluster) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 4  humidity of Provinces from 2006 to 2020") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
```


## 2.3 Kreieren einer Liste mit den Werten über alle Jahre 

```{r}
numbers <- c(1:77)
T.huss.2006 <- as.data.frame(numbers)
T.huss.2006['ADM1_EN'] = T.huss.mon.ordered$ADM1_EN

# View(T.huss.2006)

T.6 = as.data.frame(T.huss.mon.ordered[,4:15])
T.6['numbers'] = c(1:77)

# View(T.6)

T.huss.2006 =left_join(x=T.huss.2006, y=T.6 , by=NULL)

View(T.huss.2006)

T.huss.mean.2006  = as.data.frame(apply(T.huss.2006[-c(1,2)], 1, mean))
T.huss.mean.2006['numbers'] = c(1:77)
T.huss.mean.2006=right_join(T.huss.mean.2006,geography,by=NULL)

View(T.huss.mean.2006)

```

## 2.3.

```{r}
ggplot() +
  geom_sf(data = T.huss.mean.2006, aes(geometry = geometry), fill= apply(T.huss.2006[-c(1,2)], 1, mean)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "Mean Humidity in 2006") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
```

## 2.3.2 Clustering von 2006 K=3

```{r}
## Commuting forkmeans  k=3

## Um zu Clustern benötigt man eine Matrix 

T.huss.mean.2006  = as.data.frame(apply(T.huss.2006[-c(1,2)], 1, mean))

class(T.huss.mean.2006)

k <- 3
T.huss.mean.2006.clusterd.3 <- kmeans(T.huss.mean.2006, centers = k )


## print the results 
print(T.huss.mean.2006.clusterd.3$cluster)
print(T.huss.mean.2006.clusterd.3$centers)

T.huss.mean.2006.clusters.3= as.data.frame(T.huss.mean.2006.clusterd.3$cluster)

## Vorbereiten zum Plotten :

#View(T.huss.mean.2006.clusters.3)

# adding another column of numbers to merge the dataframes
T.huss.mean.2006.clusters.3['numbers'] = c(1:77)

#merging the two dataframes of the mean relative humidtiy and geography
T.huss.mean.2006.clusters.3 = merge(T.huss.mean.2006.clusters.3, geography, by= "numbers")

#View(T.huss.mean.2006.clusters.3)

ggplot() +
  geom_sf(T.huss.mean.2006.clusters.3, mapping=aes(geometry = geometry), fill= T.huss.mean.2006.clusterd.3$cluster) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "Clusters of mean Humidity in 2006 k=3 ") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

```
