---
title: "Clustering_Clima_Data_2.0"
author: "Manon Manddernach"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sp)
library(sf)
library(terra)
library(dplyr)
library(raster)
library(ggplot2)
library(exactextractr)
library(readxl)
library(colorspace)
library(viridis)
library(viridisLite)
library(reshape2)
library(cluster)


sheet_names = excel_sheets("../era5_data_2006_2020_thailand_monmean.xlsx")

T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
T.tp.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_tp_2006_2020_thailand_monm")
T.t2m.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_t2m_2006_2020_thailand_mon")

```

GEOPLOT

Ziel ( Wie kann ich Eine Series Time Analysis machen , in der Clustere
a) jedes Jahr (mean)in Cluster einteile und Plotte Done b) 3 mal 5 Jahre
c) 5 mal 3 Jahre d) Vergleiche Welche Provinzen sich geändert haben. e)
Plotten von den Centern der Cluster 1, 2 ,3 von 2006 bis 2020 Yay hab
ich hinbekommen auch mit 2 verschiedenen Methoden f) Ändern der
Farbgebung , damit Cluster und Centren nicht unterschiedlich sind YAY
Hab ich bhinbekommen mit 2 verschiedenen Methoden :) g9 anwenden von
einer Methode auf DONE hab ich auch hinbekommen g) K-Means Elbow
Clustering machen i) Alignment der Einzelnen Provinzen und wie sie
liegen h) Ordnung im Code

```{r}
View(T.huss.mon.ordered)
```

# 1. Finding K : Silhouette Method

```{r}
set.seed(123)  # Set a seed for reproducibility
k.values <- 2:10  # Define the range of K values to evaluate

# Calculate the Silhouette coefficient for each value of K

silhouette_scores <- sapply(k.values, function(k) {
  kmeans_model <- kmeans(huss.2006, centers = k, nstart = 10)
  silhouette_avg <- silhouette(kmeans_model$cluster, dist(huss.2018))
  mean(silhouette_avg[, 3])
})

# Plot the Silhouette scores
plot(k.values, silhouette_scores, type = "b", pch = 19, frame = FALSE, xlab = "Number of Clusters (K)", ylab = "Silhouette Coefficient")

# Find the optimal number of clusters with the highest Silhouette score
optimal_k <- k.values[which.max(silhouette_scores)]
print(optimal_k)
```

# 2. Cluster each Year's Humidity

## 2.1 Creating a list with yearly Humidity Data

```{r}
huss.2006<- as.data.frame(T.huss.mon.ordered[,5:16])
huss.2007<- as.data.frame(T.huss.mon.ordered[,17:28])
huss.2008<- as.data.frame(T.huss.mon.ordered[,29:40])
huss.2009<- as.data.frame(T.huss.mon.ordered[,41:52])
huss.2010<- as.data.frame(T.huss.mon.ordered[,53:64])
huss.2011<- as.data.frame(T.huss.mon.ordered[,65:76])
huss.2012<- as.data.frame(T.huss.mon.ordered[,77:88])
huss.2013<- as.data.frame(T.huss.mon.ordered[,89:100])
huss.2014<- as.data.frame(T.huss.mon.ordered[,101:112])
huss.2015<- as.data.frame(T.huss.mon.ordered[,113:124])
huss.2016<- as.data.frame(T.huss.mon.ordered[,125:136])
huss.2017<- as.data.frame(T.huss.mon.ordered[,137:148])
huss.2018<- as.data.frame(T.huss.mon.ordered[,149:150])
huss.2019<- as.data.frame(T.huss.mon.ordered[,151:162])
huss.2020<- as.data.frame(T.huss.mon.ordered[,163:174])

huss_data <- list(
  huss.2006 = huss.2006,
  huss.2007 = huss.2007,
  huss.2008 = huss.2008,
  huss.2009 = huss.2009,
  huss.2010 = huss.2010,
  huss.2011 = huss.2011,
  huss.2012 = huss.2012,
  huss.2013 = huss.2013,
  huss.2014 = huss.2014,
  huss.2015 = huss.2015,
  huss.2016 = huss.2016,
  huss.2017 = huss.2017,
  huss.2018 = huss.2018,
  huss.2019 = huss.2019,
  huss.2020 = huss.2020
)

```

## 2.2 Find diffrent Methods for Coloring the Clusters on the Maps

```{r}
## TRY OUT COLOR CODING THE CLUSTERS PART 1 

k <-3

huss.2006
huss.2006.mean<-as.data.frame(apply(huss.2006,1, mean))
huss.2006.mean ['numbers'] = c(1:77)
kmeans.2006 <-kmeans(huss.2006.mean$`apply(huss.2006, 1, mean)`, centers = k )

print(kmeans.2006$cluster)
print(kmeans.2006$centers)

kmeans.2006.cluster = as.data.frame(kmeans.2006$cluster)
kmeans.2006.cluster['numbers'] = c(1:77)
kmeans.2006.clusters.geography = merge(kmeans.2006.cluster, geography, by= "numbers")

colnames(kmeans.2006.clusters.geography)[colnames(kmeans.2006.clusters.geography) == "kmeans.2006$cluster"] <- "Clusters2006"
kmeans.2006.clusters.geography$Clusters2006[kmeans.2006.clusters.geography$Clusters2006 == 1] <- 70.68727
kmeans.2006.clusters.geography$Clusters2006[kmeans.2006.clusters.geography$Clusters2006 == 2] <- 75.06804
kmeans.2006.clusters.geography$Clusters2006[kmeans.2006.clusters.geography$Clusters2006 == 3] <- 81.61583

kmeans.2006.cluster['numbers'] = c(1:77)

print(kmeans.2006$cluster)
print(kmeans.2006$centers)


ggplot() +
  geom_sf(data = kmeans.2006.clusters.geography ,aes(geometry = geometry, fill= Clusters2006)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces 2006") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

```

## 2.3 Clustering the yearly Humidity from each Province from 2006 to 2020

```{r}

##Clustering

huss.data <- list(
  huss.2006 = huss.2006,
  huss.2007 = huss.2007,
  huss.2008 = huss.2008,
  huss.2009 = huss.2009,
  huss.2010 = huss.2010,
  huss.2011 = huss.2011,
  huss.2012 = huss.2012,
  huss.2013 = huss.2013,
  huss.2014 = huss.2014,
  huss.2015 = huss.2015,
  huss.2016 = huss.2016,
  huss.2017 = huss.2017,
  huss.2018 = huss.2018,
  huss.2019 = huss.2019,
  huss.2020 = huss.2020
)




##Clustering

for (year in 2006:2020){

k <-3

huss.year <- get(paste0("huss.",year))
huss.year.mean<-as.data.frame(apply(huss.year,1, mean))
huss.year.mean ['numbers'] = c(1:77)

kmeans.years <-kmeans(huss.year.mean$`apply(huss.year, 1, mean)`, centers = k )



## Zuordnen aller original values  of cluster 1,2,3 their corresponding center 

centers.years <- as.vector(kmeans.years$centers)
centers.clusters.years <- data.frame(Value = kmeans.years$cluster, Center = NA)

for (i in 1:length(kmeans.years$cluster)) {
  centers.clusters.years$Center[i] <- centers.years[kmeans.years$cluster[i]]
}

centers.clusters.years['numbers'] = c(1:77)
centers.clusters.years.geography = merge(centers.clusters.years, geography, by= "numbers")


## ggplot
print(ggplot() +
  geom_sf(data = centers.clusters.years.geography ,aes(geometry = geometry, fill= Center)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = paste("kmeans K = 3  humidity of Provinces",year)) +
  #scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")+
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  guides(fill = guide_legend(title = "Legende", override.aes = list(shape = 22))))


  
}


```

# 3. Plottening the Centers of the Clusters von 2006 bis 2020

## 3.1.1 Creating a new Dataframe Yearly.Centers with the centers of each year

```{r}

## Creating a Data Frame of each year with a vector of the sorted centers 

years <- c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)

Yearly.Centers <- data.frame()

for (year in years) {
  kmeans.obj <- get(paste0("kmeans.", year))# Annahme: kmeans.2006, kmeans.2007, etc do already excist
  centers.sorted <- sort(kmeans.obj$centers)
  centers.sorted.vector <- as.vector(centers.sorted)
  
  if (ncol(Yearly.Centers) == 0) {
    Yearly.Centers <- data.frame(centers.sorted.vector)
  } else {
    Yearly.Centers <- cbind(Yearly.Centers, centers.sorted.vector)
  }
}

colnames(Yearly.Centers) <- years
#print(Yearly.Centers)
```

## 3.1.2 Plotting the Clusters in 3 single Plots

```{r}

## Clusters in 3 single ggPlots 
Humidity1<- Yearly.Centers[1,]
Humidity2 <- Yearly.Centers[2,]
Humidity3 <- Yearly.Centers[3,]

ggplot() +
  plot(years, Humidity1, type="o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020", ylim = c(65, 85))+
  plot(years, Humidity2, type = "o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020", ylim = c(65, 85))+
  plot(years, Humidity3, type = "o", xlab = "Years", ylab = "Humidity ", main = " Yearly Clusters from 2006 to 2020", ylim = c(65, 85))

```

## 3.2.1 Creating a new Dataframe Yearly.Centers.Plot with the centers of each year

```{r}
##New Data Frame: Rows switch to columns and columns to rows

HumidityLow<- unlist(Yearly.Centers[1,])
HumidityLow<- unname(HumidityLow)

HumidityMedium<- unlist(Yearly.Centers[2,])
HumidityMedium<- unname(HumidityMedium)

HumidityHigh<- unlist(Yearly.Centers[3,])
HumidityHigh<- unname(HumidityHigh)

Yearly.Centers.Plot <-data.frame(Years= years,
                                 HumidityLow = HumidityLow,
                                 HumidityMedium = HumidityMedium,
                                 HumidityHigh = HumidityHigh)
#print(Yearly.Centers.Plot)
```

## 3.2.2 Plotting the Clusters in one Plot

```{r}

## Clusters in one ggplot

ggplot(Yearly.Centers.Plot,aes(x=years))+
  geom_line(aes(y=HumidityLow), color = "blue",type="o")+
  geom_line(aes(y=HumidityMedium), color = "red",type ="o")+
  geom_line(aes(y=HumidityHigh), color = "green", type= "o")+
  labs(title="Yearly Humidity Clusters ",x = "Years" , y= " Humidity")+
  geom_point(aes(y = HumidityLow), color = "blue", shape = 16) +
  geom_point(aes(y = HumidityMedium), color = "red", shape = 16) +
  geom_point(aes(y = HumidityHigh), color = "green", shape = 16, show.legend = TRUE)+
  scale_color_manual(values = c("low" = "blue", "medium" = "red", "high" = "green"),
                     labels = c("low", "medium", "high"),
                     name = "Legende")
```

# 4. Clustering 3 times 5 Years

```{r}
#print(kmeans.2006$cluster)
#print(kmeans.2006$centers)

#print(kmeans.2007$cluster)
#print(kmeans.2007$centers)

#print(kmeans.2008$cluster)
#print(kmeans.2008$centers)

#print(kmeans.2009$cluster)
#print(kmeans.2009$centers)
```

## 4.1 creating an new List with 3 times 5 years Clusterd together

```{r}
huss.2006<- as.data.frame(T.huss.mon.ordered[,5:16])
huss.2007<- as.data.frame(T.huss.mon.ordered[,17:28])
huss.2008<- as.data.frame(T.huss.mon.ordered[,29:40])
huss.2009<- as.data.frame(T.huss.mon.ordered[,41:52])
huss.2010<- as.data.frame(T.huss.mon.ordered[,53:64])

huss.2006.2010<-as.data.frame(T.huss.mon.ordered[,5:64])

huss.2011<- as.data.frame(T.huss.mon.ordered[,65:76])
huss.2012<- as.data.frame(T.huss.mon.ordered[,77:88])
huss.2013<- as.data.frame(T.huss.mon.ordered[,89:100])
huss.2014<- as.data.frame(T.huss.mon.ordered[,101:112])
huss.2015<- as.data.frame(T.huss.mon.ordered[,113:124])

huss.2011.2015<-as.data.frame(T.huss.mon.ordered[,65:124])

huss.2016<- as.data.frame(T.huss.mon.ordered[,125:136])
huss.2017<- as.data.frame(T.huss.mon.ordered[,137:148])
huss.2018<- as.data.frame(T.huss.mon.ordered[,149:150])
huss.2019<- as.data.frame(T.huss.mon.ordered[,151:162])
huss.2020<- as.data.frame(T.huss.mon.ordered[,163:174])

huss.2016.2020<-as.data.frame(T.huss.mon.ordered[,125:174])

huss3.2006 <- huss.2006.2010
huss3.2011 <- huss.2011.2015
huss3.2016 <- huss.2016.2020


huss3.<- list(
  huss3.2006 = huss.2006.2010,
  huss3.2011 = huss.2011.2015,
  huss3.2016 = huss.2016.2020
)
```

```{r}
for (year in 2006:2020){

k <-3

huss3.year <- get(paste0("huss3.",year))
huss3.year.mean<-as.data.frame(apply(huss3.year,1, mean))
huss3.year.mean ['numbers'] = c(1:77)

kmeansN.years <-kmeans(huss3.year.mean$`apply(huss.year3, 1, mean)`, centers = k )



## Zuordnen aller original values  of cluster 1,2,3 their corresponding center 

centers3.years <- as.vector(kmeans3.years$centers)
centers3.clusters.years <- data.frame(Value = kmeans3.years$cluster, Center = NA)

for (i in 1:length(kmeans3.years$cluster)) {
  centers3.clusters.years$Center[i] <- centers3.years[kmeans3.years$cluster[i]]
}


centers3.clusters.years['numbers'] = c(1:77)

centers3.clusters.years.geography <- merge(centers3.clusters.years, geography, by = "numbers")
 
## ggplot
print(ggplot() +
  geom_sf(data = centers3.clusters.years.geography ,aes(geometry = geometry, fill= Center)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = paste("kmeans K = 3  humidity of Provinces",year)) +
  #scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")+
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  guides(fill = guide_legend(title = "Legende", override.aes = list(shape = 22))))
}

```

```{r}

#PLOT 2006-2010------------------------------------------------------------------------------------
  
## Clustering  
k <-3

huss.2006.2010
huss.2006.2010.mean<-as.data.frame(apply(huss.2006.2010,1, mean))
huss.2006.2010.mean ['numbers'] = c(1:77)
kmeans.2006.2010 <-kmeans(huss.2006.2010.mean$`apply(huss.2006.2010, 1, mean)`, centers = k )

print(kmeans.2006.2010$cluster)
print(kmeans.2006.2010$centers)

kmeans.2006.2010.cluster = as.data.frame(kmeans.2006.2010$cluster)
kmeans.2006.2010.cluster['numbers'] = c(1:77)
kmeans.2006.2010.clusters.geography = merge(kmeans.2006.2010.cluster, geography, by= "numbers")
colnames(kmeans.2006.2010.clusters.geography)[colnames(kmeans.2006.2010.clusters.geography) == "kmeans.2006.2010$cluster"] <- "Clusters2006.2010"
kmeans.2006.2010.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2006.2010.clusters.geography ,aes(geometry = geometry, fill= Clusters2006.2010)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces mean 2006-2010") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2011-2015------------------------------------------------------------------------------------
  
## Clustering  
k <-3

huss.2011.2015
huss.2011.2015.mean<-as.data.frame(apply(huss.2011.2015,1, mean))
huss.2011.2015.mean ['numbers'] = c(1:77)
kmeans.2011.2015 <-kmeans(huss.2011.2015.mean$`apply(huss.2011.2015, 1, mean)`, centers = k )

print(kmeans.2011.2015$cluster)
print(kmeans.2011.2015$centers)

kmeans.2011.2015.cluster = as.data.frame(kmeans.2011.2015$cluster)
kmeans.2011.2015.cluster['numbers'] = c(1:77)
kmeans.2011.2015.clusters.geography = merge(kmeans.2011.2015.cluster, geography, by= "numbers")
colnames(kmeans.2011.2015.clusters.geography)[colnames(kmeans.2011.2015.clusters.geography) == "kmeans.2011.2015$cluster"] <- "Clusters2011.2015"
kmeans.2011.2015.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2011.2015.clusters.geography ,aes(geometry = geometry, fill= Clusters2011.2015)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces mean 2011-2015") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2016-2020------------------------------------------------------------------------------------
  
## Clustering  
k <-3

huss.2016.2020
huss.2016.2020.mean<-as.data.frame(apply(huss.2016.2020,1, mean))
huss.2016.2020.mean ['numbers'] = c(1:77)
kmeans.2016.2020 <-kmeans(huss.2016.2020.mean$`apply(huss.2016.2020, 1, mean)`, centers = k )

print(kmeans.2016.2020$cluster)
print(kmeans.2016.2020$centers)

kmeans.2016.2020.cluster = as.data.frame(kmeans.2016.2020$cluster)
kmeans.2016.2020.cluster['numbers'] = c(1:77)
kmeans.2016.2020.clusters.geography = merge(kmeans.2016.2020.cluster, geography, by= "numbers")
colnames(kmeans.2016.2020.clusters.geography)[colnames(kmeans.2016.2020.clusters.geography) == "kmeans.2016.2020$cluster"] <- "Clusters2016.2020"
kmeans.2016.2020.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2016.2020.clusters.geography ,aes(geometry = geometry, fill= Clusters2016.2020)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces mean 2016-2020") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
```

```{r}

print(kmeans.2006.2010$cluster)
print(kmeans.2006.2010$centers)

print(kmeans.2011.2015$cluster)
print(kmeans.2011.2015$centers)

print(kmeans.2016.2020$cluster)
print(kmeans.2016.2020$centers)
```

## Attached Code wich was used for presious Analysis

```{r}
#PLOT 2006--------------------------------------------------------------------------------

##Clustering

k <-3

huss.2006
huss.2006.mean<-as.data.frame(apply(huss.2006,1, mean))
huss.2006.mean ['numbers'] = c(1:77)
kmeans.2006 <-kmeans(huss.2006.mean$`apply(huss.2006, 1, mean)`, centers = k )

print(kmeans.2006$cluster)
print(kmeans.2006$centers)

kmeans.2006.cluster = as.data.frame(kmeans.2006$cluster)
kmeans.2006.cluster['numbers'] = c(1:77)
kmeans.2006.clusters.geography = merge(kmeans.2006.cluster, geography, by= "numbers")
colnames(kmeans.2006.clusters.geography)[colnames(kmeans.2006.clusters.geography) == "kmeans.2006$cluster"] <- "Clusters2006"
kmeans.2006.cluster['numbers'] = c(1:77)

print(kmeans.2006$cluster)
print(kmeans.2006$centers)


ggplot() +
  geom_sf(data = kmeans.2006.clusters.geography ,aes(geometry = geometry, fill= Clusters2006)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces 2006") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2007--------------------------------------------------------------------------------
## Clustering
k <-3

huss.2007
huss.2007.mean<-as.data.frame(apply(huss.2007,1, mean))
huss.2007.mean ['numbers'] = c(1:77)
kmeans.2007 <-kmeans(huss.2007.mean$`apply(huss.2007, 1, mean)`, centers = k )


## Zuordnen aller original values  of cluster 1,2,3 their corresponding center 

centers.2006 <- as.vector(kmeans.2006$centers)
centers.clusters.2006 <- data.frame(Value = kmeans.2006$cluster, Center = NA)

for (i in 1:length(kmeans.2006$cluster)) {
  centers.clusters.2006$Center[i] <- centers.2006[kmeans.2006$cluster[i]]
}

centers.clusters.2006['numbers'] = c(1:77)
centers.clusters.2006.geography = merge(centers.clusters.2006, geography, by= "numbers")


## ggplot

ggplot() +
  geom_sf(data = centers.clusters.2006.geography ,aes(geometry = geometry, fill= Center)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces 2006") +
  #scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")+
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  guides(fill = guide_legend(title = "Legende", override.aes = list(shape = 22)))



kmeans.2007.cluster = as.data.frame(kmeans.2007$cluster)
kmeans.2007.cluster['numbers'] = c(1:77)
kmeans.2007.clusters.geography = merge(kmeans.2007.cluster, geography, by= "numbers")
colnames(kmeans.2007.clusters.geography)[colnames(kmeans.2007.clusters.geography) == "kmeans.2007$cluster"] <- "Clusters2007"
kmeans.2007.cluster['numbers'] = c(1:77)

#View(kmeans.2006.clusters.geography)


ggplot() +
  geom_sf(data = kmeans.2007.clusters.geography ,aes(geometry = geometry, fill= Clusters2007)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2007") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2008-----------------------------------------------------------------------------------
## Clustering
k <-3

huss.2008
huss.2008.mean<-as.data.frame(apply(huss.2008,1, mean))
huss.2008.mean ['numbers'] = c(1:77)
kmeans.2008 <-kmeans(huss.2008.mean$`apply(huss.2008, 1, mean)`, centers = k )

print(kmeans.2008$cluster)
print(kmeans.2008$centers)

kmeans.2008.cluster = as.data.frame(kmeans.2008$cluster)
kmeans.2008.cluster['numbers'] = c(1:77)
kmeans.2008.clusters.geography = merge(kmeans.2008.cluster, geography, by= "numbers")
colnames(kmeans.2008.clusters.geography)[colnames(kmeans.2008.clusters.geography) == "kmeans.2008$cluster"] <- "Clusters2008"
kmeans.2008.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2008.clusters.geography ,aes(geometry = geometry, fill= Clusters2008)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2008") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2009-----------------------------------------------------------------------------------
## Clustering  
k <-3

huss.2009
huss.2009.mean<-as.data.frame(apply(huss.2009,1, mean))
huss.2009.mean ['numbers'] = c(1:77)
kmeans.2009 <-kmeans(huss.2009.mean$`apply(huss.2009, 1, mean)`, centers = k )

print(kmeans.2009$cluster)
print(kmeans.2009$centers)

kmeans.2009.cluster = as.data.frame(kmeans.2009$cluster)
kmeans.2009.cluster['numbers'] = c(1:77)
kmeans.2009.clusters.geography = merge(kmeans.2009.cluster, geography, by= "numbers")
colnames(kmeans.2009.clusters.geography)[colnames(kmeans.2009.clusters.geography) == "kmeans.2009$cluster"] <- "Clusters2009"
kmeans.2009.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2009.clusters.geography ,aes(geometry = geometry, fill= Clusters2009)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2009") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")


#PLOT 2010-------------------------------------------------------------------------------------
  ## Clustering  
k <-3

huss.2010
huss.2010.mean<-as.data.frame(apply(huss.2010,1, mean))
huss.2010.mean ['numbers'] = c(1:77)
kmeans.2010 <-kmeans(huss.2010.mean$`apply(huss.2010, 1, mean)`, centers = k )

print(kmeans.2010$cluster)
print(kmeans.2010$centers)

kmeans.2010.cluster = as.data.frame(kmeans.2010$cluster)
kmeans.2010.cluster['numbers'] = c(1:77)
kmeans.2010.clusters.geography = merge(kmeans.2010.cluster, geography, by= "numbers")
colnames(kmeans.2010.clusters.geography)[colnames(kmeans.2010.clusters.geography) == "kmeans.2010$cluster"] <- "Clusters2010"
kmeans.2010.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2010.clusters.geography ,aes(geometry = geometry, fill= Clusters2010)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2010") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2011------------------------------------------------------------------------------------
  
## Clustering  
k <-3

huss.2011
huss.2011.mean<-as.data.frame(apply(huss.2011,1, mean))
huss.2011.mean ['numbers'] = c(1:77)
kmeans.2011 <-kmeans(huss.2011.mean$`apply(huss.2011, 1, mean)`, centers = k )

print(kmeans.2011$cluster)
print(kmeans.2011$centers)

kmeans.2011.cluster = as.data.frame(kmeans.2011$cluster)
kmeans.2011.cluster['numbers'] = c(1:77)
kmeans.2011.clusters.geography = merge(kmeans.2011.cluster, geography, by= "numbers")
colnames(kmeans.2011.clusters.geography)[colnames(kmeans.2011.clusters.geography) == "kmeans.2011$cluster"] <- "Clusters2011"
kmeans.2011.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2011.clusters.geography ,aes(geometry = geometry, fill= Clusters2011)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2011") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
#PLOT 2012------------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2012
huss.2012.mean<-as.data.frame(apply(huss.2012,1, mean))
huss.2012.mean ['numbers'] = c(1:77)
kmeans.2012 <-kmeans(huss.2012.mean$`apply(huss.2012, 1, mean)`, centers = k )

print(kmeans.2012$cluster)
print(kmeans.2012$centers)

kmeans.2012.cluster = as.data.frame(kmeans.2012$cluster)
kmeans.2012.cluster['numbers'] = c(1:77)
kmeans.2012.clusters.geography = merge(kmeans.2012.cluster, geography, by= "numbers")
colnames(kmeans.2012.clusters.geography)[colnames(kmeans.2012.clusters.geography) == "kmeans.2012$cluster"] <- "Clusters2012"
kmeans.2012.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2012.clusters.geography ,aes(geometry = geometry, fill= Clusters2012)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2012") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2013-----------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2013
huss.2013.mean<-as.data.frame(apply(huss.2013,1, mean))
huss.2013.mean ['numbers'] = c(1:77)
kmeans.2013 <-kmeans(huss.2013.mean$`apply(huss.2013, 1, mean)`, centers = k )

print(kmeans.2013$cluster)
print(kmeans.2013$centers)

kmeans.2013.cluster = as.data.frame(kmeans.2013$cluster)
kmeans.2013.cluster['numbers'] = c(1:77)
kmeans.2013.clusters.geography = merge(kmeans.2013.cluster, geography, by= "numbers")
colnames(kmeans.2013.clusters.geography)[colnames(kmeans.2013.clusters.geography) == "kmeans.2013$cluster"] <- "Clusters2013"
kmeans.2013.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2013.clusters.geography ,aes(geometry = geometry, fill= Clusters2013)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2013") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2014--------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2014
huss.2014.mean<-as.data.frame(apply(huss.2014,1, mean))
huss.2014.mean ['numbers'] = c(1:77)
kmeans.2014 <-kmeans(huss.2014.mean$`apply(huss.2014, 1, mean)`, centers = k )

print(kmeans.2014$cluster)
print(kmeans.2014$centers)

kmeans.2014.cluster = as.data.frame(kmeans.2014$cluster)
kmeans.2014.cluster['numbers'] = c(1:77)
kmeans.2014.clusters.geography = merge(kmeans.2014.cluster, geography, by= "numbers")
colnames(kmeans.2014.clusters.geography)[colnames(kmeans.2014.clusters.geography) == "kmeans.2014$cluster"] <- "Clusters2014"
kmeans.2014.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2014.clusters.geography ,aes(geometry = geometry, fill= Clusters2014)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2014") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")


#PLOT 2015---------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2015
huss.2015.mean<-as.data.frame(apply(huss.2015,1, mean))
huss.2015.mean ['numbers'] = c(1:77)
kmeans.2015 <-kmeans(huss.2015.mean$`apply(huss.2015, 1, mean)`, centers = k )

print(kmeans.2015$cluster)
print(kmeans.2015$centers)

kmeans.2015.cluster = as.data.frame(kmeans.2015$cluster)
kmeans.2015.cluster['numbers'] = c(1:77)
kmeans.2015.clusters.geography = merge(kmeans.2015.cluster, geography, by= "numbers")
colnames(kmeans.2015.clusters.geography)[colnames(kmeans.2015.clusters.geography) == "kmeans.2015$cluster"] <- "Clusters2015"
kmeans.2015.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2015.clusters.geography ,aes(geometry = geometry, fill= Clusters2015)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2015") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
#PLOT 2016--------------------------------------------------------------------------------
  ## Clustering  
k <-3

huss.2016
huss.2016.mean<-as.data.frame(apply(huss.2016,1, mean))
huss.2016.mean ['numbers'] = c(1:77)
kmeans.2016 <-kmeans(huss.2016.mean$`apply(huss.2016, 1, mean)`, centers = k )

print(kmeans.2016$cluster)
print(kmeans.2016$centers)

kmeans.2016.cluster = as.data.frame(kmeans.2016$cluster)
kmeans.2016.cluster['numbers'] = c(1:77)
kmeans.2016.clusters.geography = merge(kmeans.2016.cluster, geography, by= "numbers")
colnames(kmeans.2016.clusters.geography)[colnames(kmeans.2016.clusters.geography) == "kmeans.2016$cluster"] <- "Clusters2016"
kmeans.2016.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2016.clusters.geography ,aes(geometry = geometry, fill= Clusters2016)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2016") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2017---------------------------------------------------------------------------------
  ## Clustering  
k <-3

huss.2017
huss.2017.mean<-as.data.frame(apply(huss.2017,1, mean))
huss.2017.mean ['numbers'] = c(1:77)
kmeans.2017 <-kmeans(huss.2017.mean$`apply(huss.2017, 1, mean)`, centers = k )

print(kmeans.2017$cluster)
print(kmeans.2017$centers)

kmeans.2017.cluster = as.data.frame(kmeans.2017$cluster)
kmeans.2017.cluster['numbers'] = c(1:77)
kmeans.2017.clusters.geography = merge(kmeans.2017.cluster, geography, by= "numbers")
colnames(kmeans.2017.clusters.geography)[colnames(kmeans.2017.clusters.geography) == "kmeans.2017$cluster"] <- "Clusters2017"
kmeans.2017.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2017.clusters.geography ,aes(geometry = geometry, fill= Clusters2017)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2017") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
#PLOT 2018---------------------------------------------------------------------------------
  ## Clustering  
k <-3

huss.2018
huss.2018.mean<-as.data.frame(apply(huss.2018,1, mean))
huss.2018.mean ['numbers'] = c(1:77)
kmeans.2018 <-kmeans(huss.2018.mean$`apply(huss.2018, 1, mean)`, centers = k )

print(kmeans.2018$cluster)
print(kmeans.2018$centers)

kmeans.2018.cluster = as.data.frame(kmeans.2018$cluster)
kmeans.2018.cluster['numbers'] = c(1:77)
kmeans.2018.clusters.geography = merge(kmeans.2018.cluster, geography, by= "numbers")
colnames(kmeans.2018.clusters.geography)[colnames(kmeans.2018.clusters.geography) == "kmeans.2018$cluster"] <- "Clusters2018"
kmeans.2018.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2018.clusters.geography ,aes(geometry = geometry, fill= Clusters2018)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2018") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
#PLOT 2019---------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2019
huss.2019.mean<-as.data.frame(apply(huss.2019,1, mean))
huss.2019.mean ['numbers'] = c(1:77)
kmeans.2019 <-kmeans(huss.2019.mean$`apply(huss.2019, 1, mean)`, centers = k )

print(kmeans.2019$cluster)
print(kmeans.2019$centers)

kmeans.2019.cluster = as.data.frame(kmeans.2019$cluster)
kmeans.2019.cluster['numbers'] = c(1:77)
kmeans.2019.clusters.geography = merge(kmeans.2019.cluster, geography, by= "numbers")
colnames(kmeans.2019.clusters.geography)[colnames(kmeans.2019.clusters.geography) == "kmeans.2019$cluster"] <- "Clusters2019"
kmeans.2019.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2019.clusters.geography ,aes(geometry = geometry, fill= Clusters2019)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2019") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")

#PLOT 2020---------------------------------------------------------------------------------

  ## Clustering  
k <-3

huss.2020
huss.2020.mean<-as.data.frame(apply(huss.2020,1, mean))
huss.2020.mean ['numbers'] = c(1:77)
kmeans.2020 <-kmeans(huss.2020.mean$`apply(huss.2020, 1, mean)`, centers = k )

print(kmeans.2020$cluster)
print(kmeans.2020$centers)

kmeans.2020.cluster = as.data.frame(kmeans.2020$cluster)
kmeans.2020.cluster['numbers'] = c(1:77)
kmeans.2020.clusters.geography = merge(kmeans.2020.cluster, geography, by= "numbers")
colnames(kmeans.2020.clusters.geography)[colnames(kmeans.2020.clusters.geography) == "kmeans.2020$cluster"] <- "Clusters2020"
kmeans.2020.cluster['numbers'] = c(1:77)

## Plotten

ggplot() +
  geom_sf(data = kmeans.2020.clusters.geography ,aes(geometry = geometry, fill= Clusters2020)) +
  coord_sf() +  
  theme_minimal() +
  labs(title = "kmeans K = 3  humidity of Provinces in 2020") +
  scale_fill_viridis()+
  labs(x="Longitude", y= "Latitude")
```
