---
title: "dengue_data_plotting"
output: html_document
date: "2023-06-27"
author: Lea Mrowiec
---

# 0) Preparing the Dataframes

"dengue_data_com2" is a dataframe containing the dengue cases of each month (2006 to 2020). The data is normalized according to the number of citizens in each province over each years. Therefore, the numbers show the dengue cases per 100.000 citizens.

```{r}
#loading dataframe with dengue data
library(readxl)
dengue_data_com2 = read.csv("../dengue_data/dengue_without_NAs_norm.csv")
dengue_data_com2

#loading the geo data
library(sf)
shapefilePath <- "../geo_data/gadm36_THA_shp"
layers <- st_layers(shapefilePath)
geography <- st_read(shapefilePath, layer = "gadm36_THA_1")
#geography <- st_read("/Users/leamrowiec/Desktop/Datenanalyse SS23/other data/gadm36_THA_shp")
geography= as.data.frame(geography)
geography
```

```{r}
# in 2010 the values for Bunga Khan are not set no NA in the initial dataframe. Therefore I set those values to NA. 
de.2010 = c("X2010.total", "X2010.01", "X2010.02", "X2010.03", "X2010.04", "X2010.05", "X2010.06", "X2010.07", "X2010.08", "X2010.09", "X2010.10", "X2010.11", "X2010.12")
dengue_data_com2[4, de.2010] = NA
dengue_data_com2
#combining the dataframes
geo_dengue=cbind(dengue_data_com2, geography)
geo_dengue
```

# 1) Plotting Total Dengue Cases of Each Year

Note: in 2011 a new province called "Bueng Kan" was founded. It was was part of the province "Nong Khai". Because we had no data of dengue cases for Bueng Kan before 2011 we defined them as NAs. The total dengue cases of Nong Khai and todays Bueng Kan can be seen in the map in the province Nong Khai. After 2010 both provinces have separate values.

```{r}
library(viridis)
library(ggplot2)

#2006--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2006.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Total Dengue Cases per 100.000 Citizens in 2006", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2007--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2007.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2007", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2008--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2008.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Total Dengue Cases per 100.000 Citizens in 2008", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2009--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2009.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Total Dengue Cases per 100.000 Citizens in 2009", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2010--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2010.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Total Dengue Cases per 100.000 Citizens in 2010", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2011--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2011.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2011", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2012--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2012.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2012", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2013--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2013.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue cases per 100.000 citizens in 2013", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2014--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2014.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue ases per 100.000 Citizens in 2014", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2015--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2015.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2015", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2016--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2016.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2016", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2017--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2017.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2017", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2018--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2018.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2018", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2019--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2019.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Total Dengue Cases per 100.000 Citizens in 2019", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

#2020--------------------------------------------------------------------------------------
ggplot(geo_dengue, aes(geometry=geometry)) +
  geom_sf (aes(fill = X2020.total)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Total Dengue Cases per 100.000 Citizens in 2020", fill = "cases per 100.000 citizens")+
  labs(caption = "provinces with non-available data are shown in grey")

```

# 2) Geoplots of Each Month

## 2.1) Assembling the Dataframe

```{r}
# 1) computing the mean of each month over all years
jan = geo_dengue[, c("X2006.01", "X2007.01", "X2008.01", "X2009.01", "X2010.01", "X2011.01", "X2012.01", "X2013.01", "X2014.01", "X2015.01", "X2016.01", "X2017.01", "X2018.01", "X2019.01", "X2020.01")]
feb = geo_dengue[, c("X2006.02", "X2007.02", "X2008.02", "X2009.02", "X2010.02", "X2011.02", "X2012.02", "X2013.02", "X2014.02", "X2015.02", "X2016.02", "X2017.02", "X2018.02", "X2019.02", "X2020.02")]
mar = geo_dengue[, c("X2006.03", "X2007.03", "X2008.03", "X2009.03", "X2010.03", "X2011.03", "X2012.03", "X2013.03", "X2014.03", "X2015.03", "X2016.03", "X2017.03", "X2018.03", "X2019.03", "X2020.03")]
apr = geo_dengue[, c("X2006.04", "X2007.04", "X2008.04", "X2009.04", "X2010.04", "X2011.04", "X2012.04", "X2013.04", "X2014.04", "X2015.04", "X2016.04", "X2017.04", "X2018.04", "X2019.04", "X2020.04")]
may = geo_dengue[, c("X2006.05", "X2007.05", "X2008.05", "X2009.05", "X2010.05", "X2011.05", "X2012.05", "X2013.05", "X2014.05", "X2015.05", "X2016.05", "X2017.05", "X2018.05", "X2019.05", "X2020.05")]
jun = geo_dengue[, c("X2006.06", "X2007.06", "X2008.06", "X2009.06", "X2010.06", "X2011.06", "X2012.06", "X2013.06", "X2014.06", "X2015.06", "X2016.06", "X2017.06", "X2018.06", "X2019.06", "X2020.06")]
jul = geo_dengue[, c("X2006.07", "X2007.07", "X2008.07", "X2009.07", "X2010.07", "X2011.07", "X2012.07", "X2013.07", "X2014.07", "X2015.07", "X2016.07", "X2017.07", "X2018.07", "X2019.07", "X2020.07")]
aug = geo_dengue[, c("X2006.08", "X2007.08", "X2008.08", "X2009.08", "X2010.08", "X2011.08", "X2012.08", "X2013.08", "X2014.08", "X2015.08", "X2016.08", "X2017.08", "X2018.08", "X2019.08", "X2020.08")]
sep = geo_dengue[, c("X2006.09", "X2007.09", "X2008.09", "X2009.09", "X2010.09", "X2011.09", "X2012.09", "X2013.09", "X2014.09", "X2015.09", "X2016.09", "X2017.09", "X2018.09", "X2019.09", "X2020.09")]
oct = geo_dengue[, c("X2006.10", "X2007.10", "X2008.10", "X2009.10", "X2010.10", "X2011.10", "X2012.10", "X2013.10", "X2014.10", "X2015.10", "X2016.10", "X2017.10", "X2018.10", "X2019.10", "X2020.10")]
nov = geo_dengue[, c("X2006.11", "X2007.11", "X2008.11", "X2009.11", "X2010.11", "X2011.11", "X2012.11", "X2013.11", "X2014.11", "X2015.11", "X2016.11", "X2017.11", "X2018.11", "X2019.11", "X2020.11")]
dec = geo_dengue[, c("X2006.12", "X2007.12", "X2008.12", "X2009.12", "X2010.12", "X2011.12", "X2012.12", "X2013.12", "X2014.12", "X2015.12", "X2016.12", "X2017.12", "X2018.12", "X2019.12", "X2020.12")]

## removing NA values to make sure that the mean can be calculated
jan.m = apply(jan, 1, mean, na.rm=TRUE)
feb.m = apply(feb, 1, mean, na.rm=TRUE)
mar.m = apply(mar, 1, mean, na.rm=TRUE)
apr.m = apply(apr, 1, mean, na.rm=TRUE)
may.m = apply(may, 1, mean, na.rm=TRUE)
jun.m = apply(jun, 1, mean, na.rm=TRUE)
jul.m = apply(jul, 1, mean, na.rm=TRUE)
aug.m = apply(aug, 1, mean, na.rm=TRUE)
sep.m = apply(sep, 1, mean, na.rm=TRUE)
oct.m = apply(oct, 1, mean, na.rm=TRUE)
nov.m = apply(nov, 1, mean, na.rm=TRUE)
dec.m = apply(dec, 1, mean, na.rm=TRUE)

# combining the mean values in an new dataframe 
geo_dengue_mean = cbind(geo_dengue$Reporting.areas, jan.m, feb.m, mar.m, apr.m, may.m, jun.m, jul.m, aug.m, sep.m, oct.m, nov.m, dec.m, geography)
geo_dengue_mean
```

## 2.2) Plotting Dengue Cases per 100.000 Citizens for Each Month over All Years

I plotted the mean of dengue cases of each month i each province, so: the cases for January show the mean of all cases in each province that were reported in January of the years 2006 to 2020.

```{r}
#Jan-------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = jan.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in January ", fill = "cases per 100.000 citizens")

#Feb--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = feb.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in February ", fill = "cases per 100.000 citizens")

#Mar--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = mar.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in March ", fill = "cases per 100.000 citizens")
#Apr--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = apr.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in April ", fill = "cases per 100.000 citizens")

#May--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = may.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in May ", fill = "cases per 100.000 citizens")

#Jun--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = jun.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in June ", fill = "cases per 100.000 citizens")

#Jul--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = jul.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Mean Dengue Cases per Monthper 100.000 Citizens in July ", fill = "cases per 100.000 citizens")

#Aug--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = aug.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in August ", fill = "cases per 100.000 citizens")

#Sep--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = sep.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in September ", fill = "cases per 100.000 citizens")


#Oct--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = oct.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in October ", fill = "cases per 100.000 citizens")

#Nov--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = nov.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in November ", fill = "cases per 100.000 citizens")
#Dec--------------------------------------------------------------------------------------
ggplot(geo_dengue_mean, aes(geometry=geometry)) +
  geom_sf (aes(fill = dec.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
   labs(title = "Mean Dengue Cases per Month per 100.000 Citizens in December ", fill = "cases per 100.000 citizens")

```

## 2.3) Geoplots of Mean Dengue Cases per Month During the Different Monsoon Seasons

```{r}
season_dengue_m= cbind(jan.m, feb.m, mar.m, apr.m, may.m, jun.m, jul.m, aug.m, sep.m, oct.m, nov.m, dec.m)
as.data.frame(season_dengue_m)

#pre-monsoon: march, april, may
#monsoon: june, july, august, september
#post-monsoon: october, november
pre.monsoon.d.m= apply(season_dengue_m[,c(3,4,5)], 1, mean)
monsoon.d.m= apply(season_dengue_m[,c(6,7,8,9)],1,mean)
post.monsoon.d.m= apply(season_dengue_m[,c(10,11)],1,mean)
inter.monsoon.d.m= apply(season_dengue_m[,c(12,1,2)],1,mean)
monsoon_dengue_mean_df = cbind(pre.monsoon.d.m, monsoon.d.m, post.monsoon.d.m, inter.monsoon.d.m, geography)
monsoon_dengue_mean_df

# pre monsoon----------------------------------------------------------------------------------
ggplot(monsoon_dengue_mean_df, aes(geometry=geometry)) +
  geom_sf (aes(fill = pre.monsoon.d.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Pre-Monsoon Season",fill = "mean dengue cases per 100.000 citizens")+
  labs(caption= "pre-monsoon season (March, April, May)")

# inter monsoon--------------------------------------------------------------------------------
ggplot(monsoon_dengue_mean_df, aes(geometry=geometry)) +
  geom_sf (aes(fill = monsoon.d.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Monsoon Season",fill = "mean dengue cases per 100.000 citizens")+
  labs(caption= "monsoon season (June, July, August, September)")

#post monsoon---------------------------------------------------------------------------------
ggplot(monsoon_dengue_mean_df, aes(geometry=geometry)) +
  geom_sf (aes(fill = post.monsoon.d.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Post-Monsoon Season",fill = "mean dengue cases per 100.000 citizens")+
  labs(caption= "post-monsoon season (October, November)")

#post monsoon---------------------------------------------------------------------------------
ggplot(monsoon_dengue_mean_df, aes(geometry=geometry)) +
  geom_sf (aes(fill = inter.monsoon.d.m)) +
  coord_sf() +
  theme_minimal()+
  colorspace::scale_fill_continuous_sequential(palette= "plasma")+
  labs(title = "Inter-Monsoon Season",fill = "mean dengue cases per 100.000 citizens")+
  labs(caption= "inter-monsoon season (December, January, Fabruary)")
```

#  3) Boxplot of the Total Cases of Every Year per 100.000 Citizens 

```{r}
#defining a dataframe for each year (mean of all months in one year)
d2006 = geo_dengue[,c("X2006.01", "X2006.02", "X2006.03", "X2006.04", "X2006.05", "X2006.06", "X2006.07", "X2006.08", "X2006.09","X2006.10","X2006.11","X2006.12")]
d2007 = geo_dengue[, c("X2007.01", "X2007.02", "X2007.03", "X2007.04", "X2007.05", "X2007.06", "X2007.07", "X2007.08", "X2007.09","X2007.10","X2007.11","X2007.12")]
d2008 = geo_dengue[, c("X2008.01", "X2008.02", "X2008.03", "X2008.04", "X2008.05", "X2008.06", "X2008.07", "X2008.08", "X2008.09","X2008.10","X2008.11","X2008.12")]
d2009 = geo_dengue[, c("X2009.01", "X2009.02", "X2009.03", "X2009.04", "X2009.05", "X2009.06", "X2009.07", "X2009.08", "X2009.09","X2009.10","X2009.11","X2009.12")]
d2010 = geo_dengue[, c("X2010.01", "X2010.02", "X2010.03", "X2010.04", "X2010.05", "X2010.06", "X2010.07", "X2010.08", "X2010.09","X2010.10","X2010.11","X2010.12")]
d2011 = geo_dengue[, c("X2011.01", "X2011.02", "X2011.03", "X2011.04", "X2011.05", "X2011.06", "X2011.07", "X2011.08", "X2011.09","X2011.10","X2011.11","X2011.12")]
d2012 = geo_dengue[, c("X2012.01", "X2012.02", "X2012.03", "X2012.04", "X2012.05", "X2012.06", "X2012.07", "X2012.08", "X2012.09","X2012.10","X2012.11","X2012.12")]
d2013 = geo_dengue[, c("X2013.01", "X2013.02", "X2013.03", "X2013.04", "X2013.05", "X2013.06", "X2013.07", "X2013.08", "X2013.09","X2013.10","X2013.11","X2013.12")]
d2014 = geo_dengue[, c("X2014.01", "X2014.02", "X2014.03", "X2014.04", "X2014.05", "X2014.06", "X2014.07", "X2014.08", "X2014.09","X2014.10","X2014.11","X2014.12")]
d2015 = geo_dengue[, c("X2015.01", "X2015.02", "X2015.03", "X2015.04", "X2015.05", "X2015.06", "X2015.07", "X2015.08", "X2015.09","X2015.10","X2015.11","X2015.12")]
d2016 = geo_dengue[, c("X2016.01", "X2016.02", "X2016.03", "X2016.04", "X2016.05", "X2016.06", "X2016.07", "X2016.08", "X2016.09","X2016.10","X2016.11","X2016.12")]
d2017 = geo_dengue[, c("X2017.01", "X2017.02", "X2017.03", "X2017.04", "X2017.05", "X2017.06", "X2017.07", "X2017.08", "X2017.09","X2017.10","X2017.11","X2017.12")]
d2018 = geo_dengue[, c("X2018.01", "X2018.02", "X2018.03", "X2018.04", "X2018.05", "X2018.06", "X2018.07", "X2018.08", "X2018.09","X2018.10","X2018.11","X2018.12")]
d2019 = geo_dengue[, c("X2019.01", "X2019.02", "X2019.03", "X2019.04", "X2019.05", "X2019.06", "X2019.07", "X2019.08", "X2019.09","X2019.10","X2019.11","X2019.12")]
d2020 = geo_dengue[, c("X2020.01", "X2020.02", "X2020.03", "X2020.04", "X2020.05", "X2020.06", "X2020.07", "X2020.08", "X2020.09","X2020.10","X2020.11","X2020.12")]

#computing mean of each year
m.d2006 = apply(d2006, 1, mean, na.rm=TRUE)
m.d2007 = apply(d2007, 1, mean,na.rm=TRUE)
m.d2008 = apply(d2008, 1, mean,na.rm=TRUE)
m.d2009 = apply(d2009, 1, mean,na.rm=TRUE)
m.d2010 = apply(d2010, 1, mean,na.rm=TRUE)
m.d2011 = apply(d2011, 1, mean,na.rm=TRUE)
m.d2012 = apply(d2012, 1, mean,na.rm=TRUE)
m.d2013 = apply(d2013, 1, mean,na.rm=TRUE)
m.d2014 = apply(d2014, 1, mean,na.rm=TRUE)
m.d2015 = apply(d2015, 1, mean,na.rm=TRUE)
m.d2016 = apply(d2016, 1, mean,na.rm=TRUE)
m.d2017 = apply(d2017, 1, mean,na.rm=TRUE)
m.d2018 = apply(d2018, 1, mean,na.rm=TRUE)
m.d2019 = apply(d2019, 1, mean,na.rm=TRUE)
m.d2020 = apply(d2020, 1, mean,na.rm=TRUE)


mean_dengue_yearly = cbind(m.d2006, m.d2007, m.d2008, m.d2009, m.d2010, m.d2011, m.d2012, m.d2013, m.d2014, m.d2015, m.d2016, m.d2017, m.d2018, m.d2019, m.d2020)

mean_dengue_yearly
mean_dengue_yearly <- replace(mean_dengue_yearly, mean_dengue_yearly == "NaN", NA)
mean_dengue_yearly = as.data.frame(mean_dengue_yearly)
mean_dengue_yearly

#starting to plot

library(ggplot2)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(cowplot)

#plot
plot.m.2006=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2006))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2006, y=" ")+
        ylim(50, 150)
plot.m.2007=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2007))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2007, y=" ")+
        ylim(50, 150)
plot.m.2008=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$s.d2008))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2008, y=" ")+
        ylim(50, 150)
plot.m.2009=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2009))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2009, y=" ")+
        ylim(50, 150)
plot.m.2010=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2010))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2010, y=" ")+
        ylim(50, 150)
plot.m.2011=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2011))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2011, y=" ")+
        ylim(50, 150)
plot.m.2012=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2012))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2012, y=" ")+
        ylim(50, 150)
plot.m.2013=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2013))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2013, y=" ")+
        ylim(50, 150)
plot.m.2014=ggplot(mean_dengue_yearly, aes(y=sum_dengue_yearly$m.d2014))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2014, y=" ")+
        ylim(50, 150)
plot.m.2015=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2015))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2015, y=" ")+
        ylim(50, 150)
plot.m.2016=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2016))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2016, y=" ")+
        ylim(50, 150)
plot.m.2017=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2017))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2017, y=" ")+
        ylim(50, 150)
plot.m.2018=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2018))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2018, y=" ")+
        ylim(50, 150)
plot.m.2019=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2019))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2019, y=" ")+
        ylim(50, 150)
plot.m.2020=ggplot(mean_dengue_yearly, aes(y=mean_dengue_yearly$m.d2020))+
        stat_boxplot(geom="errorbar")+
        scale_x_discrete()+
        geom_boxplot(outlier.color= "gray", outlier.size=2)+
        labs(title=2020, y=" ")+
        ylim(50, 150)

#combining the plots into one big plot makes problems -> arrangement in word dokument   
#plot_grid(plot.m.2006, plot.m.2007, plot.m.2008, plot.m.2009, plot.m.2010, plot.m.2011, plot.m.2012, plot.m.2013,  nrow = 1, align = "h")
#plot_grid(plot.m.2014, plot.m.2015, plot.m.2016, plot.m.2017, plot.m.2018, plot.m.2019, plot.m.2020, nrow=1, align="h")
#sadly plot does not work

#overview of boxplot
boxplot(mean_dengue_yearly, main="Mean Dengue Cases (2006-2020)", xlab= "year", ylab="total dengue cases")



```

# 4) Context of Dengue Outbreaks

According to the boxplots, the highest mean dengue cases were reported in 2013. to have a look at the surrounding conditions (ex. previous outbreaks) the total reported dengue cases of the years 2012, 2013 and 2014 were plotted in the course of a year.

```{r}
#loading necessary library
library(ggplot2)
library(dplyr)
#computing the mean of each month in 2013
m_jan= mean(dengue_data_com2[,"X2013.01"],na.rm=TRUE)
m_feb= mean(dengue_data_com2[,"X2013.02"],na.rm=TRUE)
m_mar= mean(dengue_data_com2[,"X2013.03"],na.rm=TRUE)
m_apr= mean(dengue_data_com2[,"X2013.04"],na.rm=TRUE)
m_may= mean(dengue_data_com2[,"X2013.05"],na.rm=TRUE)
m_jun= mean(dengue_data_com2[,"X2013.06"],na.rm=TRUE)
m_jul= mean(dengue_data_com2[,"X2013.07"],na.rm=TRUE)
m_aug= mean(dengue_data_com2[,"X2013.08"],na.rm=TRUE)
m_sep= mean(dengue_data_com2[,"X2013.09"],na.rm=TRUE)
m_oct= mean(dengue_data_com2[,"X2013.10"],na.rm=TRUE)
m_nov= mean(dengue_data_com2[,"X2013.11"],na.rm=TRUE)
m_dec= mean(dengue_data_com2[,"X2013.12"],na.rm=TRUE)

mean_2013 = apply(dengue_data_com2[,c("X2013.01","X2013.02","X2013.03", "X2013.04","X2013.05","X2013.06","X2013.07","X2013.08","X2013.09","X2013.10","X2013.11","X2013.12")], 2,function(x){mean(x, na.rm = TRUE)} )
mean_2014 = apply(dengue_data_com2[,c("X2014.01","X2014.02","X2014.03", "X2014.04","X2014.05","X2014.06","X2014.07","X2014.08","X2014.09","X2014.10","X2014.11","X2014.12")], 2,function(x){mean(x, na.rm = TRUE)} )
mean_2012 = apply(dengue_data_com2[,c("X2012.01","X2012.02","X2012.03", "X2012.04","X2012.05","X2012.06","X2012.07","X2012.08","X2012.09","X2012.10","X2012.11","X2012.12")], 2,function(x){mean(x, na.rm = TRUE)} )
as.data.frame(mean_2013)
as.data.frame(mean_2014)
as.data.frame(mean_2012)
mean_2012.2013.2014 = cbind(mean_2012, mean_2013, mean_2014)
rownames(mean_2012.2013.2014) = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
mean_2012.2013.2014_df = as.data.frame(mean_2012.2013.2014)
mean_2012.2013.2014_df


# combining data in tibbles
df3 = as_tibble(mean_2012.2013.2014_df)
df3
#na.omit(df3)

#creating the plots
df3%>%
  ggplot(aes(x=seq(1,12,1), y=mean_2012, color ="2012"), linetype=1)+
  geom_line(aes(y=mean_2012), linetype=1)+
  geom_point(aes(y=mean_2012))+
  geom_line(y=mean_2013, color="violet")+ 
  geom_point(aes(y=mean_2013,color="2013"), linetype=1)+
  geom_line(y=mean_2014, color="orange")+
  geom_point(aes(y=mean_2014,color ="2014"), linetype=1)+
  ylim(0,100)+
  ylab("mean monthly dengue cases per 100.000 citizens")+
  xlab("month")+
  ggtitle("Mean Monthly Dengue Cases per 100.000 Citizens (2012, 2013, 2014)")+
  scale_color_manual(values = c("pink", "violet", "orange"))+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```

# 5) Plotting Dengue Cases in Exemplary Provinces

## 5.1) Monthly Plot

To understand the connection between dengue cases and different kinds of areas in Thailand exemplary provinces with different characteristics (position in the country, height of land) were chosen:

-   Maha-Sarakham (north-east), 100-200m above sea level

-   Chiang Mai (north to north-west), 1000-1500m above sea level

-   Kalasin (central), 0-100m above sea level

-   Chumphon

-   Songkhla (south), 100-200m above sea level

    *Diercke Weltatlas* (1st ed.). (2008). westermann

```{r}
#install.packages("lubridate")
library(ggplot2)
library(dplyr)
library(lubridate)

# dataframe combining yearly mean of each province 
month= c("Jan", "Feb", "Mar", "Apr","May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
dengue_mean = as.data.frame(rbind(jan.m, feb.m, mar.m, apr.m, may.m, jun.m, jul.m, aug.m, sep.m, oct.m, nov.m, dec.m))

dengue_mean = cbind(month, dengue_mean)                
colnames(dengue_mean)= c("Month",geography[,"NAME_1"])
#huss.means.df = as.data.frame(huss.means)
#adding column with the names of each month

dengue_mean

#plotting
dengue_mean%>%
  ggplot(aes(x=seq(1,12,1), y= dengue_mean[,16], color ="Maha Sarakham"), linetype=1)+
  geom_line(aes(y= dengue_mean[,16], color="Maha Sarakam"), linetype=1)+
  geom_line(aes(y=dengue_mean[,12], color="Chiang Mai"))+ 
  geom_line(aes(y=dengue_mean[,62], color="Saraburi"))+
  geom_line(aes(y=dengue_mean[,15], color="Chumphon"))+
  geom_line(aes(y=dengue_mean[,66], color="Songkhla"))+
  ylim(00,70)+
  ylab(" reported dengue cases per 100.000 citizens")+
  xlab("month")+
  ggtitle("Mean Dengue Cases per Month in Different Provinces")+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```

## 5.2) Monthly Plot of Provinces with Many Dengue Cases

The geoplots above show that some provinces are highly effected by dengue. Some of them are plotted here.

```{r}
dengue_mean
dengue_mean%>%
  ggplot(aes(x=seq(1,12,1), y= dengue_mean[,69], color ="Surat Thani"), linetype=1)+
  geom_line(aes(y= dengue_mean[,69], color="Surat Thani"), linetype=1)+
  geom_line(aes(y=dengue_mean[,12], color="Chiang Mai"))+ 
  geom_line(aes(y=dengue_mean[,33], color="Nakhon Si Thammarat"))+
  geom_line(aes(y=dengue_mean[,52], color="Prachuap Khiri Khan"))+
  geom_line(aes(y=dengue_mean[,66], color="Songkhla"))+
  ylim(0,60)+
  ylab(" reported dengue cases per 100.000 citizens")+
  xlab("month")+
  ggtitle("Monthly Reported Dengue Cases per 100.000 Citizens in Different Provinces")+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

```

## 5.3) Provinves in the Gulf of Thailand

Due to the thin lad area at the Gulf of Thailand, special weather conditions might appear. Therefore, it is worth having a closer look at the provinces in this area:

```{r}
dengue_mean
dengue_mean%>%
  ggplot(aes(x=seq(1,12,1), y= dengue_mean[,69], color ="Surat Thani"), linetype=1)+
  geom_line(aes(y= dengue_mean[,69], color="Surat Thani"), linetype=1)+
  geom_line(aes(y=dengue_mean[,33], color="Nakhon Si Thammarat"))+
  geom_line(aes(y=dengue_mean[,15], color="Chumphon"))+
  geom_line(aes(y=dengue_mean[,66], color="Songkhla"))+
  ylim(0,60)+
  ylab(" reported dengue cases per 100.000 citizens")+
  xlab("month")+
  ggtitle("Monthly Reported Dengue Cases per 100.000 Citizens in Different Provinces")+
  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9,10,11,12),labels=c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

## 5.4) Seasonal Plot of Dengue Cases in All Provinces

```{r}

# computing mean of all provinces
monsoon_dengue_nogeo_m = rbind(pre.monsoon.d.m, monsoon.d.m, post.monsoon.d.m, inter.monsoon.d.m)
monsoon_tot_dengue_m = apply(monsoon_dengue_nogeo_m, 1, mean)
monsoon_tot_dengue_m = as.data.frame(monsoon_tot_dengue_m)
colnames(monsoon_tot_dengue_m) = "cases per 100.000"
monsoon_tot_dengue_m


#plotting the meanof each season
monsoon_tot_dengue_m %>%
  ggplot(aes(x = seq(1, 4,  1), y = `total dengue cases per 100.000 citizens`, color = "dengue cases per 100.000")) +
  geom_line(aes(y = `cases per 100.000`)) +
  geom_point(aes(y = `cases per 100.000`))+
  ylab("reported dengue cases per 100.000 citizens per month") +
  xlab("season") +
  ggtitle("Monthly Dengue Cases in the Monsoon Seasons in the Whole Country")+
  labs(caption= "1: pre-monsoon, 2: monsoon, 3: post-monsoon, 4: inter-monsoon")

#computing general mean of all provinces and years
mean.gen = apply(monsoon_tot_dengue_m, 2, mean)
mean.gen
```
