---
title: "Correlation_testing"
author: "Paula"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The humidity and dengue case data is now sorted into their respective monsoon seasons and stacked atop of each other. Let's load it again and do some basic correlation tests 
First load the packages we need and the data 
```{r}
dengue_norm <- read.csv("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\dengue_data\\dengue_norm.csv")
T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
stacked_hum_den <- read.csv("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\Correlation_hypothesis\\mon_dengue_hum.csv")
stacked_hum_den
```
First let's look at the correlation between the respective seasons, so between the intermonsoon humdidity and the intermonsoon cases from the years 2006-2020, as well as all other seasons 
```{r}
#first non-lagged correlation 
##inter-monsoon season
cor(stacked_hum_den$intermon, stacked_hum_den$intermonH, use = "pairwise.complete.obs")
##pre-monsoon season
cor(stacked_hum_den$premon, stacked_hum_den$premonH, use = "pairwise.complete.obs")
##monsoon season 
cor(stacked_hum_den$mon, stacked_hum_den$monH, use = "pairwise.complete.obs")
##post-monsoon season
cor(stacked_hum_den$postmon, stacked_hum_den$postmonH, use = "pairwise.complete.obs")
```
We can see no significant correlation within the seasons.
Therefore we can now look at the lagged correlation for 1 season and 2 seasons 

```{r}
#lag of one season
#correlation between inter-monsoon humdity and pre-monsoon cases 
cor(stacked_hum_den$intermonH, stacked_hum_den$premon, use = "pairwise.complete.obs" )

#correlation between the pre-monsoon humidity season and the monsoon dengue cases 
cor(stacked_hum_den$premonH, stacked_hum_den$mon, use = "pairwise.complete.obs")

#correlation between monsoon humidity and post-monsoon dengue cases
cor(stacked_hum_den$premonH, stacked_hum_den$postmon, use = "pairwise.complete.obs")

#correlation between post-monsoon humidity and inter-monsoon dengue cases 
cor(stacked_hum_den$postmonH, stacked_hum_den$intermon, use = "pairwise.complete.obs")


#lag of 2 seasons 
#correlation between inter-monsoon humidity and monsoon dengue cases 
cor(stacked_hum_den$intermonH, stacked_hum_den$mon, use = "pairwise.complete.obs")

#correlation between pre-monsoon humidity and post-monsoon dengue cases 
cor(stacked_hum_den$premonH, stacked_hum_den$postmon, use = "pairwise.complete.obs")

#correlation between monsoon humidity and inter-monsoon dengue cases 
cor(stacked_hum_den$monH, stacked_hum_den$intermon, use = "pairwise.complete.obs")

#correlation between post-monsoon humidity and pre-monsoon dengue cases 
cor(stacked_hum_den$postmonH, stacked_hum_den$premon, use = "pairwise.complete.obs")
```

The highest 3 correlations are between pre-monsoon humidity and monsoon dengue cases, monsoon humidity and post-monsoon dengue cases, and pre-monsoon humidity and post-monsoon dengue cases. 
All correlations are still very small (between -0.145 and -0.164) and interestingly also negative. 

Let's visualize these with scatterplots 
```{r}
#a linear model can be used to add the correlation to the scatterplot 
lm1 <- lm(stacked_hum_den$premonH ~ stacked_hum_den$mon)
lm2 <- lm(stacked_hum_den$monH ~ stacked_hum_den$postmon)
lm3 <- lm(stacked_hum_den$premonH ~ stacked_hum_den$postmon)

plot(stacked_hum_den$premonH, stacked_hum_den$mon,
     pch=20,
     xlab='pre-monsoon humidity',
     ylab='monsoon dengue cases')
abline(lm1, col = 4, lwd = 3)

plot(stacked_hum_den$monH, stacked_hum_den$postmon,pch=20,
     xlab='monsoon humidity',
     ylab='post-monsoon dengue cases')
abline(lm2, col = 4, lwd = 3)

plot(stacked_hum_den$premonH, stacked_hum_den$postmon,pch=20,
     xlab='pre-monsoon humidity',
     ylab='post-monsoon dengue cases')
abline(lm3, col = 4, lwd = 3)
```

The most interesting correlation would be the correlation between pre-monsoon humidity and monsoon dengue cases or between the monsoon humidity and the post monsoon dengue cases.
The correlation does not seem great, however let's test this with a correlation test to see if it might actually not exist at all.
The H0 hypothesis assumes that the actual correlation between the pre-monsoon humidity and the monsoon season dengue cases is equal to zero 
The H1 hypothesis assumes that the actual correlation between the pre-monsoon humidity and the monsoon season dengue cases is less than zero

The same H0 and H1 Hypothesis apply to the correlation between monsoon humidity and post-monsoon dengue cases.
```{r}
#correlation between pre-monsoon humidity and monsoon dengue cases 
cor.test(stacked_hum_den$premonH, stacked_hum_den$mon, use = "pairwise.complete.obs", method = "pearson", alternative = "less")

#correlation between monsoon humidity and post-monsoon dengue cases
cor.test(stacked_hum_den$monH, stacked_hum_den$postmon, use = "pairwise.complete.obs", method = "pearson", alternative = "less")
```

The correlation test returns a p-value that is smaller than 0.05 for both tests, which indicates that the H0 hypothesis is rejected and the H1 hypothesis is assumed. This is not what one would expect.

Correlation between the monsoon seasons within the variables 
A paper mentions, that even slightly higher pre-epidemic case numbers have a significant impact on the epidemic case numbers, maybe this same principle can be seen within the monsoon seasons if we assume the pre-monsoon season as the pre-epidemic stage and the monsoon season as the epidemic stage 
```{r}
cor(stacked_hum_den$intermon, stacked_hum_den$premon, use = "pairwise.complete.obs")
cor(stacked_hum_den$premon, stacked_hum_den$mon, use = "pairwise.complete.obs" )
cor(stacked_hum_den$mon, stacked_hum_den$postmon, use = "pairwise.complete.obs")
cor(stacked_hum_den$postmon, stacked_hum_den$intermon, use = "pairwise.complete.obs")
```
The correlation between the pre-monsoon dengue cases and the monsoon dengue cases is by far the highest with an r-value of 0.73. 
```{r}
model <- lm(stacked_hum_den$mon ~ stacked_hum_den$premon)
plot(stacked_hum_den$premon, stacked_hum_den$mon, pch=20,
     xlab='pre-monsoon cases',
     ylab='monsoon dengue cases'
     )
abline(model, col = 4, lwd = 3)
```

Let's test the correlation again with a correlation test 
H0: the actual correlation between the pre-monsoon cases and the monsoon season dengue cases is equal to zero 
H1: the actual correlation between the pre-monsoon cases and the monsoon season dengue cases is greater than zero 
```{r}
cor.test(stacked_hum_den$premon, stacked_hum_den$mon, use = "pairwise.complete.obs", method = "pearson", alternative = "greater")
```

```{r}
#the linear model was used before to plot an Abline in the scatterplot, we just need to summerize it
summary(model)
```
We can see that the regression model can explain 54% of the variance in this relationship as well as with a p-value smaller than 0.05 using the linear regression model here would be significantly better to predict the monsoon dengue cases than the zero model. 

```{r}
cases = c()
dengue_monthly = data.frame()


for(i in 0:14){
    position = as.vector(5+(13*i))
    cases = dengue_norm[,position]
    if (i == 0){
      dengue = cases
    }
      else {
        dengue = c(dengue_jan, cases)
      }
}  
  print(dengue)
```

