---
title: "Correlation_statisitcal_tests"
author: "Paula"
date: "2023-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First load the data frame with the dengue cases 
```{r}
filePath = paste(getwd(), "dengue_data_com.csv",sep = "/")
combined_df <- read.csv("../dengue_data/dengue_data_com.csv")
print(combined_df)
#convertieren der dengue cases zu zahlen statt charaktern
combined_df$dengue_cases <- as.integer(combined_df$dengue_cases)
print(combined_df)
```

Convert missing values to value zero 
```{r}
combined_df[is.na(combined_df)] <- 0
print(combined_df)
```
load the climate data 
```{r}
install.packages("readxl")
library(readxl)

sheet_names = excel_sheets("../era5_data_2006_2020_thailand_monmean.xlsx")

#load the humidity data 
T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")
```

Humidity dataframe unwichtige Daten entfernen und correlation für Januar 2006 als Beispiel berechnen
```{r}
#unwichtige Columns entfernen 
i.remove <- which(colnames(T.huss.mon) %in% c("Number","Longitude", "Latitude"))
T.huss.mon1 <- T.huss.mon[,-i.remove]

#alphabetisch ordnen
sorted_df <- T.huss.mon1[order(T.huss.mon1$ADM1_EN), ]
print(sorted_df)

#Bueng Kan ist zu Beginn nicht vorhanden in dengue dataframe 
sorted_df <- sorted_df[-4,]

#Beispiel an Januar 2006
dengue_2006_jan <- subset(combined_df, combined_df$Year == "2006" & combined_df$Month == "Jan")
cor(sorted_df$'2006.01.', dengue_2006_jan$dengue_cases, method='pearson')
cor(sorted_df$'2006.01.', dengue_2006_jan$dengue_cases, method='spearman')

#plotten der Daten
plot(sorted_df$'2006.01.', dengue_2006_jan$dengue_cases,
     pch=20,
     xlab='humidity',
     ylab='dengue cases')

```
Bangkok is an outlier, try to calculate the correlation without Bangkok 
```{r}
#Bangkok entfernen aus beiden Dataframes 
without_bangkok_d <- dengue_2006_jan[-3,]
without_bangkok_c <- sorted_df[-3,]

#correlation of humidity and dengue cases 
cor(without_bangkok_c$'2006.01.', without_bangkok_d$dengue_cases, method='pearson')
cor(without_bangkok_c$'2006.01.', without_bangkok_d$dengue_cases, method='spearman')

#pearson correlation much higher without bangkok but spearman correlation is a little lower 
#plot without bangkok 
plot(without_bangkok_c$'2006.01.', without_bangkok_d$dengue_cases,
     pch=20,
     xlab='humidity',
     ylab='dengue cases')
```
Let's try the lagged correlation for 1 month, 2 months and 3 months
```{r}
#get February 2006 from dengue dataframe
dengue_2006_feb <-  subset(combined_df, combined_df$Year == "2006" & combined_df$Month == "Feb")
#get March 2006 from dengue dataframe 
dengue_2006_mar <- subset(combined_df, combined_df$Year == "2006" & combined_df$Month == "Mar")
#get April 2006 from dengue dataframe
dengue_2006_apr <- subset(combined_df, combined_df$Year == "2006" & combined_df$Month == "Apr")

#correlation for 1 month lag, humidity data one month before dengue cases 
cor(sorted_df$'2006.01.', dengue_2006_feb$dengue_cases, method='pearson')
cor(sorted_df$'2006.01.', dengue_2006_feb$dengue_cases, method='spearman')

#two months lagged
cor(sorted_df$'2006.01.', dengue_2006_mar$dengue_cases, method='pearson')
cor(sorted_df$'2006.01.', dengue_2006_mar$dengue_cases, method='spearman')

#three months lagged 
cor(sorted_df$'2006.01.', dengue_2006_apr$dengue_cases, method='pearson')
cor(sorted_df$'2006.01.', dengue_2006_apr$dengue_cases, method='spearman')
```
Populationsdaten laden 
```{r}
popu_06 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2006_population.xlsx")
popu_07 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2007_population.xlsx")
popu_08 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2008_population.xlsx")
popu_09 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2009_population.xlsx")
popu_10 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2010_population.xlsx")
popu_11 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2011_population.xlsx")
popu_12_21 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2012-2021-population-sector_01_11102_EN_.xlsx")
```

Normalisieren der dengue cases auf 100'000 personen 
```{r}

```

