---
title: "Correlation_statisitcal_tests"
author: "Paula"
date: "2023-06-20"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sorting of Dataframes

This markdown is mainly used to normalize and sort data into dataframes most useful for correlation tests and other test.

```{r}
#install and load needed Packages 
install.packages("readxl")
library(readxl)

#load population data 
popu_06 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2006_population.xlsx")
popu_07 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2007_population.xlsx")
popu_08 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2008_population.xlsx")
popu_09 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2009_population.xlsx")
popu_10 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2010_population.xlsx")
popu_11 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2011_population.xlsx")
popu_12_21 <- read_excel("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\population_data\\2012-2021-population-sector_01_11102_EN_.xlsx")

#load the humidity data 
sheet_names = excel_sheets("../era5_data_2006_2020_thailand_monmean.xlsx")
T.huss.mon = read_excel("../era5_data_2006_2020_thailand_monmean.xlsx", sheet="era5_huss_2006_2020_thailand_mo")

#load the dengue data 
dengue <- read.csv("C:\\Users\\paula\\Documents\\Uni Heidelberg\\4. Semester\\Bioinformatik\\github\\topic04_team01\\dengue_data\\dengue_data_com2.csv")
```

## Creating a normalized dengue Data frame, without removing the NA's

The humidity data is not sorted alphabetically and contains columns that aren't needed, let's remove them and sort the dataframe

```{r}
#remove columns that aren't needed
i.remove <- which(colnames(T.huss.mon) %in% c("Number","Longitude", "Latitude"))
T.huss.mon1 <- T.huss.mon[,-i.remove]
#sort them alphabetically
T.huss.mon1 <- T.huss.mon1[order(T.huss.mon1$ADM1_EN), ]
```

The population data is stored in different excel files and to work with them, we need them in one single dataframe

```{r}
#bind the coloumns in the different excel files together in one new dataframe
popu_06_20 <- cbind(popu_06, "2007" = popu_07$`Total population`, "2008" = popu_08$`Total population`, "2009" = popu_09$`Total population`, "2010" = popu_10$`Total population`, "2011" =  popu_11$`Total population`, popu_12_21[,3:12])

#rename the coloumn that contains the population data from 2006 to "2006"
names(popu_06_20)[names(popu_06_20) == "Total population"] <- "2006"

#to avoid having to run the entire code to get this version of the dataframe, let's save this as a csv file as well
filePath = paste(dirname(getwd()), "population_data", "population_06_20.csv",sep = "/")
write.csv(popu_06_20, filePath)

```

To avoid the correlation being distorted by the differences in population, it is necessary to normalize the dengue data to cases per 100'000 people.
To do this the cases in each year and province are divided by the population of the respective province and respective year and then multiplied by 100'000.

```{r}
#create a new dataframe, where the normalized data will be saved 
dengue_norm = data.frame(dengue)

# loops over all years in population dataframe
for(i in 3 : ncol(popu_06_20)) {
  # stores yearly population in local variable
  yearly_pop = popu_06_20[,i]
  # loops over total and monthly columns of each year
  for(j in 0:12) {
    # calculates index of corresponding monthly cases
    # cases start at column index 3
    # every year is displayed over 13 columns, thus every year cursor jumps by 13 columns (i-3)*13
    # as years each are displayed in 13 columns, each month the cursor jumps over by 1
    case_index = 3 + (i-3)*13 + j
    
    selected_cases = dengue[,case_index]
    normalized_cases = selected_cases / yearly_pop * 100000
    
    dengue_norm[,case_index] = normalized_cases
  }
}

#let's save the normalized data in the repository 
filePath = paste(dirname(getwd()), "dengue_data", "dengue_norm.csv",sep = "/")
write.csv(dengue_norm, filePath)
```

## Dengue Dataframe that does not contain NA's

This normalized dataframe might be useful for other observations but for the correlations between dengue seasons, it is important to replace NA's with an educated guess to avoid distortion of the correlation as the monsoon seasons of the dengue cases are calculated via the sum of the respective monsoon season months.
This educated guess is constructed as followed:

1\.
For Provinces that are missing entire years and do not have a total per year given, the values from January to December for each year are replaced by the mean cases of each month during the year.

2\.
The new total then has to be calculated and added to the total cases column of the year and respective province.

3\.
For Provinces where only some variables are missing and the total per year is given, we might try a different approach.

3.1 Firstly we need to calculate a dataframe with the percentages of the cases in each month in comparision to the total cases per year.
For example the percentage of cases in january 2006 in the Amnat Charoen Province is calculated by deviding the cases of january 2006 by the total cases of 2006.

3.2 Secondly we need to calculate the mean of these percentages for each month.
If we take the example of january 2006 again this means adding up all percentages in the january 2006 and deviding them by the number of values that where added up.

3.3 Lastly the single NA's are replaced by the percentage average of the respective month in the respective year multiplied by the total cases of the year.
Again for example if the Amnat Charoen Province is missing a value for january 2006, one would multiply the mean of the january 2006 percentage as described in step 3.2 with the total cases in Amnat Charoen in 2006.

I am aware that this process might still not match the actual cases exactly but it is a good attempt to minimize the effects of NA's on the correlation tests when the monsoon seasons (based on sums) are used as a base for the correlation tests.

### Step 1 and 2

```{r}
#make a copy of the dengue_data_com2 dataframe 
dengue_wo_NAs = data.frame(dengue)

#Provinces that are missing entire years are the Bungkan province from 2006-2010 and the Kalasin province form 2008-2009 and 2011-2013. For these provinces the step described in 1. will be applied 

##the province Bungkan is in row number 4  and December 2010 is in column 67
##Kalasin is in row 14, total 2008 to december 2009 are in columns 29 til 54, and total 2011 to december 2013 are in the columns 68 til 106

# loops over all years in dengue dataframe
for(i in 2006 : 2013) {
  
  # loops over total and monthly columns of each year
  for(j in 1:12) {
    # calculates index of corresponding monthly cases
    # cases start at column index 3
    # every year is displayed over 13 columns, thus every year cursor jumps by 13 columns (i-3)*13
    # as years each are displayed in 13 columns, each month the cursor jumps over by 1
    case_index = 3 + (i-2006)*13 + j
    
    selected_cases = dengue_wo_NAs[,case_index]
    #calculate the mean of each month of each year
    mean_cases = mean(selected_cases, na.rm = TRUE)

    #replace the missing values with the means just calculated for the specific provinces and years listed above
    if(i >= 2006 && i <= 2011) {
      dengue_wo_NAs[4,case_index] = mean_cases
    }

    if((i >= 2008 && i <= 2009) || (i >= 2011 && i <= 2013)) {
      dengue_wo_NAs[14,case_index] = mean_cases
    }
    
    if (j == 12) {
      # after multiplying the definition of the arithmetic mean by the number of elements over which the
      # mean is calculated, one finds the sum of number of cases --> easier than using a dedicated sum
      # function ; )
      lower_bound = case_index - 11
      
      if(is.na(dengue_wo_NAs[4,case_index-j])) {
        yearly_total = mean(as.numeric(dengue_wo_NAs[4,lower_bound:case_index]))*12
        dengue_wo_NAs[4,case_index-j] = yearly_total
      }
      if(is.na(dengue_wo_NAs[14,case_index-j])) {
        yearly_total = mean(as.numeric(dengue_wo_NAs[14,lower_bound:case_index]))*12
        dengue_wo_NAs[14,case_index-j] = yearly_total
      }
    }
    
  }
  
}

```

### Step 3 Step 3.1

```{r}
#make percentage dataframe
dengue_percentages = data.frame(dengue)


for(i in 2006 : 2020) {
  
  total_index = 3 + (i-2006)*13
  
  # loops over total and monthly columns of each year
  for(j in 0:12) {
    
    month_index = 3 + (i-2006)*13 + j
    
    dengue_percentages[,month_index] = dengue_wo_NAs[,month_index] / dengue_wo_NAs[,total_index]
    
  }
  
}
```

### step 3.2 and 3.3

```{r}
# loops over all years in dengue dataframe without NAs 
for(i in 2006 : 2020) {

total_index = 3 + (i-2006)*13
total_cases = dengue_wo_NAs[,total_index]

# loops over total and monthly columns of each year
  for(j in 1:12) {
    
    month_index = 3 + (i-2006)*13 + j
    column_mean_percent = mean(dengue_percentages[,month_index], na.rm = TRUE)
    replacement_value = column_mean_percent * total_cases
    
    dengue_wo_NAs[, month_index] <- ifelse(is.na(dengue_wo_NAs[, month_index]), replacement_value, dengue_wo_NAs[, month_index])
      
  }

}
  
#save as csv file 
filePath = paste(dirname(getwd()), "dengue_data", "dengue_without_NAs.csv",sep = "/")
write.csv(dengue_wo_NAs, filePath)


#Through replacing the NA's with mean values, the total values might differ a little bit. However as these might be needed in further analysis, I would rather not distort these actual recorded case numbers
```

## Normalized dataframe without NA's

Again this dataframe needs to be normalized with the population data

```{r}
#create a new dataframe, where the normalized data will be saved 
dengue_wo_NA_norm = data.frame(dengue_wo_NAs)

# loops over all years in population dataframe
for(i in 3 : ncol(popu_06_20)) {
  # stores yearly population in local variable
  yearly_pop = popu_06_20[,i]
  # loops over total and monthly columns of each year
  for(j in 0:12) {
    # calculates index of corresponding monthly cases
    # cases start at column index 3
    # every year is displayed over 13 columns, thus every year cursor jumps by 13 columns (i-3)*13
    # as years each are displayed in 13 columns, each month the cursor jumps over by 1
    case_index = 3 + (i-3)*13 + j
    
    selected_cases = dengue_wo_NA_norm[,case_index]
    normalized_cases = selected_cases / yearly_pop * 100000
    
    dengue_wo_NA_norm[,case_index] = normalized_cases
  }
}

dengue_wo_NA_norm[sapply(dengue_wo_NA_norm, is.infinite)] <- NA

#let's save the normalized data in the repository 
filePath = paste(dirname(getwd()), "dengue_data", "dengue_without_NAs_norm.csv",sep = "/")
write.csv(dengue_wo_NA_norm, filePath)
```

## Monsoon season dataframes

### Normalized dengue Dataframe without NA's sorted into monsoon seasons

As Thailand is strongly influenced by its monsoon seasons, we sort the dengue cases and humidity data into 4 different monsoon seasons.
The inter-monsoon season months are December, January and February, the pre-monsoon seasons are March, April and May, the monsoon seasons are June, July, August and September and finally the post-monsoon seasons are October and November.
The Dengue monsoon seasons will be calculated by summing the respective months up.
As Bungkan has again gained NAs because the population data from 2006-2011 is missing, I decided to loose this one province.
The inter-monsoon season of 2006 will also have to be discarded in the correlation test, as it only comprises of a sum of 2 months instead of 3.
For the humidity data mean for the months mentioned above is used.
To avoid clashes with different vector lengths, Bungkan will be removed here as well

```{r}
#normalized dengue data without NA's
##the bungkan province needs to be removed in row 4 
dengue_wo_NA_norm 

## means of the inter-monsoon seasons over the years 
intermon_06_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2006.01","X2006.02")])
intermon_07_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2006.12", "X2007.01", "X2007.02")])
intermon_08_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2007.12", "X2008.01", "X2008.02")])
intermon_09_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2008.12", "X2009.01", "X2009.02")])
intermon_10_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2009.12", "X2010.01", "X2010.02")])
intermon_11_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2010.12", "X2011.01", "X2011.02")])
intermon_12_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2011.12", "X2012.01", "X2012.02")])
intermon_13_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2012.12", "X2013.01", "X2013.02")])
intermon_14_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2013.12", "X2014.01", "X2014.02")])
intermon_15_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2014.12", "X2015.01", "X2015.02")])
intermon_16_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2015.12", "X2016.01", "X2016.02")])
intermon_17_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2016.12", "X2017.01", "X2017.02")])
intermon_18_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2017.12", "X2018.01", "X2018.02")])
intermon_19_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2018.12", "X2019.01", "X2019.02")])
intermon_20_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2019.12", "X2020.01", "X2020.02")])

##means of the pre-monsoon seasons over the years 
premon_06_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2006.03", "X2006.04", "X2006.05")])
premon_07_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2007.03", "X2007.04", "X2007.05")])
premon_08_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2008.03", "X2008.04", "X2008.05")])
premon_09_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2009.03", "X2009.04", "X2009.05")])
premon_10_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2010.03", "X2010.04", "X2010.05")])
premon_11_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2011.03", "X2011.04", "X2011.05")])
premon_12_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2012.03", "X2012.04", "X2012.05")])
premon_13_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2013.03", "X2013.04", "X2013.05")])
premon_14_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2014.03", "X2014.04", "X2014.05")])
premon_15_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2015.03", "X2015.04", "X2015.05")])
premon_16_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2016.03", "X2016.04", "X2016.05")])
premon_17_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2017.03", "X2017.04", "X2017.05")])
premon_18_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2018.03", "X2018.04", "X2018.05")])
premon_19_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2019.03", "X2019.04", "X2019.05")])
premon_20_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2020.03", "X2020.04", "X2020.05")])

##means of the monsoon seasons over the years 
mon_06_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2006.06", "X2006.07", "X2006.08", "X2006.09")])
mon_07_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2007.06", "X2007.07", "X2007.08", "X2007.09")])
mon_08_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2008.06", "X2008.07", "X2008.08", "X2008.09")])
mon_09_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2009.06", "X2009.07", "X2009.08", "X2009.09")])
mon_10_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2010.06", "X2010.07", "X2010.08", "X2010.09")])
mon_11_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2011.06", "X2011.07", "X2011.08", "X2011.09")])
mon_12_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2012.06", "X2012.07", "X2012.08", "X2012.09")])
mon_13_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2013.06", "X2013.07", "X2013.08", "X2013.09")])
mon_14_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2014.06", "X2014.07", "X2014.08", "X2014.09")])
mon_15_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2015.06", "X2015.07", "X2015.08", "X2015.09")])
mon_16_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2016.06", "X2016.07", "X2016.08", "X2016.09")])
mon_17_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2017.06", "X2017.07", "X2017.08", "X2017.09")])
mon_18_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2018.06", "X2018.07", "X2018.08", "X2018.09")])
mon_19_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2019.06", "X2019.07", "X2019.08", "X2019.09")])
mon_20_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2020.06", "X2020.07", "X2020.08", "X2020.09")])

##means of the post-monsoon seasons over the years
postmon_06_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2006.10", "X2006.11")])
postmon_07_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2007.10", "X2007.11")])
postmon_08_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2008.10", "X2008.11")])
postmon_09_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2009.10", "X2009.11")])
postmon_10_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2010.10", "X2010.11")])
postmon_11_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2011.10", "X2011.11")])
postmon_12_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2012.10", "X2012.11")])
postmon_13_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2013.10", "X2013.11")])
postmon_14_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2014.10", "X2014.11")])
postmon_15_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2015.10", "X2015.11")])
postmon_16_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2016.10", "X2016.11")])
postmon_17_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2017.10", "X2017.11")])
postmon_18_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2018.10", "X2018.11")])
postmon_19_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2019.10", "X2019.11")])
postmon_20_mean <- rowMeans(dengue_wo_NA_norm[, c( "X2020.10", "X2020.11")])
```

let's combine all these vectors into a dataframe

```{r}
#we need the provinces 
Provinces <- dengue_wo_NA_norm$Reporting.areas

#let's save them all in a new dataframe 
dengue_mon_mean <- cbind(Provinces, intermon_06_mean, premon_06_mean, mon_06_mean, postmon_06_mean,intermon_07_mean, premon_07_mean, mon_07_mean, postmon_07_mean,  intermon_08_mean, premon_08_mean, mon_08_mean, postmon_08_mean, intermon_09_mean, premon_09_mean, mon_09_mean, postmon_09_mean, intermon_10_mean, premon_10_mean, mon_10_mean, postmon_10_mean, intermon_11_mean, premon_11_mean, mon_11_mean, postmon_11_mean, intermon_12_mean, premon_12_mean, mon_12_mean, postmon_12_mean, intermon_13_mean, premon_13_mean, mon_13_mean, postmon_13_mean, intermon_14_mean, premon_14_mean, mon_14_mean, postmon_14_mean, intermon_15_mean, premon_15_mean, mon_15_mean, postmon_15_mean, intermon_16_mean, premon_16_mean, mon_16_mean, postmon_16_mean, intermon_17_mean, premon_17_mean, mon_17_mean, postmon_17_mean, intermon_18_mean, premon_18_mean, mon_18_mean, postmon_18_mean, intermon_19_mean, premon_19_mean, mon_19_mean, postmon_19_mean, intermon_20_mean, premon_20_mean, mon_20_mean, postmon_20_mean)

#turn it into a dataframe
dengue_mon_mean <- as.data.frame(dengue_mon_mean)

#turn the character columns into numeric values 
numColumns1 <- names(dengue_mon_mean)[-which(names(dengue_mon_mean) == "Provinces")]
dengue_mon_mean[numColumns1] <- lapply(dengue_mon_mean[numColumns1], as.numeric)
str(dengue_mon_mean)

#let's save yet another dataframe here 
filePath = paste(dirname(getwd()), "dengue_data", "dengue_mon_mean.csv",sep = "/")
write.csv(dengue_mon_mean, filePath)
```

### Humidity Dataframe without NA's sorted into monsoon seasons

We need another dataframe with the means of the humidity during each monsoon season to be able to calculate the correlation between the two, so let's do that

```{r}
#first let's calculate the mean of the intermonsoon seasons 
intermonH_06 <- rowMeans(T.huss.mon1[,c("2006.01.", "2006.02.")])
intermonH_07 <- rowMeans(T.huss.mon1[,c("2006.12.", "2007.01.", "2007.02.")])
intermonH_08 <- rowMeans(T.huss.mon1[,c("2007.12.", "2008.01.", "2008.02.")])
intermonH_09 <- rowMeans(T.huss.mon1[,c("2008.12.", "2009.01.", "2009.02.")])
intermonH_10 <- rowMeans(T.huss.mon1[,c("2009.12.", "2010.01.", "2010.02.")])
intermonH_11 <- rowMeans(T.huss.mon1[,c("2010.12.", "2011.01.", "2011.02.")])
intermonH_12 <- rowMeans(T.huss.mon1[,c("2011.12.", "2012.01.", "2012.02.")])
intermonH_13 <- rowMeans(T.huss.mon1[,c("2012.12.", "2013.01.", "2013.02.")])
intermonH_14 <- rowMeans(T.huss.mon1[,c("2013.12.", "2014.01.", "2014.02.")])
intermonH_15 <- rowMeans(T.huss.mon1[,c("2014.12.", "2015.01.", "2015.02.")])
intermonH_16 <- rowMeans(T.huss.mon1[,c("2015.12.", "2016.01.", "2016.02.")])
intermonH_17 <- rowMeans(T.huss.mon1[,c("2016.12.", "2017.01.", "2017.02.")])
intermonH_18 <- rowMeans(T.huss.mon1[,c("2017.12.", "2018.01.", "2018.02.")])
intermonH_19 <- rowMeans(T.huss.mon1[,c("2018.12.", "2019.01.", "2019.02.")])
intermonH_20 <- rowMeans(T.huss.mon1[,c("2019.12.", "2020.01.", "2020.02.")])

#Do the same for the pre-monsoon months
premonH_06 <- rowMeans(T.huss.mon1[,c("2006.03.", "2006.04.", "2006.05.")])
premonH_07 <- rowMeans(T.huss.mon1[,c("2007.03.", "2007.04.", "2007.05.")])
premonH_08 <- rowMeans(T.huss.mon1[,c("2008.03.", "2008.04.", "2008.05.")])
premonH_09 <- rowMeans(T.huss.mon1[,c("2009.03.", "2009.04.", "2009.05.")])
premonH_10 <- rowMeans(T.huss.mon1[,c("2010.03.", "2010.04.", "2010.05.")])
premonH_11 <- rowMeans(T.huss.mon1[,c("2011.03.", "2011.04.", "2011.05.")])
premonH_12 <- rowMeans(T.huss.mon1[,c("2012.03.", "2012.04.", "2012.05.")])
premonH_13 <- rowMeans(T.huss.mon1[,c("2013.03.", "2013.04.", "2013.05.")])
premonH_14 <- rowMeans(T.huss.mon1[,c("2014.03.", "2014.04.", "2014.05.")])
premonH_15 <- rowMeans(T.huss.mon1[,c("2015.03.", "2015.04.", "2015.05.")])
premonH_16 <- rowMeans(T.huss.mon1[,c("2016.03.", "2016.04.", "2016.05.")])
premonH_17 <- rowMeans(T.huss.mon1[,c("2017.03.", "2017.04.", "2017.05.")])
premonH_18 <- rowMeans(T.huss.mon1[,c("2018.03.", "2018.04.", "2018.05.")])
premonH_19 <- rowMeans(T.huss.mon1[,c("2019.03.", "2019.04.", "2019.05.")])
premonH_20 <- rowMeans(T.huss.mon1[,c("2020.03.", "2020.04.", "2020.05.")])

#again for the monsoon months
monH_06 <- rowMeans(T.huss.mon1[,c("2006.06.", "2006.07.", "2006.08.", "2006.09.")])
monH_07 <- rowMeans(T.huss.mon1[,c("2007.06.", "2007.07.", "2007.08.", "2007.09.")])
monH_08 <- rowMeans(T.huss.mon1[,c("2008.06.", "2008.07.", "2008.08.", "2008.09.")])
monH_09 <- rowMeans(T.huss.mon1[,c("2009.06.", "2009.07.", "2009.08.", "2009.09.")])
monH_10 <- rowMeans(T.huss.mon1[,c("2010.06.", "2010.07.", "2010.08.", "2010.09.")])
monH_11 <- rowMeans(T.huss.mon1[,c("2011.06.", "2011.07.", "2011.08.", "2011.09.")])
monH_12 <- rowMeans(T.huss.mon1[,c("2012.06.", "2012.07.", "2012.08.", "2012.09.")])
monH_13 <- rowMeans(T.huss.mon1[,c("2013.06.", "2013.07.", "2013.08.", "2013.09.")])
monH_14 <- rowMeans(T.huss.mon1[,c("2014.06.", "2014.07.", "2014.08.", "2014.09.")])
monH_15 <- rowMeans(T.huss.mon1[,c("2015.06.", "2015.07.", "2015.08.", "2015.09.")])
monH_16 <- rowMeans(T.huss.mon1[,c("2016.06.", "2016.07.", "2016.08.", "2016.09.")])
monH_17 <- rowMeans(T.huss.mon1[,c("2017.06.", "2017.07.", "2017.08.", "2017.09.")])
monH_18 <- rowMeans(T.huss.mon1[,c("2018.06.", "2018.07.", "2018.08.", "2018.09.")])
monH_19 <- rowMeans(T.huss.mon1[,c("2019.06.", "2019.07.", "2019.08.", "2019.09.")])
monH_20 <- rowMeans(T.huss.mon1[,c("2020.06.", "2020.07.", "2020.08.", "2020.09.")])

#again for the postmonsoon seasons
postmonH_06 <- rowMeans(T.huss.mon1[,c("2006.10.", "2006.11.")])
postmonH_07 <- rowMeans(T.huss.mon1[,c("2007.10.", "2007.11.")])
postmonH_08 <- rowMeans(T.huss.mon1[,c("2008.10.", "2008.11.")])
postmonH_09 <- rowMeans(T.huss.mon1[,c("2009.10.", "2009.11.")])
postmonH_10 <- rowMeans(T.huss.mon1[,c("2010.10.", "2010.11.")])
postmonH_11 <- rowMeans(T.huss.mon1[,c("2011.10.", "2011.11.")])
postmonH_12 <- rowMeans(T.huss.mon1[,c("2012.10.", "2012.11.")])
postmonH_13 <- rowMeans(T.huss.mon1[,c("2013.10.", "2013.11.")])
postmonH_14 <- rowMeans(T.huss.mon1[,c("2014.10.", "2014.11.")])
postmonH_15 <- rowMeans(T.huss.mon1[,c("2015.10.", "2015.11.")])
postmonH_16 <- rowMeans(T.huss.mon1[,c("2016.10.", "2016.11.")])
postmonH_17 <- rowMeans(T.huss.mon1[,c("2017.10.", "2017.11.")])
postmonH_18 <- rowMeans(T.huss.mon1[,c("2018.10.", "2018.11.")])
postmonH_19 <- rowMeans(T.huss.mon1[,c("2019.10.", "2019.11.")])
postmonH_20 <- rowMeans(T.huss.mon1[,c("2020.10.", "2020.11.")])

#put all the means together in a new dataframe
Provinces <- T.huss.mon1$ADM1_EN
humidity_mon <- cbind(Provinces, intermonH_06, premonH_06, monH_06, postmonH_06, intermonH_07, premonH_07, monH_07, postmonH_07, intermonH_08, premonH_08, monH_08, postmonH_08, intermonH_09, premonH_09, monH_09, postmonH_09, intermonH_10, premonH_10, monH_10, postmonH_10, intermonH_11, premonH_11, monH_11, postmonH_11, intermonH_12, premonH_12, monH_12, postmonH_12, intermonH_13, premonH_13, monH_13, postmonH_13, intermonH_14, premonH_14, monH_14, postmonH_14, intermonH_15, premonH_15, monH_15, postmonH_15, intermonH_16, premonH_16, monH_16, postmonH_16, intermonH_17, premonH_17, monH_17, postmonH_17, intermonH_18, premonH_18, monH_18, postmonH_18, intermonH_19, premonH_19, monH_19, postmonH_19, intermonH_20, premonH_20, monH_20, postmonH_20)

#turn it into a dataframe
humidity_mon <- as.data.frame(humidity_mon)

#save the dataframe 
filePath = paste(dirname(getwd()), "geo_data", "humidity_mon.csv",sep = "/")
write.csv(humidity_mon, filePath)
```

## Yearly stacked monsoon season dataframe with both humidity and dengue cases

let's just quickly stack them atop each other

```{r}
##let's stack the data for the years atop of each other for easier calculation with the cor() function 

#start with the inter-monsoon dengue cases
stacked_intermon_mean <- sapply(sprintf("intermon_%02d_mean", 6:20), function(var_name) get(var_name))

#next pre-monsoon dengue cases 
stacked_premon_mean <- sapply(sprintf("premon_%02d_mean", 6:20), function(var_name) get(var_name))

#next monsoon dengue cases 
stacked_mon_mean <- sapply(sprintf("mon_%02d_mean", 6:20), function(var_name) get(var_name))

#next post-monsoon dengue cases 
stacked_postmon_mean <- sapply(sprintf("postmon_%02d_mean", 6:20), function(var_name) get(var_name))

#we need the provinces in the stacked dataframe too
Provinces <- T.huss.mon1$ADM1_EN


#stack them 14 times 
stacked_provinces <- rep(Provinces, times = 15)

#make another dataframe to calculate the correlation 
stacked_mon_mean <- cbind(stacked_provinces, stacked_intermon_mean, stacked_premon_mean, stacked_mon_mean, stacked_postmon_mean)

```

```{r}
#normalized dengue data 
##the normalized dengue data contains NA's, if the sum is used, these NA's distort the sum and therefore the correlation, we first need to replace the NA's with mean values of the respective year and month 
## sums of the inter-monsoon seasons over the years 
intermon_06 <- dengue_norm$X2006.01 + dengue_norm$X2006.02
intermon_07 <- rowSums(dengue_norm[, c( "X2006.12", "X2007.01", "X2007.02")], na.rm = FALSE)
intermon_08 <- rowSums(dengue_norm[, c( "X2007.12", "X2008.01", "X2008.02")], na.rm = FALSE)
intermon_09 <- rowSums(dengue_norm[, c( "X2008.12", "X2009.01", "X2009.02")], na.rm = FALSE)
intermon_10 <- rowSums(dengue_norm[, c( "X2009.12", "X2010.01", "X2010.02")], na.rm = FALSE)
intermon_11 <- rowSums(dengue_norm[, c( "X2010.12", "X2011.01", "X2011.02")], na.rm = FALSE)
intermon_12 <- rowSums(dengue_norm[, c( "X2011.12", "X2012.01", "X2012.02")], na.rm = FALSE)
intermon_13 <- rowSums(dengue_norm[, c( "X2012.12", "X2013.01", "X2013.02")], na.rm = FALSE)
intermon_14 <- rowSums(dengue_norm[, c( "X2013.12", "X2014.01", "X2014.02")], na.rm = FALSE)
intermon_15 <- rowSums(dengue_norm[, c( "X2014.12", "X2015.01", "X2015.02")], na.rm = FALSE)
intermon_16 <- rowSums(dengue_norm[, c( "X2015.12", "X2016.01", "X2016.02")], na.rm = FALSE)
intermon_17 <- rowSums(dengue_norm[, c( "X2016.12", "X2017.01", "X2017.02")], na.rm = FALSE)
intermon_18 <- rowSums(dengue_norm[, c( "X2017.12", "X2018.01", "X2018.02")], na.rm = FALSE)
intermon_19 <- rowSums(dengue_norm[, c( "X2018.12", "X2019.01", "X2019.02")], na.rm = FALSE)
intermon_20 <- rowSums(dengue_norm[, c( "X2019.12", "X2020.01", "X2020.02")], na.rm = FALSE)

##sums of the pre-monsoon seasons over the years 
premon_06 <- rowSums(dengue_norm[, c( "X2006.03", "X2006.04", "X2006.05")], na.rm = FALSE)
premon_07 <- rowSums(dengue_norm[, c( "X2007.03", "X2007.04", "X2007.05")], na.rm = FALSE)
premon_08 <- rowSums(dengue_norm[, c( "X2008.03", "X2008.04", "X2008.05")], na.rm = FALSE)
premon_09 <- rowSums(dengue_norm[, c( "X2009.03", "X2009.04", "X2009.05")], na.rm = FALSE)
premon_10 <- rowSums(dengue_norm[, c( "X2010.03", "X2010.04", "X2010.05")], na.rm = FALSE)
premon_11 <- rowSums(dengue_norm[, c( "X2011.03", "X2011.04", "X2011.05")], na.rm = FALSE)
premon_12 <- rowSums(dengue_norm[, c( "X2012.03", "X2012.04", "X2012.05")], na.rm = FALSE)
premon_13 <- rowSums(dengue_norm[, c( "X2013.03", "X2013.04", "X2013.05")], na.rm = FALSE)
premon_14 <- rowSums(dengue_norm[, c( "X2014.03", "X2014.04", "X2014.05")], na.rm = FALSE)
premon_15 <- rowSums(dengue_norm[, c( "X2015.03", "X2015.04", "X2015.05")], na.rm = FALSE)
premon_16 <- rowSums(dengue_norm[, c( "X2016.03", "X2016.04", "X2016.05")], na.rm = FALSE)
premon_17 <- rowSums(dengue_norm[, c( "X2017.03", "X2017.04", "X2017.05")], na.rm = FALSE)
premon_18 <- rowSums(dengue_norm[, c( "X2018.03", "X2018.04", "X2018.05")], na.rm = FALSE)
premon_19 <- rowSums(dengue_norm[, c( "X2019.03", "X2019.04", "X2019.05")], na.rm = FALSE)
premon_20 <- rowSums(dengue_norm[, c( "X2020.03", "X2020.04", "X2020.05")], na.rm = FALSE)

##sums of the monsoon seasons over the years 
mon_06 <- rowSums(dengue_norm[, c( "X2006.06", "X2006.07", "X2006.08", "X2006.09")], na.rm = FALSE)
mon_07 <- rowSums(dengue_norm[, c( "X2007.06", "X2007.07", "X2007.08", "X2007.09")], na.rm = FALSE)
mon_08 <- rowSums(dengue_norm[, c( "X2008.06", "X2008.07", "X2008.08", "X2008.09")], na.rm = FALSE)
mon_09 <- rowSums(dengue_norm[, c( "X2009.06", "X2009.07", "X2009.08", "X2009.09")], na.rm = FALSE)
mon_10 <- rowSums(dengue_norm[, c( "X2010.06", "X2010.07", "X2010.08", "X2010.09")], na.rm = FALSE)
mon_11 <- rowSums(dengue_norm[, c( "X2011.06", "X2011.07", "X2011.08", "X2011.09")], na.rm = FALSE)
mon_12 <- rowSums(dengue_norm[, c( "X2012.06", "X2012.07", "X2012.08", "X2012.09")], na.rm = FALSE)
mon_13 <- rowSums(dengue_norm[, c( "X2013.06", "X2013.07", "X2013.08", "X2013.09")], na.rm = FALSE)
mon_14 <- rowSums(dengue_norm[, c( "X2014.06", "X2014.07", "X2014.08", "X2014.09")], na.rm = FALSE)
mon_15 <- rowSums(dengue_norm[, c( "X2015.06", "X2015.07", "X2015.08", "X2015.09")], na.rm = FALSE)
mon_16 <- rowSums(dengue_norm[, c( "X2016.06", "X2016.07", "X2016.08", "X2016.09")], na.rm = FALSE)
mon_17 <- rowSums(dengue_norm[, c( "X2017.06", "X2017.07", "X2017.08", "X2017.09")], na.rm = FALSE)
mon_18 <- rowSums(dengue_norm[, c( "X2018.06", "X2018.07", "X2018.08", "X2018.09")], na.rm = FALSE)
mon_19 <- rowSums(dengue_norm[, c( "X2019.06", "X2019.07", "X2019.08", "X2019.09")], na.rm = FALSE)
mon_20 <- rowSums(dengue_norm[, c( "X2020.06", "X2020.07", "X2020.08", "X2020.09")], na.rm = FALSE)

##sums of the post-monsoon seasons over the years
postmon_06 <- rowSums(dengue_norm[, c( "X2006.10", "X2006.11")], na.rm = FALSE)
postmon_07 <- rowSums(dengue_norm[, c( "X2007.10", "X2007.11")], na.rm = FALSE)
postmon_08 <- rowSums(dengue_norm[, c( "X2008.10", "X2008.11")], na.rm = FALSE)
postmon_09 <- rowSums(dengue_norm[, c( "X2009.10", "X2009.11")], na.rm = FALSE)
postmon_10 <- rowSums(dengue_norm[, c( "X2010.10", "X2010.11")], na.rm = FALSE)
postmon_11 <- rowSums(dengue_norm[, c( "X2011.10", "X2011.11")], na.rm = FALSE)
postmon_12 <- rowSums(dengue_norm[, c( "X2012.10", "X2012.11")], na.rm = FALSE)
postmon_13 <- rowSums(dengue_norm[, c( "X2013.10", "X2013.11")], na.rm = FALSE)
postmon_14 <- rowSums(dengue_norm[, c( "X2014.10", "X2014.11")], na.rm = FALSE)
postmon_15 <- rowSums(dengue_norm[, c( "X2015.10", "X2015.11")], na.rm = FALSE)
postmon_16 <- rowSums(dengue_norm[, c( "X2016.10", "X2016.11")], na.rm = FALSE)
postmon_17 <- rowSums(dengue_norm[, c( "X2017.10", "X2017.11")], na.rm = FALSE)
postmon_18 <- rowSums(dengue_norm[, c( "X2018.10", "X2018.11")], na.rm = FALSE)
postmon_19 <- rowSums(dengue_norm[, c( "X2019.10", "X2019.11")], na.rm = FALSE)
postmon_20 <- rowSums(dengue_norm[, c( "X2020.10", "X2020.11")], na.rm = FALSE)
```

let's combine all these vectors into a dataframe

```{r}
Provinces <- dengue_norm$Reporting.areas
dengue_mon <- cbind(Provinces, intermon_06, premon_06, mon_06, postmon_06,intermon_07, premon_07, mon_07, postmon_07,  intermon_08, premon_08, mon_08, postmon_08, intermon_09, premon_09, mon_09, postmon_09, intermon_10, premon_10, mon_10, postmon_10, intermon_11, premon_11, mon_11, postmon_11, intermon_12, premon_12, mon_12, postmon_12, intermon_13, premon_13, mon_13, postmon_13, intermon_14, premon_14, mon_14, postmon_14, intermon_15, premon_15, mon_15, postmon_15, intermon_16, premon_16, mon_16, postmon_16, intermon_17, premon_17, mon_17, postmon_17, intermon_18, premon_18, mon_18, postmon_18, intermon_19, premon_19, mon_19, postmon_19, intermon_20, premon_20, mon_20, postmon_20)

#turn it into a dataframe
dengue_mon <- as.data.frame(dengue_mon)

#turn the character columns into numeric values 
numColumns1 <- names(dengue_mon)[-which(names(dengue_mon) == "Provinces")]
dengue_mon[numColumns1] <- lapply(dengue_mon[numColumns1], as.numeric)
str(dengue_mon)
```

save the new dataframe that contains the normalized dengue cases merged into monsoon seasons

```{r}
filePath = paste(dirname(getwd()), "dengue_data", "dengue_mon.csv",sep = "/")
write.csv(dengue_mon, filePath)
```

Now that the data is in a format we can work with let's finally look at the correlation between humidity and dengue cases

```{r}
#example for monsoon season 2006
cor(dengue_mon$mon_06, humidity_mon$monH_06, use = "pairwise.complete.obs")

#let's try the correlation throughout all years 
##let's stack the data for the years atop of each other for easier calculation with the cor() function 

#start with the inter-monsoon dengue cases
stacked_intermon <- c(intermon_06, intermon_07, intermon_09, intermon_10, intermon_11, intermon_12, intermon_13, intermon_14, intermon_15, intermon_16, intermon_17, intermon_18, intermon_19, intermon_20)

#next pre-monsoon dengue cases 
stacked_premon <- c(premon_06, premon_07, premon_09, premon_10, premon_11, premon_12, premon_13, premon_14, premon_15, premon_16, premon_17, premon_18, premon_19, premon_20)

#next monsoon dengue cases 
stacked_mon <- c(mon_06, mon_07, mon_09, mon_10, mon_11, mon_12, mon_13, mon_14, mon_15, mon_16, mon_17, mon_18, mon_19, mon_20)

#next post-monsoon dengue cases 
stacked_postmon <- c(postmon_06, postmon_07, postmon_09, postmon_10, postmon_11, postmon_12, postmon_13, postmon_14, postmon_15, postmon_16, postmon_17, postmon_18, postmon_19, postmon_20)


##let's do the same for the humidity monsoon seasons 
#inter-monsoon seasons 
stacked_intermonH <- c(intermonH_06, intermonH_07, intermonH_09, intermonH_10, intermonH_11, intermonH_12, intermonH_13, intermonH_14, intermonH_15, intermonH_16, intermonH_17, intermonH_18, intermonH_19, intermonH_20)

#pre-monsoon seasons
stacked_premonH <- c(premonH_06, premonH_07, premonH_09, premonH_10, premonH_11, premonH_12, premonH_13, premonH_14, premonH_15, premonH_16, premonH_17, premonH_18, premonH_19, premonH_20)

#monsoon seasons 
stacked_monH <-c(monH_06, monH_07, monH_09, monH_10, monH_11, monH_12, monH_13, monH_14, monH_15, monH_16, monH_17, monH_18, monH_19, monH_20)

#post-monsoon seasons 
stacked_postmonH <- c(postmonH_06, postmonH_07, postmonH_09, postmonH_10, postmonH_11, postmonH_12, postmonH_13, postmonH_14, postmonH_15, postmonH_16, postmonH_17, postmonH_18, postmonH_19, postmonH_20)

##we need the provinces as well
#sort the provinces alphabetically
Provinces <- sort(Provinces)
#stack them 14 times 
stacked_provinces <- c(Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces,Provinces)

#make another dataframe to calculate the correlation 
stacked_hum_den <- cbind(stacked_provinces, stacked_intermon, stacked_premon, stacked_mon, stacked_postmon, stacked_intermonH, stacked_premonH, stacked_monH, stacked_postmonH)

#rename the coloumns 
colnames(stacked_hum_den) <- c("Provinces", "intermon", "premon", "mon", "postmon", "intermonH", "premonH", "monH", "postmonH")

#make it into a dataframe 
stacked_hum_den <- as.data.frame(stacked_hum_den)

#convert it to numeric 
numColumns2 <- names(stacked_hum_den)[-which(names(stacked_hum_den) == "Provinces")]
stacked_hum_den[numColumns2] <- lapply(stacked_hum_den[numColumns2], as.numeric)
str(stacked_hum_den)

#let's save this dataframe as well 
filePath1 <- paste(dirname(getwd()), "Correlation_hypothesis", "mon_dengue_hum.csv",sep = "/")
write.csv(stacked_hum_den, filePath1)  
```
